### 2024-02-05 outputs/paper/paper.qmd
Added in paper title and authors

---
title: "Why did the US fare so poorly during COVID-19?"
author: 
  - Adrian Ly
  - Sakhil Goel
  - Hannah Yu
thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-02-08 outputs/paper/paper.qmd
updated LLM usage and paper

@@
 # Data {#sec-data}
+
+## Data Measurement
 
@@
 
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
 # Results
@@
 
-## First discussion point {#sec-first-point}
+## Influence of political polarization on adherence to health guidelines.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Impact of government transparency and consistent communication on public trust.
 
-## Second discussion point
+## Role of social vulnerabilities and healthcare access disparities in pandemic impact.
 
-## Third discussion point
+## Strategies for improving real-time data collection and sharing for public health decisions.
 

### 2024-02-10 scripts/00-simulate_data.R
updated references and paper

@@ -1,15 +1,22 @@
 #### Preamble ####
 # Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Adrian Ly
+# Date: 9 February 2024 
+# Contact: adrian.ly@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
 # [...UPDATE THIS...]
 
 #### Simulate data ####

### 2024-02-10 scripts/01-download_data.R

@@ -1,18 +1,21 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Adrian Ly
+# Date: 9 February 2024 
+# Contact: adrian.ly@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
 
 #### Workspace setup ####
-library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
-
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
 #### Download data ####
 # [...ADD CODE HERE TO DOWNLOAD...]
 

### 2024-02-10 scripts/02-data_cleaning.R

@@ -1,14 +1,22 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Adrian Ly
+# Date: 9 February 2024 
+# Contact: adrian.ly@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+
 
 #### Workspace setup ####
 library(tidyverse)
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
 
 #### Clean data ####
 raw_data <- read_csv("inputs/data/plane_data.csv")

### 2024-02-10 scripts/03-test_data.R

@@ -1,15 +1,22 @@
 #### Preamble ####
 # Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Adrian Ly
+# Date: 9 February 2024 
+# Contact: adrian.ly@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
 
 #### Workspace setup ####
 library(tidyverse)
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
+
 # [...UPDATE THIS...]
 
 #### Test data ####

### 2024-02-10 outputs/paper/paper.qmd
updated references and paper
upadated llm usage and paper

@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 
+#### Workspace setup ####
 library(tidyverse)
-library(palmerpenguins)
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
+library(scales)
+
+#### Reading files ####
+
+ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
 ```
@@
 
-You can and should cross-reference sections and sub-sections.
-
-The remainder of this paper is structured as follows. @sec-data....
+*This reproduction was performed after a replication on the Social Science Reproduction platform: ****[link here]()***
 
@@
 
+## Source
+
+The datasets utilized throughout this paper was obtained from the original paper.
+
+## Methodology
+
 ## Data Measurement
+
+## Data cleaning
 
@@
 #| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
 #| label: tbl-modelresults
@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+ghs_data <- ghs_data %>%
+  mutate(COUNTRY = case_when(
+    Country == "United States of America" ~ "United States",
+    Country == "Bolivia (Plurinational State of)" ~ "Bolivia",
+    # Add other necessary name changes here
+    TRUE ~ Country
+  ))
+
+world_ghs <- merge(world_map, ghs_data, by.x = "COUNTRY", by.y = "COUNTRY", all.x = TRUE)
+
+ggplot(data = world_ghs) +
+  geom_sf(aes(fill = `OVERALL SCORE`), color = "white") +
+  scale_fill_distiller(palette = "Spectral", na.value = "lightgrey", 
+                       name = "GHS Index Score", direction = 1,
+                       limits = c(min(ghs_data$`OVERALL SCORE`, na.rm = TRUE), max(ghs_data$`OVERALL SCORE`, na.rm = TRUE)),
+                       labels = comma) +  # Use comma to format numbers with commas
+  labs(title = "Global Health Security Index Scores by Country") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5))
+
+# Save the plot to a file
+ggsave("GHS_Index_Map.png", width = 11, height = 8)
 ```

### 2024-02-10 outputs/paper/paper.qmd
updated paper

@@
 ---
-title: "Why did the US fare so poorly during COVID-19?"
+title: "How did local health infrastructure and socio-political factors within different states and counties in the United States affect the disparities in COVID-19 outcomes, and what lessons can be learned for more targeted public health preparedness and response strategies in future pandemics?"
 author: 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
 # Results
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in @fig-GHS.
 
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: fig-GHS
+#| fig-cap: Global Health Security Index Scores by Country
 #| warning: false
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
-\newpage
-
 # References

### 2024-02-10 outputs/paper/paper.qmd
added references

@@
 
-The datasets utilized throughout this paper was obtained from the original paper.
+The datasets utilized throughout this paper was obtained from the `original paper` [@citearticle].
 
 ## Methodology
+
+`R` [@citeR] was the language and enviroment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], and `scales` [@citeScales].
 

### 2024-02-10 outputs/paper/paper.qmd
Made figure 2 and 3

--- 2024-02-10 outputs/paper/paper.qmd
@@
 #| message: false
-
 #### Workspace setup ####
@@
 world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
 ```
@@
 
+```{r fig.width=7, fig.height=3.5}
+#| echo: false
+#| eval: true
+#| label: fig-ELE
+#| fig-cap: Estimates of Life Expectancy at Birth, by Race 2006-2021
+#| warning: false
+
+# Reshape the data if necessary
+life_exp_melted <- reshape(
+  life_exp,
+  idvar = "year_id",
+  varying = list(names(life_exp)[-1]), 
+  v.names = "life_exp",
+  times = names(life_exp)[-1],
+  new.row.names = 1:1e5,
+  direction = "long"
+)
+life_exp_melted$race <- gsub("_", " ", life_exp_melted$time)  # Convert "_" to " " in race names
+life_exp_melted <- life_exp_melted[, c("year_id", "race", "life_exp")]
+
+life_exp_melted[life_exp_melted$race == 'All', "race"] <- 'All races and origins'
+
+## CREATE PANEL A
+est_life <- ggplot(data = life_exp_melted[!(life_exp_melted$race %like% 'AIAN|Asian'), ], aes(x = year_id, y = life_exp, color = race)) +
+  geom_point(shape = 16, size = 3.5, alpha = 0.9) +  # Circular points with size 2
+  geom_line(aes(group = race), alpha = 0.56, size = 1.5) +  # Add group aesthetic for lines
+  scale_y_continuous(name = 'Life expectancy at birth', limits = c(65, 85)) +
+  scale_x_continuous(name = '', breaks = seq(2006, 2020, by = 2), limits = c(2005.5, 2021.5)) +
+  scale_color_manual(values = c('#0c4928', '#0b3669', '#4c2962', '#5a0f1c'), 
+                     labels = c("All races and origins", "Hispanic", "Black", "White")) +
+  # Darker color values
+  theme_minimal() +
+  theme(axis.title = element_text(size = 14),  # Adjusting the size of the y-axis label
+        axis.text = element_text(size = 12),
+        legend.position = 'bottom',  # Set legend position to bottom
+        legend.justification = 'center',  # Center the legend
+        legend.text = element_text(size = 10),  # Adjusting the font size for legend text
+        legend.title = element_text(size = 12, face = 'bold')) +
+  labs(color = "Race:")  # Set the legend title
+
+est_life
+# Save the plot to a file
+ggsave("Life_Exp.png", width = 10, height = 8)
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+# Create a dataframe to store the changes by race
+change_by_race <- data.frame(
+  race = c("All races and origins", "Hispanic", "American Indians and Alaska Native", "Asian", "Black", "White"),
+  change_2020_2019 = c(77 - 78.8, 77.9 - 81.9, 67.1 - 71.8, 83.6 - 85.6, 71.5 - 74.8, 77.4 - 78.8),
+  change_2021_2020 = c(76.1 - 77, 77.7 - 77.9, 65.2 - 67.1, 83.5 - 83.6, 70.8 - 71.5, 76.4 - 77.4)
+)
+
+```
+
+```{r fig.width=8, fig.height=4}
+#| echo: false
+#| eval: true
+#| label: fig-CLE
+#| fig-cap: Change in Life Expectancy at Birth from the Previous Year
+#| warning: false
+
+# Melt the dataframe
+tmp <- melt(change_by_race, id.vars = "race", measure.vars = c("change_2020_2019", "change_2021_2020"), 
+             variable.name = "period", value.name = "diff")
+
+# Create PANEL B
+change_life_exp <- ggplot(data = tmp, aes(x = race, y = diff, fill = period)) +
+  geom_bar(stat = "identity", position = "dodge", alpha = 0.71) +
+  scale_y_continuous(name = "Change (years)", breaks = -5:0, limits = c(-5.5, 0)) +
+  scale_fill_manual(name = "Time period", values = c('#243f29', '#186d38'), labels = c('2019-2020', '2020-2021')) + 
+  scale_color_manual(name = "Time period", values = c('#243f29', '#186d38'), labels = c('2019-2020', '2020-2021')) + 
+  theme_minimal() + labs(x = '') +
+  geom_text(aes(label = sprintf("%.1f", diff)), position = position_dodge(width = 0.9), vjust = 1.3, size = 4) +
+  theme(
+    axis.title = element_text(size = 14),
+    axis.text.x = element_text(size = 10, hjust = 1),
+    axis.text.y = element_text(size = 8),
+    legend.text = element_text(size = 12),
+    legend.title = element_text(size = 12, face = 'bold')) + 
+    scale_x_discrete(labels = function(x) str_wrap(x, width = 15))
+
+change_life_exp
+ggsave("Change_Life_Exp.png", width = 10, height = 8)
+
+```
+
 # Discussion

### 2024-02-10 outputs/paper/paper.qmd
updated paper

--- 2024-02-10 outputs/paper/paper.qmd
@@
 
-The datasets utilized throughout this paper was obtained from the original paper.
+The datasets utilized throughout this paper was obtained from the `original paper` [@citearticle].
 
 ## Methodology
+
+`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], and `scales` [@citeScales].
 
@@
                        labels = comma) +  # Use comma to format numbers with commas
-  labs(title = "Global Health Security Index Scores by Country") +
   theme_minimal() +

### 2024-02-11 scripts/00-simulate_data.R
create replication in scripts

@@ -1,26 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Adrian Ly
-# Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca
-# License: MIT
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(sf)
-library(readxl)
-library(knitr)
-library(janitor)
-library(lubridate)
-library(dplyr)
-library(data.table)
-library(RColorBrewer)
-library(ggpubr)
-# [...UPDATE THIS...]
-
-#### Simulate data ####
-# [...ADD CODE HERE...]
-
-
-

### 2024-02-11 scripts/Replication.R

@@ -1,0 +1,131 @@
+#### Preamble ####
+# Purpose: Paper Replication
+# Author: Adrian Ly, Hannah Yu
+# Date: 9 February 2024 
+# Contact: adrian.ly@mail.utoronto.ca
+# License: MIT
+
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
+
+#### Reading files ####
+ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+
+## CREATE FIGURE 1
+ghs_data <- ghs_data %>%
+  mutate(COUNTRY = case_when(
+    Country == "United States of America" ~ "United States",
+    Country == "Bolivia (Plurinational State of)" ~ "Bolivia",
+    # Add other necessary name changes here
+    TRUE ~ Country
+  ))
+
+
+#| Fig-cap:  Global Health Security Index Scores by Country
+
+world_ghs <- merge(world_map, ghs_data, by.x = "COUNTRY", by.y = "COUNTRY", all.x = TRUE)
+
+
+ggplot(data = world_ghs) +
+  geom_sf(aes(fill = `OVERALL SCORE`), color = "white") +
+  scale_fill_distiller(palette = "Spectral", na.value = "lightgrey", 
+                       name = "GHS Index Score", direction = 1,
+                       limits = c(min(ghs_data$`OVERALL SCORE`, na.rm = TRUE), max(ghs_data$`OVERALL SCORE`, na.rm = TRUE)),
+                       labels = comma) +  # Use comma to format numbers with commas
+  labs(title = "Global Health Security Index Scores by Country") +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5))
+
+# Save the plot to a file
+ggsave("GHS_Index_Map.png", width = 11, height = 8)
+
+#| fig-cap: Estimates of Life Expectancy at Birth, by Race 2006-2021
+
+# Reshape the data if necessary
+life_exp_melted <- reshape(
+  life_exp,
+  idvar = "year_id",
[…]

### 2024-02-11 outputs/paper/paper.qmd
added vote share graph

@@
 life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
 ```
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+
+vote_data_by_state <- vote_share %>%
+  group_by(state_name) %>%
+  summarize(
+    total_votes = sum(total_votes),
+    total_gop_votes = sum(votes_gop),
+    total_dem_votes = sum(votes_dem),
+    total_diff = sum(diff),
+    avg_per_gop = mean(per_gop),
+    avg_per_dem = mean(per_dem),
+    avg_per_point_diff = mean(per_point_diff)
+  )
+
+vote_data_by_state <- vote_data_by_state |>
+  mutate(gop_greater = if_else(total_gop_votes > total_dem_votes, 1, 0))
+
+```
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+
+# Create plot summarizing total number of votes by top 10 republican state
+sorted_states <- vote_data_by_state[order(-vote_data_by_state$total_diff), ]
+
+# Extract the top 10 states with the biggest difference
+top_10_states <- head(sorted_states, 10)
+# Create a data frame for plotting
+plot_data <- data.frame(
+  state_name = top_10_states$state_name,
+  republican_votes = top_10_states$total_gop_votes,
+  democrat_votes = top_10_states$total_dem_votes
+)
+
+# Convert data to long format for ggplot
+plot_data_long <- tidyr::pivot_longer(plot_data, cols = c(republican_votes, democrat_votes), names_to = "party", values_to = "votes")
+
+# Create the stacked bar plot
+ggplot(plot_data_long, aes(x = state_name, y = votes, fill = party)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Total Number of Votes by Top 10 Republican State", x = "State", y = "Total Votes", fill = "Party") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
+  scale_fill_manual(values = c("republican_votes" = "red", "democrat_votes" = "blue"),
+                    labels = c("Republican Votes", "Democratic Votes"))
+
+```
+
 # Discussion

### 2024-02-11 outputs/paper/paper.qmd
added ssrp link

--- 2024-02-10 outputs/paper/paper.qmd
@@
 
-*This reproduction was performed after a replication on the Social Science Reproduction platform: ****[link here]()***
+*This reproduction was performed after a replication on the Social Science Reproduction platform: ****[link here](https://www.socialsciencereproduction.org/reproductions/c35e8e98-762d-4c00-a1a3-544ed0b2008d/index)***
 

### 2024-02-11 outputs/paper/paper.qmd
updated paper

@@
 
+This begs the question as to why we are seeing these results. There isn't exactly a single answer to this question, however we can certainly point out some considerable factors to this result.
+
 ## Influence of political polarization on adherence to health guidelines.
+
+Political polarization has significantly impacted the adherence to health guidelines during the COVID-19 pandemic.
+The divergence in political ideologies has translated into differing attitudes towards health directives, including mask mandates, social distancing, and vaccination uptake.
+Various studies and our own results have shown that areas with higher support for one political party exhibited distinct behaviors and compliance levels with health recommendations, which directly correlated with COVID-19 case rates and mortality.
+An news article from `ABC News` [@citesource1] shows that the top states with the highest COVID-19 deaths are Arizona, and Washington with 581 deaths and 526 deaths respectively per 100,000 people.
+According to 2020 presidential voting data published by `CNN`, we have both states having the electoral vote of democrat with Washington wining by 58% [@citesource2] and Arizona winning by 49.4% [@citesource3].
+Another news article by `ContagionLive` [@citesource4] also makes the claim of both Arizona and Washington having the highest COVID-19 mortality.
+This polarization has not only influenced individual behavior but also shaped state and local health policies, further entrenching the disparities in health outcomes.
 

### 2024-02-11 outputs/paper/paper.qmd
updated discussion and llm usage

@@
 This polarization has not only influenced individual behavior but also shaped state and local health policies, further entrenching the disparities in health outcomes.
+The adherence to health guidelines are evident in the varied health outcomes observed across the United States.
+Regions with lower compliance to health directives, often influenced by political leanings, have experienced higher rates of COVID-19 transmission, hospitalizations, and deaths.
+The disparities in vaccine uptake, driven by political affiliations, have further exacerbated these outcomes, leaving certain communities more vulnerable to the virus and its variants.
+In order to mitigate the influence of political polarization on public health, it is imperative to depoliticize health guidelines and focus on evidence-based approaches to disease prevention and control.
+Building trust in health institutions and promoting bipartisan support for public health measures are essential steps towards achieving higher compliance and better health outcomes.
+Engaging trusted community leaders and utilizing targeted communication strategies can also help bridge the divide and encourage adherence to health guidelines.
 
 ## Impact of government transparency and consistent communication on public trust.
+
+The politicization of health guidelines and mixed messages from political and health leaders during the COVID-19 pandemic have significantly undermined the effectiveness of public health messaging, leading to confusion, skepticism, and eroded trust among the public.
+Initially, inconsistencies in recommendations, such as on mask usage, challenged the principle of clear, consistent, and science-based communication essential for an effective public health response.
+Moreover, the transparency of government actions and decision-making processes is crucial in building and maintaining public trust, especially during health crises.
+The level of public trust was greatly affected by the openness and accuracy with which governments, at all levels, communicated about the evolving situation, the reasoning behind guidelines, and the measures taken to combat the virus, emphasizing the importance of transparent reporting of data related to case counts, hospitalizations, vaccine distribution, and side effects.
+Furthermore, consistent communication from public health officials and government leaders is key to ensuring adherence to health guidelines, where inconsistencies, such as changes in mask-wearing guidelines without clear explanations, have led to public confusion.
+The direct correlation between government transparency, consistent communication, and public behavior is self-evident, with populations receiving clear and transparent information being more likely to adhere to guidelines, participate in testing and tracing efforts, and accept vaccination.
+Drawing lessons from the pandemic, strategies for improving government transparency and communication in future health emergencies should include establishing centralized information hubs, ensuring regular and predictable communication from health authorities, engaging community leaders in information dissemination, and harnessing digital platforms and social media to amplify public health messages, thus reinforcing public trust and compliance.
 

### 2024-02-11 outputs/paper/paper.qmd
- updated replication and paper - deleted unecessary files - changed file names

@@
 
-# Save the plot to a file
-ggsave("GHS_Index_Map.png", width = 11, height = 8)
 ```
@@
 est_life
-# Save the plot to a file
-ggsave("Life_Exp.png", width = 10, height = 8)
+
 ```
@@
 change_life_exp
-ggsave("Change_Life_Exp.png", width = 10, height = 8)
 
@@
 ```
+
 ```{r}
@@
 #| eval: true
+#| label: fig-Vote
+#| fig-cap: Total Number of Votes by Top 10 Republican State
 #| warning: false

### 2024-02-11 scripts/01-simulate_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
+# Purpose: Simulates... [...UPDATE THIS...]
 # Author: Adrian Ly
 # Date: 9 February 2024 
 # Contact: adrian.ly@mail.utoronto.ca
@@ -16,14 +16,8 @@ library(dplyr)
 library(data.table)
 library(RColorBrewer)
 library(ggpubr)
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+#### Simulate data ####
+# [...ADD CODE HERE...]
 
          

### 2024-02-11 scripts/02-data_cleaning.R

@@ -1,52 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Adrian Ly
-# Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca
-# License: MIT
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(sf)
-library(readxl)
-library(knitr)
-library(janitor)
-library(lubridate)
-library(dplyr)
-library(data.table)
-library(RColorBrewer)
-library(ggpubr)
-
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
-
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
-
-#### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")

### 2024-02-11 scripts/Replication.R

@@ -6,7 +6,6 @@
 # License: MIT
 
 
-
 #### Workspace setup ####
 library(tidyverse)
 library(sf)
@@ -23,6 +22,7 @@ library(ggpubr)
 ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
 world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
 life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
 
 ## CREATE FIGURE 1
 ghs_data <- ghs_data %>%
@@ -33,9 +33,6 @@ ghs_data <- ghs_data %>%
     TRUE ~ Country
   ))
 
-
-#| Fig-cap:  Global Health Security Index Scores by Country
-
 world_ghs <- merge(world_map, ghs_data, by.x = "COUNTRY", by.y = "COUNTRY", all.x = TRUE)
 
 
@@ -45,15 +42,14 @@ ggplot(data = world_ghs) +
                        name = "GHS Index Score", direction = 1,
                        limits = c(min(ghs_data$`OVERALL SCORE`, na.rm = TRUE), max(ghs_data$`OVERALL SCORE`, na.rm = TRUE)),
                        labels = comma) +  # Use comma to format numbers with commas
-  labs(title = "Global Health Security Index Scores by Country") +
   theme_minimal() +
   theme(legend.position = "bottom",
         plot.title = element_text(hjust = 0.5))
 
 # Save the plot to a file
-ggsave("GHS_Index_Map.png", width = 11, height = 8)
+ggsave("outputs/data/GHS_Index_Map.png", width = 11, height = 8)
 
-#| fig-cap: Estimates of Life Expectancy at Birth, by Race 2006-2021
+## CREATE FIGURE 2
 
 # Reshape the data if necessary
 life_exp_melted <- reshape(
@@ -70,7 +66,7 @@ life_exp_melted <- life_exp_melted[, c("year_id", "race", "life_exp")]
 
 life_exp_melted[life_exp_melted$race == 'All', "race"] <- 'All races and origins'
 
-## CREATE FIGURE 2 PANEL A
+## CREATE PANEL A
 est_life <- ggplot(data = life_exp_melted[!(life_exp_melted$race %like% 'AIAN|Asian'), ], aes(x = year_id, y = life_exp, color = race)) +
[…]

### 2024-02-11 outputs/paper/paper.qmd
simulated data and tests

@@
 library(scales)
+library(here)
 
@@
 
-ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+#ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+#world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+#life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+#vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
+world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
+life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
+vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
+
 ```

### 2024-02-11 scripts/01-simulate_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Adrian Ly
+# Purpose: Simulate Data of Interest
+# Author: Adrian Ly, Hannah Yu
 # Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca
+# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca
 # License: MIT
 
 #### Workspace setup ####
@@ -18,6 +18,88 @@ library(RColorBrewer)
 library(ggpubr)
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
 
-         
+# Simulate the average life expectancy of people by ethnicity
+# Set seed for reproducibility
+set.seed(123)
+
+# Define the years
+years <- 2006:2021
+
+# Define life expectancy ranges for each group
+life_expectancy_ranges <- list(
+  "All" = c(77, 80),
+  "Hispanic" = c(80, 83),
+  "Non-Hispanic AIAN" = c(65, 72),
+  "Non-Hispanic Asian" = c(83, 87),
+  "Non-Hispanic Black" = c(70, 76),
+  "Non-Hispanic White" = c(77, 81)
+)
+
+# Generate the dataset
+data <- matrix(NA, nrow = length(years), ncol = length(life_expectancy_ranges) + 1)
+colnames(data) <- c("year_id", names(life_expectancy_ranges))
+data[, 1] <- years
+
+for (i in 1:length(years)) {
+  for (group in names(life_expectancy_ranges)) {
+    if (group == "All") {
+      life_expectancy <- runif(1, life_expectancy_ranges[[group]][1], life_expectancy_ranges[[group]][2])
+    } else {
+      life_expectancy <- runif(1, life_expectancy_ranges[[group]][1], life_expectancy_ranges[[group]][2])
+    }
+    data[i, which(colnames(data) == group)] <- round(life_expectancy, 1)
+  }
+}
+
+# Save the dataset to a CSV file
+write.csv(data, "simulated_life_data.csv", row.names = FALSE)
+
+# Print the dataset to the console
+print(data)
+
+
+# Simulate the Vote at Each State
+set.seed(123)
+# Define the state names
+state_names <- c(
+  "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", "Georgia",
[…]

### 2024-02-11 scripts/02-test_data.R

@@ -1,8 +1,8 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
+# Purpose: Tests the Simulated Data
 # Author: Adrian Ly
 # Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca
+# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca
 # License: MIT
 
 #### Workspace setup ####
@@ -17,6 +17,61 @@ library(data.table)
 library(RColorBrewer)
 library(ggpubr)
 
-# [...UPDATE THIS...]
-
 #### Test data ####
+## Test simulated_life_data
+# Load the dataset generated in R
+data1 <- read.csv("outputs/simulated_life_exp.csv", stringsAsFactors = FALSE)
+
+# 1. Data Structure Check
+# Check the number of rows and columns
+structure_check <- dim(data)[1] == length(years) & dim(data)[2] == length(life_expectancy_ranges) + 1
+print(paste("Data Structure Check:", ifelse(structure_check, "Passed", "Failed")))
+
+# 2. Range Check
+# Define the expected ranges
+expected_ranges <- c(77:80, 80:83, 65:72, 83:87, 70:76, 77:81)
+# Check if all values fall within the expected ranges
+range_check <- all(sapply(data[, -1], function(x) all(x >= min(expected_ranges) & x <= max(expected_ranges))))
+print(paste("Range Check:", ifelse(range_check, "Passed", "Failed")))
+
+# 3. Numeric Check
+# Check if all values are numeric
+numeric_check <- all(sapply(data, is.numeric))
+print(paste("Numeric Check:", ifelse(numeric_check, "Passed", "Failed")))
+
+## Test simulated_vote_data
+# Load the dataset generated in R
+data2 <- read.csv("outputs/simulated_vote_data.csv", stringsAsFactors = FALSE)
+
+# Test 1: Check number of rows and columns
+num_rows <- nrow(data2)
+num_cols <- ncol(data2)
+
+expected_num_rows <- 50 # Expecting 50 states
+expected_num_cols <- 8  # Expecting 7 columns (plus 1 for state_name)
+
+if (num_rows == expected_num_rows && num_cols == expected_num_cols) {
+  print("Test 1 passed: Correct number of rows and columns.")
+} else {
+  print("Test 1 failed: Incorrect number of rows or columns.")
+}
+
+# Test 2: Check for missing values
+if (!anyNA(data2)) {
[…]

### 2024-02-12 outputs/paper/paper.qmd
add reference added tables of covid deaths added data source

--- 2024-02-11 outputs/paper/paper.qmd
@@
 toc: true
+prefer-html: true
 bibliography: references.bib
@@
 library(scales)
+library(here)
+library(readxl)
+library(kableExtra)
 
@@
 
-ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+#ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+#world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+#life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+#vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
+world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
+life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
+vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
+deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
+
 ```
@@
 
-The datasets utilized throughout this paper was obtained from the `original paper` [@citearticle].
+The datasets utilized in this paper were mainly obtained from the 'original paper' [@citearticle].
+Additionally, to address the original paper's lack of US Covid statistics and political party support data, we incorporated information from @citedata1 and @citedata2.
+
+@citedata1 discusses the long-term impacts of COVID-related school closures. From this source, we utilized the dataset on voting shares during the 2020 election by county.
+@citedata2 provides summaries of COVID-19 death rates in the United States as of March 2023, organized by state.
+Analyzing results from both datasets allows us to explore the relationship between political affiliation and COVID-19 outcomes.Our reproduction aims to fill these gaps and also includes tables and graphs that were not presented in the original paper to support our findings.
 
@@
 
-# Results
-
-Our results are summarized in @fig-GHS.
+## Data Visualization
 
@@
 ```
+
+# Results
+
+Our results are summarized in @fig-GHS.
 
@@
 #| label: fig-Vote
-#| fig-cap: Total Number of Votes by Top 10 Republican State
+#| fig-cap: Total Number of Votes in Top 10 States with Highest Proportion of Republican Votes 
 #| warning: false
@@
 # Create plot summarizing total number of votes by top 10 republican state
-sorted_states <- vote_data_by_state[order(-vote_data_by_state$total_diff), ]
+#sorted_states <- vote_data_by_state[order(-vote_data_by_state$total_diff), ]
+
+# Calculate proportion of Republican votes
+sorted_states <- vote_data_by_state
+sorted_states$proportion_republican <- sorted_states$total_gop_votes / sorted_states$total_votes
+
+# Sort by proportion of Republican votes
+sorted_states <- sorted_states[order(-sorted_states$proportion_republican), ]
+
+# Select top 10 states
+top_10_states <- head(sorted_states, 10)
 
@@
   scale_fill_manual(values = c("republican_votes" = "red", "democrat_votes" = "blue"),
-                    labels = c("Republican Votes", "Democratic Votes"))
+                    labels = c("Democratic Votes", "Republican Votes"))
 
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl- deaths
+#| tbl-cap: Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)
+#| warning: false
+deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
+
+# Assign column names
+colnames(deaths) <- c("State", "Death_Rate")
+
+# Rank the states based on death rate
+deaths <- deaths[order(-deaths$Death_Rate), ]
+deaths$Rank <- 1:nrow(deaths)
+
+# Create the table
+table <- head(data.frame(Rank = deaths$Rank, State = deaths$State, Deaths = deaths$Death_Rate), 10)
+
+# Print the table using kableExtra
+kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c", format = "html") %>%
+  kable_styling(full_width = FALSE)
 ```
@@
 The divergence in political ideologies has translated into differing attitudes towards health directives, including mask mandates, social distancing, and vaccination uptake.
+
 Various studies and our own results have shown that areas with higher support for one political party exhibited distinct behaviors and compliance levels with health recommendations, which directly correlated with COVID-19 case rates and mortality.
@@
 This polarization has not only influenced individual behavior but also shaped state and local health policies, further entrenching the disparities in health outcomes.
+
 The adherence to health guidelines are evident in the varied health outcomes observed across the United States.

### 2024-02-12 outputs/paper/paper.qmd
fixed paper and folders

@@
 toc: true
-prefer-html: true
 bibliography: references.bib
@@
 #| message: false
+
 #### Workspace setup ####
@@
 library(here)
-library(readxl)
 library(kableExtra)
@@
 
-#ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-#world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-#life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-#vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
-ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
-world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
-life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
-vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
-deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
+ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
 
@@
 
-`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], and `scales` [@citeScales].
+`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], and `scales` [@citeScales].
 
@@
 #| eval: true
-#| label: tbl- deaths
+#| label: tbl-deaths
 #| tbl-cap: Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)
 #| warning: false
-deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
 
@@
 # Print the table using kableExtra
-kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c", format = "html") %>%
+kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c") %>%
   kable_styling(full_width = FALSE)
+
 ```
@@
 
-## Weaknesses and next steps
+## Weaknesses and next stepsa
 

### 2024-02-12 scripts/01-simulate_data.R

@@ -53,7 +53,7 @@ for (i in 1:length(years)) {
 }
 
 # Save the dataset to a CSV file
-write.csv(data, "simulated_life_data.csv", row.names = FALSE)
+write.csv(data, "outputs/data/simulated_life_exp.csv", row.names = FALSE)
 
 # Print the dataset to the console
 print(data)
@@ -95,7 +95,7 @@ simulated_vote_df$state_name <- state_names
 print(simulated_vote_df)
 
 # Define file path
-file_path <- "simulated_vote_data.csv"
+file_path <- "outputs/data/simulated_vote_data.csv"
 
 # Write simulated vote data to CSV
 write.csv(simulated_vote_df, file = file_path, row.names = FALSE)

### 2024-02-12 scripts/02-test_data.R

@@ -20,7 +20,7 @@ library(ggpubr)
 #### Test data ####
 ## Test simulated_life_data
 # Load the dataset generated in R
-data1 <- read.csv("outputs/simulated_life_exp.csv", stringsAsFactors = FALSE)
+data1 <- read.csv("outputs/data/simulated_life_exp.csv", stringsAsFactors = FALSE)
 
 # 1. Data Structure Check
 # Check the number of rows and columns
@@ -41,7 +41,7 @@ print(paste("Numeric Check:", ifelse(numeric_check, "Passed", "Failed")))
 
 ## Test simulated_vote_data
 # Load the dataset generated in R
-data2 <- read.csv("outputs/simulated_vote_data.csv", stringsAsFactors = FALSE)
+data2 <- read.csv("outputs/data/simulated_vote_data.csv", stringsAsFactors = FALSE)
 
 # Test 1: Check number of rows and columns
 num_rows <- nrow(data2)

### 2024-02-12 outputs/paper/paper.qmd
fixed file paths and paper

@@
 
-The datasets utilized in this paper were mainly obtained from the 'original paper' [@citearticle].
+The datasets utilized in this paper were mainly obtained from the `original paper` [@citearticle].
 Additionally, to address the original paper's lack of US Covid statistics and political party support data, we incorporated information from @citedata1 and @citedata2.
@@
 
-`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], and `scales` [@citeScales].
+`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], and `scales` [@citeScales].
 
@@
 
+\newpage
+
 # Discussion

### 2024-02-12 scripts/01-simulate_data.R

@@ -16,6 +16,9 @@ library(dplyr)
 library(data.table)
 library(RColorBrewer)
 library(ggpubr)
+library(scales)
+library(here)
+library(kableExtra)
 
 #### Simulate data ####
 

### 2024-02-12 scripts/02-test_data.R

@@ -16,6 +16,9 @@ library(dplyr)
 library(data.table)
 library(RColorBrewer)
 library(ggpubr)
+library(scales)
+library(here)
+library(kableExtra)
 
 #### Test data ####
 ## Test simulated_life_data

### 2024-02-12 scripts/Replication.R

@@ -17,12 +17,18 @@ library(dplyr)
 library(data.table)
 library(RColorBrewer)
 library(ggpubr)
+library(scales)
+library(here)
+library(kableExtra)
+library(webshot)
+library(webshot2)
 
 #### Reading files ####
 ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
 world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
 life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
 vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
 
 ## CREATE FIGURE 1
 ghs_data <- ghs_data %>%
@@ -159,3 +165,24 @@ ggplot(plot_data_long, aes(x = state_name, y = votes, fill = party)) +
 
 ggsave("outputs/data/Voter_Data.png", width = 10, height = 8)
 
+
+## CREATE TABLE 1
+
+# Assign column names
+colnames(deaths) <- c("State", "Death_Rate")
+
+# Rank the states based on death rate
+deaths <- deaths[order(-deaths$Death_Rate), ]
+deaths$Rank <- 1:nrow(deaths)
+
+# Create the table
+table <- head(data.frame(Rank = deaths$Rank, State = deaths$State, Deaths = deaths$Death_Rate), 10)
+
+# Print the table using kableExtra
+kable_img <- kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c") %>%
+  kable_styling(full_width = FALSE)
+
+save_kable(kable_img, "outputs/data/table.html")
+
+# Capture the table as an image
+webshot::webshot("outputs/data/table.html", "outputs/data/table.png")

### 2024-02-12 scripts/Replication.R
fixed replication

@@ -53,7 +53,7 @@ ggplot(data = world_ghs) +
         plot.title = element_text(hjust = 0.5))
 
 # Save the plot to a file
-ggsave("outputs/data/GHS_Index_Map.png", width = 11, height = 8)
+ggsave("outputs/graphs/GHS_Index_Map.png", width = 11, height = 8)
 
 ## CREATE FIGURE 2
 
@@ -91,7 +91,7 @@ est_life <- ggplot(data = life_exp_melted[!(life_exp_melted$race %like% 'AIAN|As
   labs(color = "Race:")  # Set the legend title
 
 # Save the plot to a file
-ggsave("outputs/data/Life_Exp.png", width = 10, height = 8)
+ggsave("outputs/graphs/Life_Exp.png", width = 10, height = 8)
 
 ## CREATE FIGURE 3
 
@@ -120,7 +120,7 @@ change_life_exp <- ggplot(data = tmp, aes(x = race, y = diff, fill = period)) +
     legend.title = element_text(size = 12, face = 'bold')) + 
   scale_x_discrete(labels = function(x) str_wrap(x, width = 15))
 
-ggsave("outputs/data/Change_Life_Exp.png", width = 10, height = 8)
+ggsave("outputs/graphs/Change_Life_Exp.png", width = 10, height = 8)
 
 ## CREATE FIGURE 4
 
@@ -163,7 +163,7 @@ ggplot(plot_data_long, aes(x = state_name, y = votes, fill = party)) +
   scale_fill_manual(values = c("republican_votes" = "red", "democrat_votes" = "blue"),
                     labels = c("Republican Votes", "Democratic Votes"))
 
-ggsave("outputs/data/Voter_Data.png", width = 10, height = 8)
+ggsave("outputs/graphs/Voter_Data.png", width = 10, height = 8)
 
 
 ## CREATE TABLE 1
@@ -182,7 +182,8 @@ table <- head(data.frame(Rank = deaths$Rank, State = deaths$State, Deaths = deat
 kable_img <- kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c") %>%
   kable_styling(full_width = FALSE)
 
-save_kable(kable_img, "outputs/data/table.html")
+save_kable(kable_img, "outputs/graphs/table.html")
 
 # Capture the table as an image
-webshot::webshot("outputs/data/table.html", "outputs/data/table.png")
+webshot::webshot("outputs/graphs/table.html", "outputs/graphs/table.png")
+

### 2024-02-12 outputs/paper/paper.qmd
updated paper

@@
 
+The COVID-19 pandemic starkly highlighted how social vulnerabilities and disparities in healthcare access exacerbated the impact of global health crises, contributing to significant variations in disease outcomes and underscoring the need for targeted public health strategies that address these disparities' root causes.
+Social vulnerabilities, such as socioeconomic status, race, ethnicity, and housing conditions, critically determined COVID-19 outcomes' severity, with populations in crowded housing, limited access to sanitation, and lower socioeconomic brackets experiencing higher transmission rates due to social distancing and hygiene maintenance challenges.
+An independent study done by the `Government of Canada` states that it "identified that the risk of COVID-19 related deaths in Black, Asian and minority ethnic groups was nearly 1.5 times higher than White individuals" [@citesource5].
+Another news article done by the `MSN` [@citesource6] states that 'racial minority participants reported more negative impacts on health status, activity, and absence from work as compared to the White population.' The pandemic's economic toll further limited these groups' healthcare access, amplifying vulnerabilities.
+Disparities in healthcare access played a significant role in influencing COVID-19 morbidity and mortality, with communities facing healthcare facility shortages, provider scarcities, and barriers due to insurance or financial constraints at heightened risk.
+These disparities were evident in the uneven vaccine distribution and access, highlighting the advantages of regions with strong healthcare infrastructure.
+Marginalized populations, including racial and ethnic minorities, faced compounded risks from social vulnerabilities and healthcare disparities, evidenced by higher infection, hospitalization, and death rates due to factors like essential service employment and prevalent pre-existing conditions.
+A news article from `CNN` [@citesource7] talks about how many essential workers, who cannot work from home, are from black and Latinx communities.
+These include healthcare professionals, grocery cashiers, delivery workers, and public transport employees.
+Despite their crucial roles, they often lack adequate pay, protection, and respect.
+Addressing these disparities in future pandemics requires public health strategies that prioritize equity and inclusivity, including community-based healthcare investments, enhanced vulnerable community outreach, and equitable healthcare resource access policies.
+By incorporating social determinants of health into public health preparedness plans, responses can effectively protect at-risk populations, making future public health responses more resilient, inclusive, and effective in safeguarding all population segments.
+
 ## Strategies for improving real-time data collection and sharing for public health decisions.

### 2024-02-12 outputs/paper/paper.qmd
wrote results and other

--- 2024-02-12 outputs/paper/paper.qmd
@@
 library(kableExtra)
+library(here)
+library(gt)
 
 #### Reading files ####
+# Set the working directory to the correct location
+ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
+world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
+life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
+vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
+deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
 
-ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
-deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+# ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+# world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+# life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+# vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+# deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
 
@@
 ## Data cleaning
-
+rename columns
 ## Data Visualization
@@
 
-Our results are summarized in @fig-GHS.
+Our results are summarized in the following figures.
+@fig-ELE illustrates the trend of life expectancy at birth across different racial groups over time.
+@tbl-life_exp provides a more detailed breakdown of the life expectancy before and during COVID for...
+Unfortunately, due to the unavailability of data over time for Asians and American Indian and Alaska Native communities, they have been excluded from the time series graph.
+
+An intriguing observation is the consistently higher life expectancy among Hispanic individuals compared to other groups, even amidst the challenges posed by COVID-19.
+On the other hand, Black individuals l have consistently exhibited lower life expectancy, which further declined notably in 2021, reaching just over 70.8 years old.
+The life expectancy trends of white people and all other races and origins remain close together throughout the 14 year time period, with minimum variance.
+Specifically, the life expectancy for White individuals decreased from 78.8 to 76.4 years old from pre-pandemic levels in 2019 to 2021, while for Hispanic individuals, it dropped by 4.2 years, and for Black individuals, it declined by 4 years during the same period.
 
@@
 #| eval: true
+#| label: tbl-life_exp
+#| tbl-cap: Life Expectancy by Race (2019-2021)
+#| warning: false
+
+# Filter the data for the years 2019 to 2021
+life_exp_subset <- life_exp_melted[life_exp_melted$year_id >= 2019 & life_exp_melted$year_id <= 2021, ]
+
+# Pivot the data to create a table and rename columns
+life_exp_table <- reshape2::dcast(life_exp_subset, year_id ~ race, value.var = "life_exp")
+colnames(life_exp_table) <- c("Year", "All races and origins", "Hispanic", "AIAN", "Asian", "Black", "White")
+
+# Create a gt table
+life_exp_gt <- gt(life_exp_table, rowname_col = "Race") %>%
+  tab_header(
+    title = "Life Expectancy by Race (2019-2021)"
+  )
+
+# Print the gt table
+life_exp_gt
+
+```
+@fig-CLE serves as a valuable addition to our previous analysis with the inclusion of Asians and American Indians and Alaska Natives during the critical period from 2019 to 2021.
+With these additional ethnic categories, we can discern that American Indians and Alaska Natives experienced significant impacts from the pandemic, with a decrease of 4.7 years in the first year and 1.9 years in the second year, totaling 6.6 years.
+
+With a chart that depicts the change of life expectancy each year by ethnicity, we are able to better gain a more comprehensive understanding of the impact of COVID by ethnicity.
+While the declines in life expectancy are smaller in magnitude from 2020 to 2021, they are notably minimal for Asians and Hispanics, with reductions of -0.1 and -0.2 years respectively.
+(Insert some sources tmrw)
+
+```{r}
+#| echo: false
+#| eval: true
 #| warning: false
@@
 ```
+Numerous sources have indicated a correlation between a state's political affiliation and its handling of COVID issues (sources to be inserted tomorrow).
+To testify the claim, we collected the voting data for each of the 50 states for the 2020 presidential election.
+We then selected the ten states with the highest proportions of Republican votes.These voting patterns are visualized in Figure @fig-Vote, where red symbols represent Republican votes and blue symbols represent Democratic votes.
+Among the top ten states, Tennessee recorded the highest total number of votes, while Wyoming boasted the highest proportion of Republican votes.
 
@@
   mutate(gop_greater = if_else(total_gop_votes > total_dem_votes, 1, 0))
+write.csv(vote_data_by_state, "vote_data_by_state.csv", row.names = FALSE)
 
@@
 ```
+Given the variation in population sizes among states, direct comparisons of COVID case numbers are inherently flawed.
+As an alternative, we employed death rates per 100,000 people as a metric for evaluating each state's COVID preparedness and situation.
+We then produced @tbl-deaths that ranks the top ten states with the highest death rates from COVID per 100,000 people to examine the potential correlation between party preferences and COVID related deaths.
+Notably, We found that six of the ten top Republican states made a reappearance in the top death rates table; these states are Oklahoma, West Virginia, Arkansas, Alabama, Tennessee, and Kentucky.
+(Insert quotes about republican party and covid tmrw)
 
@@
 deaths$Rank <- 1:nrow(deaths)
+write.csv(deaths, "deaths.csv", row.names = FALSE)
 
@@
 
-# Print the table using kableExtra
-kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c") %>%
-  kable_styling(full_width = FALSE)
+```
+
+Following from our previous analysis regarding individuals' political affiliations, we have developed Figure @fig-deaths_rates_votes, which encompasses all 50 states of the US along with their political leanings based on which party garnered the majority votes.
+This information is juxtaposed against their respective COVID death rates.
+
+While we cannot make any definitive assertions about stark differences, we do observe that the Republican-leaning states are slightly more clustered around higher death rates ranging from  350 to 450 deaths, while the Democratic-leaning states appear to be more evenly distributed, and notably one Democratic-leaning state has the lowest death rate.An intriguing observation is that although many Republican-leaning states demonstrate higher COVID death rates, it is noteworthy that Arizona, typically considered a Democratic-leaning state, records the highest death rate among all states.
+
+(may add more)
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-deaths_rates_votes
+#| fig-cap: COVID-19 Death Rates vs. Republican Votes
+#| warning: false
+merged_data <- merge(deaths, vote_data_by_state, by.x = "State", by.y = "state_name")
+# Create a scatter plot of COVID-19 death rates against the percentage of Republican votes
+ggplot(merged_data, aes(x = as.factor(gop_greater), y = Death_Rate)) +
+  geom_point(shape = 21, fill = "skyblue", color = "darkblue", size = 3) + # Use filled circles with blue color
+  labs(x = "Republican Votes", y = "COVID-19 Death Rate") +
+  ggtitle("COVID-19 Death Rates vs. Republican Votes") +
+  theme_minimal() + # Use minimal theme for cleaner appearance
+  theme(plot.title = element_text(size = 16, face = "bold"), # Adjust title appearance
+        axis.title = element_text(size = 14)) + # Adjust axis labels appearance
+  scale_x_discrete(labels = c("Democratic Majority", "Republican Majority")) # Set custom labels for x-axis
 

### 2024-02-12 outputs/paper/paper.qmd
updated most of discussion

--- 2024-02-12 outputs/paper/paper.qmd
@@
 
+*The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper.
+A reader should be able to read only the introduction and know what was done, why, and what was found.
+Likely 3 or 4 paragraphs, or 10 per cent of total.
+The estimand is clearly stated in the introduction.*
+
 # Data {#sec-data}
@@
 
-## Data cleaning
+*A thorough discussion of measurement, relating to the dataset, is provided in the data section.*
 
@@
 
+# Conclusion 
 \newpage

### 2024-02-12 outputs/paper/paper.qmd
fixed paper

@@
 
-## Weaknesses and next stepsa
+To address the fragmentation in data collection and sharing witnessed during the pandemic, it's crucial to establish integrated data platforms that enable seamless health data exchange among various health agencies and stakeholders, utilizing cloud computing and APIs for real-time accessibility and usability.
+Equally important is enhancing data standardization and interoperability through universal standards like FHIR (Fast Healthcare Interoperability Resources) to facilitate efficient data sharing.
+Investing in digital surveillance systems, which employ AI and machine learning to sift through diverse data sources for early outbreak detection, is essential for rapid response to health threats.
+Furthermore, fostering public-private partnerships can harness the agility of the private sector and the public health expertise of governmental agencies to enhance data analytics capabilities.
+Ensuring the privacy and security of health data through robust governance frameworks and advanced encryption is paramount to maintaining public trust.
+Engaging communities in these initiatives ensures their relevance and fosters trust, while building global data sharing networks encourages international collaboration, crucial for a concerted response to pandemics.
+Collectively, these strategies are fundamental to bolstering public health decision-making and preparedness, making our health systems more resilient against the challenges posed by emerging infectious diseases.
+
+## Weaknesses and next steps
 

### 2024-02-12 outputs/paper/paper.qmd

--- 2024-02-12 outputs/paper/paper.qmd
@@
 library(kableExtra)
-library(here)
 library(gt)
@@
 #### Reading files ####
-# Set the working directory to the correct location
-ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
-world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
-life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
-vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
-deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
 
-# ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-# world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-# life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-# vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
-# deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
 
@@
 
+*The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper.
+A reader should be able to read only the introduction and know what was done, why, and what was found.
+Likely 3 or 4 paragraphs, or 10 per cent of total.
+The estimand is clearly stated in the introduction.*
+
 # Data {#sec-data}
+
+Our results are summarized in @fig:fig-GHS.
 
@@
 
-`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], and `scales` [@citeScales].
+`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], `gt` [@citeGT], and `scales` [@citeScales].
 
@@
 
+*A thorough discussion of measurement, relating to the dataset, is provided in the data section.*
+
 ## Data cleaning
-rename columns
-## Data Visualization
 
@@
 
-est_life
-
 ```
@@
 # Create a gt table
-life_exp_gt <- gt(life_exp_table, rowname_col = "Race") %>%
-  tab_header(
-    title = "Life Expectancy by Race (2019-2021)"
-  )
-
-# Print the gt table
-life_exp_gt
+life_exp_gt <- gt(life_exp_table, rowname_col = "Race")
 
 ```
-@fig-CLE serves as a valuable addition to our previous analysis with the inclusion of Asians and American Indians and Alaska Natives during the critical period from 2019 to 2021.
-With these additional ethnic categories, we can discern that American Indians and Alaska Natives experienced significant impacts from the pandemic, with a decrease of 4.7 years in the first year and 1.9 years in the second year, totaling 6.6 years.
-
-With a chart that depicts the change of life expectancy each year by ethnicity, we are able to better gain a more comprehensive understanding of the impact of COVID by ethnicity.
-While the declines in life expectancy are smaller in magnitude from 2020 to 2021, they are notably minimal for Asians and Hispanics, with reductions of -0.1 and -0.2 years respectively.
-(Insert some sources tmrw)
 
@@
 ```
+
+@fig-CLE serves as a valuable addition to our previous analysis with the inclusion of Asians and American Indians and Alaska Natives during the critical period from 2019 to 2021.
+With these additional ethnic categories, we can discern that American Indians and Alaska Natives experienced significant impacts from the pandemic, with a decrease of 4.7 years in the first year and 1.9 years in the second year, totaling 6.6 years.
+
+With a chart that depicts the change of life expectancy each year by ethnicity, we are able to better gain a more comprehensive understanding of the impact of COVID by ethnicity.
+While the declines in life expectancy are smaller in magnitude from 2020 to 2021, they are notably minimal for Asians and Hispanics, with reductions of -0.1 and -0.2 years respectively.
+(Insert some sources tmrw)
+
 Numerous sources have indicated a correlation between a state's political affiliation and its handling of COVID issues (sources to be inserted tomorrow).
@@
 
+table
 ```

### 2024-02-12 outputs/paper/paper.tex
idskdk

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={How did local health infrastructure and socio-political factors within different states and counties in the United States affect the disparities in COVID-19 outcomes, and what lessons can be learned for more targeted public health preparedness and response strategies in future pandemics?},
  pdfauthor={Adrian Ly; Sakhil Goel; Hannah Yu},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{How did local health infrastructure and socio-political factors
within different states and counties in the United States affect the
disparities in COVID-19 outcomes, and what lessons can be learned for
more targeted public health preparedness and response strategies in
future pandemics?\thanks{Code and data are available at:
https://github.com/hannahyu07/US-Covid-Analysis.git}}
\author{Adrian Ly \and Sakhil Goel \and Hannah Yu}
\date{February 12, 2024}

\begin{document}
\maketitle
\begin{abstract}
First sentence. Second sentence. Third sentence. Fourth sentence.
\end{abstract}

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\section{Introduction}\label{introduction}

\emph{This reproduction was performed after a replication on the Social
Science Reproduction platform:
}\textbf{\emph{\href{https://www.socialsciencereproduction.org/reproductions/c35e8e98-762d-4c00-a1a3-544ed0b2008d/index}{link
here}}}

\emph{The introduction is self-contained and tells a reader everything
they need to know including: 1) broader context to motivate; 2) some
detail about what the paper is about; 3) a clear gap that needs to be
filled; 4) what was done; 5) what was found; 6) why it is important; 7)
the structure of the paper. A reader should be able to read only the
introduction and know what was done, why, and what was found. Likely 3
or 4 paragraphs, or 10 per cent of total. The estimand is clearly stated
in the introduction.}

\section{Data}\label{sec-data}

Our results are summarized in (\textbf{fig:fig-GHS?}).

\subsection{Source}\label{source}

The datasets utilized in this paper were mainly obtained from the
\texttt{original\ paper} (Nuzzo and Ledesma 2023). Additionally, to
address the original paper's lack of US Covid statistics and political
party support data, we incorporated information from Jack and Oster
(2023) and Elflein (2023).

Jack and Oster (2023) discusses the long-term impacts of COVID-related
school closures. From this source, we utilized the dataset on voting
shares during the 2020 election by county. Elflein (2023) provides
summaries of COVID-19 death rates in the United States as of March 2023,
organized by state. Analyzing results from both datasets allows us to
explore the relationship between political affiliation and COVID-19
outcomes.Our reproduction aims to fill these gaps and also includes
tables and graphs that were not presented in the original paper to
support our findings.

\subsection{Methodology}\label{methodology}

\texttt{R} (R Core Team 2022) was the language and environment used for
the bulk of this analysis, alongside \texttt{tidyverse} (Wickham et al.
2019), \texttt{sf} (Pebesma 2018), \texttt{readxl} (Wickham and Bryan
2023), \texttt{knitr} (Xie 2014), \texttt{janitor} (Firke 2023),
\texttt{lubridate} (Grolemund and Wickham 2011), \texttt{dplyr} (Wickham
et al. 2023), \texttt{data.table} (Barrett et al. 2024),
\texttt{RColorBrewer} (Neuwirth 2022), \texttt{ggpubr} (Kassambara
2023), \texttt{ggplot2} (Wickham 2016), \texttt{here} (Müller 2020),
\texttt{kableExtra} (Zhu 2024), \texttt{webshot} (Chang 2023a),
\texttt{webshot2} (Chang 2023b), \texttt{gt} (Iannone et al. 2024), and
\texttt{scales} (Wickham, Pedersen, and Seidel 2023).

\subsection{Data Measurement}\label{data-measurement}

\emph{A thorough discussion of measurement, relating to the dataset, is
provided in the data section.}

\subsection{Data cleaning}\label{data-cleaning}

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-GHS-1.pdf}

}

\caption{\label{fig-GHS}Global Health Security Index Scores by Country}

\end{figure}%

\section{Results}\label{results}

Our results are summarized in the following figures. \textbf{?@fig-ELE}
illustrates the trend of life expectancy at birth across different
racial groups over time. Table~\ref{tbl-life_exp} provides a more
detailed breakdown of the life expectancy before and during COVID
for\ldots{} Unfortunately, due to the unavailability of data over time
for Asians and American Indian and Alaska Native communities, they have
been excluded from the time series graph.

An intriguing observation is the consistently higher life expectancy
among Hispanic individuals compared to other groups, even amidst the
challenges posed by COVID-19. On the other hand, Black individuals l
have consistently exhibited lower life expectancy, which further
declined notably in 2021, reaching just over 70.8 years old. The life
expectancy trends of white people and all other races and origins remain
close together throughout the 14 year time period, with minimum
variance. Specifically, the life expectancy for White individuals
decreased from 78.8 to 76.4 years old from pre-pandemic levels in 2019
to 2021, while for Hispanic individuals, it dropped by 4.2 years, and
for Black individuals, it declined by 4 years during the same period.

\begin{table}

\caption{\label{tbl-life\_exp}Life Expectancy by Race (2019-2021)}

\centering{

}

\end{table}%

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-CLE-1.pdf}

}

\caption{\label{fig-CLE}Change in Life Expectancy at Birth from the
Previous Year}

\end{figure}%

Figure~\ref{fig-CLE} serves as a valuable addition to our previous
analysis with the inclusion of Asians and American Indians and Alaska
Natives during the critical period from 2019 to 2021. With these
additional ethnic categories, we can discern that American Indians and
Alaska Natives experienced significant impacts from the pandemic, with a
decrease of 4.7 years in the first year and 1.9 years in the second
year, totaling 6.6 years.

With a chart that depicts the change of life expectancy each year by
ethnicity, we are able to better gain a more comprehensive understanding
of the impact of COVID by ethnicity. While the declines in life
expectancy are smaller in magnitude from 2020 to 2021, they are notably
minimal for Asians and Hispanics, with reductions of -0.1 and -0.2 years
respectively. (Insert some sources tmrw)

Numerous sources have indicated a correlation between a state's
political affiliation and its handling of COVID issues (sources to be
inserted tomorrow). To testify the claim, we collected the voting data
for each of the 50 states for the 2020 presidential election. We then
selected the ten states with the highest proportions of Republican
votes.These voting patterns are visualized in Figure
Figure~\ref{fig-Vote}, where red symbols represent Republican votes and
blue symbols represent Democratic votes. Among the top ten states,
Tennessee recorded the highest total number of votes, while Wyoming
boasted the highest proportion of Republican votes.

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-Vote-1.pdf}

}

\caption{\label{fig-Vote}Total Number of Votes in Top 10 States with
Highest Proportion of Republican Votes}

\end{figure}%

Given the variation in population sizes among states, direct comparisons
of COVID case numbers are inherently flawed. As an alternative, we
employed death rates per 100,000 people as a metric for evaluating each
state's COVID preparedness and situation. We then produced
Table~\ref{tbl-deaths} that ranks the top ten states with the highest
death rates from COVID per 100,000 people to examine the potential
correlation between party preferences and COVID related deaths. Notably,
We found that six of the ten top Republican states made a reappearance
in the top death rates table; these states are Oklahoma, West Virginia,
Arkansas, Alabama, Tennessee, and Kentucky. (Insert quotes about
republican party and covid tmrw)

\begin{table}

\caption{\label{tbl-deaths}Top 10 States with Highest Death Rates from
COVID-19 (per 100,000 people)}

\centering{

}

\end{table}%

Following from our previous analysis regarding individuals' political
affiliations, we have developed Figure
Figure~\ref{fig-deaths_rates_votes}, which encompasses all 50 states of
the US along with their political leanings based on which party garnered
the majority votes. This information is juxtaposed against their
respective COVID death rates.

While we cannot make any definitive assertions about stark differences,
we do observe that the Republican-leaning states are slightly more
clustered around higher death rates ranging from 350 to 450 deaths,
while the Democratic-leaning states appear to be more evenly
distributed, and notably one Democratic-leaning state has the lowest
death rate.An intriguing observation is that although many
Republican-leaning states demonstrate higher COVID death rates, it is
noteworthy that Arizona, typically considered a Democratic-leaning
state, records the highest death rate among all states.

(may add more)

\begin{figure}

\centering{

\includegraphics{paper_files/figure-pdf/fig-deaths_rates_votes-1.pdf}

}

\caption{\label{fig-deaths\_rates\_votes}COVID-19 Death Rates
vs.~Republican Votes}

\end{figure}%

\newpage

\section{Discussion}\label{discussion}

This begs the question as to why we are seeing these results. There
isn't exactly a single answer to this question, however we can certainly
point out some considerable factors to this result.

\subsection{Influence of political polarization on adherence to health
guidelines.}\label{influence-of-political-polarization-on-adherence-to-health-guidelines.}

Political polarization has significantly impacted the adherence to
health guidelines during the COVID-19 pandemic. The divergence in
political ideologies has translated into differing attitudes towards
health directives, including mask mandates, social distancing, and
vaccination uptake.

Various studies and our own results have shown that areas with higher
support for one political party exhibited distinct behaviors and
compliance levels with health recommendations, which directly correlated
with COVID-19 case rates and mortality. An news article from
\texttt{ABC\ News} (Diab and Kumar 2023) shows that the top states with
the highest COVID-19 deaths are Arizona, and Washington with 581 deaths
and 526 deaths respectively per 100,000 people. According to 2020
presidential voting data published by \texttt{CNN}, we have both states
having the electoral vote of democrat with Washington wining by 58\%
(\emph{2020 Election Results by State, Washington} 2020) and Arizona
winning by 49.4\% (\emph{2020 Election Results by State, Arizona} 2020).
Another news article by \texttt{ContagionLive} (Parkinson 2023) also
makes the claim of both Arizona and Washington having the highest
COVID-19 mortality. This polarization has not only influenced individual
behavior but also shaped state and local health policies, further
entrenching the disparities in health outcomes.

The adherence to health guidelines are evident in the varied health
outcomes observed across the United States. Regions with lower
compliance to health directives, often influenced by political leanings,
have experienced higher rates of COVID-19 transmission,
hospitalizations, and deaths. The disparities in vaccine uptake, driven
by political affiliations, have further exacerbated these outcomes,
leaving certain communities more vulnerable to the virus and its
variants. In order to mitigate the influence of political polarization
on public health, it is imperative to depoliticize health guidelines and
focus on evidence-based approaches to disease prevention and control.
Building trust in health institutions and promoting bipartisan support
for public health measures are essential steps towards achieving higher
compliance and better health outcomes. Engaging trusted community
leaders and utilizing targeted communication strategies can also help
bridge the divide and encourage adherence to health guidelines.

\subsection{Impact of government transparency and consistent
communication on public
trust.}\label{impact-of-government-transparency-and-consistent-communication-on-public-trust.}

The politicization of health guidelines and mixed messages from
political and health leaders during the COVID-19 pandemic have
significantly undermined the effectiveness of public health messaging,
leading to confusion, skepticism, and eroded trust among the public.
Initially, inconsistencies in recommendations, such as on mask usage,
challenged the principle of clear, consistent, and science-based
communication essential for an effective public health response.
Moreover, the transparency of government actions and decision-making
processes is crucial in building and maintaining public trust,
especially during health crises. The level of public trust was greatly
affected by the openness and accuracy with which governments, at all
levels, communicated about the evolving situation, the reasoning behind
guidelines, and the measures taken to combat the virus, emphasizing the
importance of transparent reporting of data related to case counts,
hospitalizations, vaccine distribution, and side effects. Furthermore,
consistent communication from public health officials and government
leaders is key to ensuring adherence to health guidelines, where
inconsistencies, such as changes in mask-wearing guidelines without
clear explanations, have led to public confusion. The direct correlation
between government transparency, consistent communication, and public
behavior is self-evident, with populations receiving clear and
transparent information being more likely to adhere to guidelines,
participate in testing and tracing efforts, and accept vaccination.
Drawing lessons from the pandemic, strategies for improving government
transparency and communication in future health emergencies should
include establishing centralized information hubs, ensuring regular and
predictable communication from health authorities, engaging community
leaders in information dissemination, and harnessing digital platforms
and social media to amplify public health messages, thus reinforcing
public trust and compliance.

\subsection{Role of social vulnerabilities and healthcare access
disparities in pandemic
impact.}\label{role-of-social-vulnerabilities-and-healthcare-access-disparities-in-pandemic-impact.}

\subsection{Strategies for improving real-time data collection and
sharing for public health
decisions.}\label{strategies-for-improving-real-time-data-collection-and-sharing-for-public-health-decisions.}

To address the fragmentation in data collection and sharing witnessed
during the pandemic, it's crucial to establish integrated data platforms
that enable seamless health data exchange among various health agencies
and stakeholders, utilizing cloud computing and APIs for real-time
accessibility and usability. Equally important is enhancing data
standardization and interoperability through universal standards like
FHIR (Fast Healthcare Interoperability Resources) to facilitate
efficient data sharing. Investing in digital surveillance systems, which
employ AI and machine learning to sift through diverse data sources for
early outbreak detection, is essential for rapid response to health
threats. Furthermore, fostering public-private partnerships can harness
the agility of the private sector and the public health expertise of
governmental agencies to enhance data analytics capabilities. Ensuring
the privacy and security of health data through robust governance
frameworks and advanced encryption is paramount to maintaining public
trust. Engaging communities in these initiatives ensures their relevance
and fosters trust, while building global data sharing networks
encourages international collaboration, crucial for a concerted response
to pandemics. Collectively, these strategies are fundamental to
bolstering public health decision-making and preparedness, making our
health systems more resilient against the challenges posed by emerging
infectious diseases.

\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}

Weaknesses and next steps should also be included.

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-citesource3}
\emph{2020 Election Results by State, Arizona}. 2020. CNN.
\url{https://www.cnn.com/election/2020/results/state/arizona}.

\bibitem[\citeproctext]{ref-citesource2}
\emph{2020 Election Results by State, Washington}. 2020. CNN.
\url{https://www.cnn.com/election/2020/results/state/washington}.

\bibitem[\citeproctext]{ref-citeDataTable}
Barrett, Tyson, Matt Dowle, Arun Srinivasan, Jan Gorecki, Michael
Chirico, and Toby Hocking. 2024. \emph{Data.table: Extension of
`Data.frame`}. \url{https://CRAN.R-project.org/package=data.table}.

\bibitem[\citeproctext]{ref-citewebshot}
Chang, Winston. 2023a. \emph{Webshot: Take Screenshots of Web Pages}.
\url{https://CRAN.R-project.org/package=webshot}.

\bibitem[\citeproctext]{ref-citewebshot2}
---------. 2023b. \emph{Webshot2: Take Screenshots of Web Pages}.
\url{https://CRAN.R-project.org/package=webshot2}.

\bibitem[\citeproctext]{ref-citesource1}
Diab, Dr. Alaa, and Dr. Keerthana Kumar. 2023. \emph{COVID-19 Death
Rates Varied Dramatically Across US, Major Analysis Finds}. ABC News.
\url{https://abcnews.go.com/Health/covid-19-death-rates-varied-dramatically-us-major/story?id=98055024}.

\bibitem[\citeproctext]{ref-citedata2}
Elflein, John. 2023. {``U.s. COVID-19 Death Rate by State.''}
\emph{Statista}.
\url{https://www.statista.com/statistics/1109011/coronavirus-covid19-death-rates-us-by-state/}.

\bibitem[\citeproctext]{ref-citeJanitor}
Firke, Sam. 2023. \emph{Janitor: Simple Tools for Examining and Cleaning
Dirty Data}. \url{https://CRAN.R-project.org/package=janitor}.

\bibitem[\citeproctext]{ref-citeLubridate}
Grolemund, Garrett, and Hadley Wickham. 2011. {``Dates and Times Made
Easy with {lubridate}.''} \emph{Journal of Statistical Software} 40 (3):
1--25. \url{https://www.jstatsoft.org/v40/i03/}.

\bibitem[\citeproctext]{ref-citeGT}
Iannone, Richard, Joe Cheng, Barret Schloerke, Ellis Hughes, Alexandra
Lauer, and JooYoung Seo. 2024. \emph{Gt: Easily Create
Presentation-Ready Display Tables}.
\url{https://CRAN.R-project.org/package=gt}.

\bibitem[\citeproctext]{ref-citedata1}
Jack, Rebecca, and Emily Oster. 2023. {``COVID-19, School Closures, and
Outcomes.''} \emph{Journal of Economic Perspectives} 37 (4): 51--70.
https://doi.org/\url{https://doi.org/10.1257/jep.37.4.51}.

\bibitem[\citeproctext]{ref-citeggpubr}
Kassambara, Alboukadel. 2023. \emph{Ggpubr: 'Ggplot2' Based Publication
Ready Plots}. \url{https://CRAN.R-project.org/package=ggpubr}.

\bibitem[\citeproctext]{ref-citeHere}
Müller, Kirill. 2020. \emph{Here: A Simpler Way to Find Your Files}.
\url{https://CRAN.R-project.org/package=here}.

\bibitem[\citeproctext]{ref-citeRColorBrewer}
Neuwirth, Erich. 2022. \emph{RColorBrewer: ColorBrewer Palettes}.
\url{https://CRAN.R-project.org/package=RColorBrewer}.

\bibitem[\citeproctext]{ref-citearticle}
Nuzzo, Jennifer B., and Jorge R. Ledesma. 2023. \emph{Why Did the Best
Prepared Country in the World Fare so Poorly During COVID?}
\url{https://www.aeaweb.org/articles?id=10.1257/jep.37.4.3}.

\bibitem[\citeproctext]{ref-citesource4}
Parkinson, John. 2023. \emph{Which States Saw Greatest COVID-19
Mortality?} ContagionLive.
\url{https://www.contagionlive.com/view/which-states-saw-greatest-covid-19-mortality-}.

\bibitem[\citeproctext]{ref-citeSF}
Pebesma, Edzer. 2018. {``{Simple Features for R: Standardized Support
for Spatial Vector Data}.''} \emph{{The R Journal}} 10 (1): 439--46.
\url{https://doi.org/10.32614/RJ-2018-009}.

\bibitem[\citeproctext]{ref-citeR}
R Core Team. 2022. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-citeggplot}
Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
Analysis}. Springer-Verlag New York.
\url{https://ggplot2.tidyverse.org}.

\bibitem[\citeproctext]{ref-citeTidyVerse}
Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
D'Agostino McGowan, Romain François, Garrett Grolemund, et al. 2019.
{``Welcome to the {tidyverse}.''} \emph{Journal of Open Source Software}
4 (43): 1686. \url{https://doi.org/10.21105/joss.01686}.

\bibitem[\citeproctext]{ref-citeReadXL}
Wickham, Hadley, and Jennifer Bryan. 2023. \emph{Readxl: Read Excel
Files}. \url{https://CRAN.R-project.org/package=readxl}.

\bibitem[\citeproctext]{ref-citeDplyr}
Wickham, Hadley, Romain François, Lionel Henry, Kirill Müller, and Davis
Vaughan. 2023. \emph{Dplyr: A Grammar of Data Manipulation}.
\url{https://CRAN.R-project.org/package=dplyr}.

\bibitem[\citeproctext]{ref-citeScales}
Wickham, Hadley, Thomas Lin Pedersen, and Dana Seidel. 2023.
\emph{Scales: Scale Functions for Visualization}.
\url{https://CRAN.R-project.org/package=scales}.

\bibitem[\citeproctext]{ref-citeKnitR}
Xie, Yihui. 2014. {``Knitr: A Comprehensive Tool for Reproducible
Research in {R}.''} In \emph{Implementing Reproducible Computational
Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
Peng. Chapman; Hall/CRC.

\bibitem[\citeproctext]{ref-citeKableExtra}
Zhu, Hao. 2024. \emph{kableExtra: Construct Complex Table with 'Kable'
and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{CSLReferences}

\end{document}

### 2024-02-12 scripts/01-simulate_data.R

@@ -1,108 +1,0 @@
-#### Preamble ####
-# Purpose: Simulate Data of Interest
-# Author: Adrian Ly, Hannah Yu
-# Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca
-# License: MIT
-
-#### Workspace setup ####
-library(tidyverse)
-library(sf)
-library(readxl)
-library(knitr)
-library(janitor)
-library(lubridate)
-library(dplyr)
-library(data.table)
-library(RColorBrewer)
-library(ggpubr)
-library(scales)
-library(here)
-library(kableExtra)
-
-#### Simulate data ####
-
-# Simulate the average life expectancy of people by ethnicity
-# Set seed for reproducibility
-set.seed(123)
-
-# Define the years
-years <- 2006:2021
-
-# Define life expectancy ranges for each group
-life_expectancy_ranges <- list(
-  "All" = c(77, 80),
-  "Hispanic" = c(80, 83),
-  "Non-Hispanic AIAN" = c(65, 72),
-  "Non-Hispanic Asian" = c(83, 87),
-  "Non-Hispanic Black" = c(70, 76),
-  "Non-Hispanic White" = c(77, 81)
-)
-
-# Generate the dataset
-data <- matrix(NA, nrow = length(years), ncol = length(life_expectancy_ranges) + 1)
-colnames(data) <- c("year_id", names(life_expectancy_ranges))
-data[, 1] <- years
-
-for (i in 1:length(years)) {
-  for (group in names(life_expectancy_ranges)) {
-    if (group == "All") {
-      life_expectancy <- runif(1, life_expectancy_ranges[[group]][1], life_expectancy_ranges[[group]][2])
-    } else {
-      life_expectancy <- runif(1, life_expectancy_ranges[[group]][1], life_expectancy_ranges[[group]][2])
-    }
-    data[i, which(colnames(data) == group)] <- round(life_expectancy, 1)
-  }
-}
-
-# Save the dataset to a CSV file
-write.csv(data, "outputs/data/simulated_life_exp.csv", row.names = FALSE)
-
-# Print the dataset to the console
-print(data)
-
-
-# Simulate the Vote at Each State
-set.seed(123)
-# Define the state names
-state_names <- c(
-  "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", "Georgia",
[…]

### 2024-02-12 scripts/02-test_data.R

@@ -1,80 +1,0 @@
-#### Preamble ####
-# Purpose: Tests the Simulated Data
-# Author: Adrian Ly
-# Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca
-# License: MIT
-
-#### Workspace setup ####
-library(tidyverse)
-library(sf)
-library(readxl)
-library(knitr)
-library(janitor)
-library(lubridate)
-library(dplyr)
-library(data.table)
-library(RColorBrewer)
-library(ggpubr)
-library(scales)
-library(here)
-library(kableExtra)
-
-#### Test data ####
-## Test simulated_life_data
-# Load the dataset generated in R
-data1 <- read.csv("outputs/data/simulated_life_exp.csv", stringsAsFactors = FALSE)
-
-# 1. Data Structure Check
-# Check the number of rows and columns
-structure_check <- dim(data)[1] == length(years) & dim(data)[2] == length(life_expectancy_ranges) + 1
-print(paste("Data Structure Check:", ifelse(structure_check, "Passed", "Failed")))
-
-# 2. Range Check
-# Define the expected ranges
-expected_ranges <- c(77:80, 80:83, 65:72, 83:87, 70:76, 77:81)
-# Check if all values fall within the expected ranges
-range_check <- all(sapply(data[, -1], function(x) all(x >= min(expected_ranges) & x <= max(expected_ranges))))
-print(paste("Range Check:", ifelse(range_check, "Passed", "Failed")))
-
-# 3. Numeric Check
-# Check if all values are numeric
-numeric_check <- all(sapply(data, is.numeric))
-print(paste("Numeric Check:", ifelse(numeric_check, "Passed", "Failed")))
-
-## Test simulated_vote_data
-# Load the dataset generated in R
-data2 <- read.csv("outputs/data/simulated_vote_data.csv", stringsAsFactors = FALSE)
-
-# Test 1: Check number of rows and columns
-num_rows <- nrow(data2)
-num_cols <- ncol(data2)
-
-expected_num_rows <- 50 # Expecting 50 states
-expected_num_cols <- 8  # Expecting 7 columns (plus 1 for state_name)
-
-if (num_rows == expected_num_rows && num_cols == expected_num_cols) {
-  print("Test 1 passed: Correct number of rows and columns.")
-} else {
-  print("Test 1 failed: Incorrect number of rows or columns.")
-}
-
-# Test 2: Check for missing values
[…]

### 2024-02-12 scripts/Replication.R

@@ -1,189 +1,0 @@
-#### Preamble ####
-# Purpose: Paper Replication
-# Author: Adrian Ly, Hannah Yu
-# Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca
-# License: MIT
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(sf)
-library(readxl)
-library(knitr)
-library(janitor)
-library(lubridate)
-library(dplyr)
-library(data.table)
-library(RColorBrewer)
-library(ggpubr)
-library(scales)
-library(here)
-library(kableExtra)
-library(webshot)
-library(webshot2)
-
-#### Reading files ####
-ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
-deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
-
-## CREATE FIGURE 1
-ghs_data <- ghs_data %>%
-  mutate(COUNTRY = case_when(
-    Country == "United States of America" ~ "United States",
-    Country == "Bolivia (Plurinational State of)" ~ "Bolivia",
-    # Add other necessary name changes here
-    TRUE ~ Country
-  ))
-
-world_ghs <- merge(world_map, ghs_data, by.x = "COUNTRY", by.y = "COUNTRY", all.x = TRUE)
-
-
-ggplot(data = world_ghs) +
-  geom_sf(aes(fill = `OVERALL SCORE`), color = "white") +
-  scale_fill_distiller(palette = "Spectral", na.value = "lightgrey", 
-                       name = "GHS Index Score", direction = 1,
-                       limits = c(min(ghs_data$`OVERALL SCORE`, na.rm = TRUE), max(ghs_data$`OVERALL SCORE`, na.rm = TRUE)),
-                       labels = comma) +  # Use comma to format numbers with commas
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        plot.title = element_text(hjust = 0.5))
-
-# Save the plot to a file
-ggsave("outputs/graphs/GHS_Index_Map.png", width = 11, height = 8)
-
-## CREATE FIGURE 2
-
[…]

### 2024-02-12 outputs/paper/paper.qmd
Descriptive message about the changes

--- 2024-02-12 outputs/paper/paper.qmd
@@
 
-*The introduction is self-contained and tells a reader everything they need to know including: 1) broader context to motivate; 2) some detail about what the paper is about; 3) a clear gap that needs to be filled; 4) what was done; 5) what was found; 6) why it is important; 7) the structure of the paper.
-A reader should be able to read only the introduction and know what was done, why, and what was found.
-Likely 3 or 4 paragraphs, or 10 per cent of total.
-The estimand is clearly stated in the introduction.*
-
 # Data {#sec-data}
@@
 The datasets utilized in this paper were mainly obtained from the `original paper` [@citearticle].
-Additionally, to address the original paper's lack of US Covid statistics and political party support data, we incorporated information from @citedata1 and @citedata2.
+Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and @citedata2.
 
 @citedata1 discusses the long-term impacts of COVID-related school closures. From this source, we utilized the dataset on voting shares during the 2020 election by county.
-@citedata2 provides summaries of COVID-19 death rates in the United States as of March 2023, organized by state.
-Analyzing results from both datasets allows us to explore the relationship between political affiliation and COVID-19 outcomes.Our reproduction aims to fill these gaps and also includes tables and graphs that were not presented in the original paper to support our findings.
+@citedata2 summarizes COVID-19 death rates in the United States as of March 2023, organized by state.
+Analyzing results from both datasets allows us to explore the relationship between political affiliation and COVID-19 outcomes.
+Our reproduction aims to fill these gaps and includes tables and graphs not presented in the original paper to support our findings.
 
@@
 
-*A thorough discussion of measurement, relating to the dataset, is provided in the data section.*
+## Data cleaning
 
@@
 
-Our results are summarized in @fig-GHS.
+Our results are summarized in the following figures.
+@fig-ELE illustrates the trend of life expectancy at birth across different racial groups over time.
+@tbl-life-exp provides a more detailed breakdown of the life expectancy before and during COVID for...
+
+Unfortunately, due to the unavailability of data over time for Asian, American Indian, and Alaska Native communities, they have been excluded from the time series graph.
+
+An intriguing observation is the consistently higher life expectancy among Hispanic individuals compared to other groups, even amidst the challenges posed by COVID-19.
+On the other hand, Black individuals l have consistently exhibited lower life expectancy, which further declined notably in 2021, reaching just over 70.8 years old.
+The life expectancy trends of white people and all other races and origins remain close together throughout the 14 years, with minimum variance.
+Specifically, the life expectancy for White individuals decreased from 78.8 to 76.4 years old from pre-pandemic levels in 2019 to 2021, while for Hispanic individuals, it dropped by 4.2 years, and for Black individuals, it declined by 4 years during the same period.
 
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-life-exp
+#| tbl-cap: Life Expectancy by Race (2019-2021)
+#| warning: false
+# Filter the data for the years 2019 to 2021
+life_exp_subset <- life_exp_melted[life_exp_melted$year_id >= 2019 & life_exp_melted$year_id <= 2021, ]
+
+# Pivot the data to create a table and rename columns
+life_exp_table <- reshape2::dcast(life_exp_subset, year_id ~ race, value.var = "life_exp")
+colnames(life_exp_table) <- c("Year", "All races and origins", "Hispanic", "AIAN", "Asian", "Black", "White")
+
+# Create a table using knitr::kable
+
+kable(life_exp_table, col.names = c("Year", "All Races and Origins", "Hispanic", "AIAN", "Asian", "Black", "White"))
+
+```
+
+@fig-CLE serves as a valuable addition to our previous analysis with the inclusion of Asians, American Indians, and Alaska Natives during the critical period from 2019 to 2021.
+With these additional ethnic categories, we can discern that American Indians and Alaska Natives experienced significant impacts from the pandemic, with a decrease of 4.7 years in the first year and 1.9 years in the second year, totaling 6.6 years.
+
+With a chart that depicts the change in life expectancy each year by ethnicity, we can better gain a more comprehensive understanding of the impact of COVID by ethnicity.
+While the declines in life expectancy are smaller in magnitude from 2020 to 2021, they are notably minimal for Asians and Hispanics, with reductions of -0.1 and -0.2 years respectively.
+(Insert some sources tmrw)
+
+Numerous sources have indicated a correlation between a state's political affiliation and its handling of COVID issues (sources to be inserted tomorrow).
+To testify to the claim, we collected the voting data for each of the 50 states for the 2020 presidential election.
+We then selected the ten states with the highest proportions of Republican votes.
+These voting patterns are visualized in Figure @fig-Vote, where red symbols represent Republican votes and blue symbols represent Democratic votes.
+Among the top ten states, Tennessee recorded the highest total number of votes, while Wyoming boasted the highest proportion of Republican votes.
+
+Given the variation in population sizes among states, direct comparisons of COVID-19 case numbers are inherently flawed.
+As an alternative, we employed death rates per 100,000 people as a metric for evaluating each state's COVID preparedness and situation.
+We then produced @tbl-deaths that rank the top ten states with the highest death rates from COVID-19 per 100,000 people to examine the potential correlation between party preferences and COVID-related deaths.
+Notably, We found that six of the ten top Republican states made a reappearance in the top death rates table; these states are Oklahoma, West Virginia, Arkansas, Alabama, Tennessee, and Kentucky.
+(Insert quotes about the republican party and COVID tmrw)
+
+Following from our previous analysis regarding individuals' political affiliations, we have developed Figure @fig-deaths-rates-votes, which encompasses all 50 states of the US along with their political leanings based on which party garnered the majority votes.
+This information is juxtaposed against their respective COVID-19 death rates.
+
+While we cannot make any definitive assertions about stark differences, we do observe that the Republican-leaning states are slightly more clustered around higher death rates ranging from  350 to 450 deaths, while the Democratic-leaning states appear to be more evenly distributed, and notably one Democratic-leaning state has the lowest death rate.
+An intriguing observation is that although many Republican-leaning states demonstrate higher COVID death rates, it is noteworthy that Arizona, typically considered a Democratic-leaning state, records the highest death rate among all states.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-deaths-rates-votes
+#| fig-cap: COVID-19 Death Rates vs. Republican Votes
+#| warning: false
+merged_data <- merge(deaths, vote_data_by_state, by.x = "State", by.y = "state_name")
+# Create a scatter plot of COVID-19 death rates against the percentage of Republican votes
+ggplot(merged_data, aes(x = as.factor(gop_greater), y = Death_Rate)) +
+  geom_point(shape = 21, fill = "skyblue", color = "darkblue", size = 3) + # Use filled circles with blue color
+  labs(x = "Republican Votes", y = "COVID-19 Death Rate") +
+  ggtitle("COVID-19 Death Rates vs. Republican Votes") +
+  theme_minimal() + # Use minimal theme for cleaner appearance
+  theme(plot.title = element_text(size = 16, face = "bold"), # Adjust title appearance
+        axis.title = element_text(size = 14)) + # Adjust axis labels appearance
+  scale_x_discrete(labels = c("Democratic Majority", "Republican Majority")) # Set custom labels for x-axis
+```
+
 \newpage
@@
 
-This begs the question as to why we are seeing these results. There isn't exactly a single answer to this question, however we can certainly point out some considerable factors to this result.
+This begs the question as to why we are seeing these results. There isn't exactly a single answer to this question, however, we can certainly point out some considerable factors to this result.
 
@@
 
-Various studies and our own results have shown that areas with higher support for one political party exhibited distinct behaviors and compliance levels with health recommendations, which directly correlated with COVID-19 case rates and mortality.
-An news article from `ABC News` [@citesource1] shows that the top states with the highest COVID-19 deaths are Arizona, and Washington with 581 deaths and 526 deaths respectively per 100,000 people.
-According to 2020 presidential voting data published by `CNN`, we have both states having the electoral vote of democrat with Washington wining by 58% [@citesource2] and Arizona winning by 49.4% [@citesource3].
-Another news article by `ContagionLive` [@citesource4] also makes the claim of both Arizona and Washington having the highest COVID-19 mortality.
+Various studies and our results have shown that areas with higher support for one political party exhibited distinct behaviors and compliance levels with health recommendations, which directly correlated with COVID-19 case rates and mortality.
+A news article from `ABC News` [@citesource1] shows that the top states with the highest COVID-19 deaths are Arizona, and Washington with 581 deaths and 526 deaths respectively per 100,000 people.
+According to 2020 presidential voting data published by `CNN`, we have both states having the electoral vote of Democrat with Washington winning by 58% [@citesource2] and Arizona winning by 49.4% [@citesource3].
+Another news article by `ContagionLive` [@citesource4] also claims both Arizona and Washington have the highest COVID-19 mortality.
 This polarization has not only influenced individual behavior but also shaped state and local health policies, further entrenching the disparities in health outcomes.
 
-The adherence to health guidelines are evident in the varied health outcomes observed across the United States.
+The adherence to health guidelines is evident in the varied health outcomes observed across the United States.
 Regions with lower compliance to health directives, often influenced by political leanings, have experienced higher rates of COVID-19 transmission, hospitalizations, and deaths.
 The disparities in vaccine uptake, driven by political affiliations, have further exacerbated these outcomes, leaving certain communities more vulnerable to the virus and its variants.
-In order to mitigate the influence of political polarization on public health, it is imperative to depoliticize health guidelines and focus on evidence-based approaches to disease prevention and control.
-Building trust in health institutions and promoting bipartisan support for public health measures are essential steps towards achieving higher compliance and better health outcomes.
+To mitigate the influence of political polarization on public health, it is imperative to depoliticize health guidelines and focus on evidence-based approaches to disease prevention and control.
+Building trust in health institutions and promoting bipartisan support for public health measures are essential steps toward achieving higher compliance and better health outcomes.
 Engaging trusted community leaders and utilizing targeted communication strategies can also help bridge the divide and encourage adherence to health guidelines.
@@
 
-## Role of social vulnerabilities and healthcare access disparities in pandemic impact.
+## Role of social vulnerabilities and healthcare access disparities in pandemic impact
 
@@
 To address the fragmentation in data collection and sharing witnessed during the pandemic, it's crucial to establish integrated data platforms that enable seamless health data exchange among various health agencies and stakeholders, utilizing cloud computing and APIs for real-time accessibility and usability.
-Equally important is enhancing data standardization and interoperability through universal standards like FHIR (Fast Healthcare Interoperability Resources) to facilitate efficient data sharing.
+Equally important is enhancing data standardization and interoperability through universal standards like FHIR (Fast Healthcare Interoperability Resources) to facilitate efficient data sharing [@citesource8].
 Investing in digital surveillance systems, which employ AI and machine learning to sift through diverse data sources for early outbreak detection, is essential for rapid response to health threats.
@@
 Ensuring the privacy and security of health data through robust governance frameworks and advanced encryption is paramount to maintaining public trust.
-Engaging communities in these initiatives ensures their relevance and fosters trust, while building global data sharing networks encourages international collaboration, crucial for a concerted response to pandemics.
+Engaging communities in these initiatives ensures their relevance and fosters trust while building global data-sharing networks encourages international collaboration, crucial for a concerted response to pandemics.
 Collectively, these strategies are fundamental to bolstering public health decision-making and preparedness, making our health systems more resilient against the challenges posed by emerging infectious diseases.

### 2024-02-12 scripts/01-simulate_data.R
fixed???

@@ -1,0 +1,108 @@
+#### Preamble ####
+# Purpose: Simulate Data of Interest
+# Author: Adrian Ly, Hannah Yu
+# Date: 9 February 2024 
+# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
+library(scales)
+library(here)
+library(kableExtra)
+
+#### Simulate data ####
+
+# Simulate the average life expectancy of people by ethnicity
+# Set seed for reproducibility
+set.seed(123)
+
+# Define the years
+years <- 2006:2021
+
+# Define life expectancy ranges for each group
+life_expectancy_ranges <- list(
+  "All" = c(77, 80),
+  "Hispanic" = c(80, 83),
+  "Non-Hispanic AIAN" = c(65, 72),
+  "Non-Hispanic Asian" = c(83, 87),
+  "Non-Hispanic Black" = c(70, 76),
+  "Non-Hispanic White" = c(77, 81)
+)
+
+# Generate the dataset
+data <- matrix(NA, nrow = length(years), ncol = length(life_expectancy_ranges) + 1)
+colnames(data) <- c("year_id", names(life_expectancy_ranges))
+data[, 1] <- years
+
+for (i in 1:length(years)) {
+  for (group in names(life_expectancy_ranges)) {
+    if (group == "All") {
+      life_expectancy <- runif(1, life_expectancy_ranges[[group]][1], life_expectancy_ranges[[group]][2])
+    } else {
+      life_expectancy <- runif(1, life_expectancy_ranges[[group]][1], life_expectancy_ranges[[group]][2])
+    }
+    data[i, which(colnames(data) == group)] <- round(life_expectancy, 1)
+  }
+}
+
+# Save the dataset to a CSV file
+write.csv(data, "outputs/data/simulated_life_exp.csv", row.names = FALSE)
+
+# Print the dataset to the console
+print(data)
+
+
+# Simulate the Vote at Each State
+set.seed(123)
+# Define the state names
+state_names <- c(
+  "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", "Georgia",
[…]

### 2024-02-12 scripts/02-test_data.R

@@ -1,0 +1,80 @@
+#### Preamble ####
+# Purpose: Tests the Simulated Data
+# Author: Adrian Ly
+# Date: 9 February 2024 
+# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca
+# License: MIT
+
+#### Workspace setup ####
+library(tidyverse)
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
+library(scales)
+library(here)
+library(kableExtra)
+
+#### Test data ####
+## Test simulated_life_data
+# Load the dataset generated in R
+data1 <- read.csv("outputs/data/simulated_life_exp.csv", stringsAsFactors = FALSE)
+
+# 1. Data Structure Check
+# Check the number of rows and columns
+structure_check <- dim(data)[1] == length(years) & dim(data)[2] == length(life_expectancy_ranges) + 1
+print(paste("Data Structure Check:", ifelse(structure_check, "Passed", "Failed")))
+
+# 2. Range Check
+# Define the expected ranges
+expected_ranges <- c(77:80, 80:83, 65:72, 83:87, 70:76, 77:81)
+# Check if all values fall within the expected ranges
+range_check <- all(sapply(data[, -1], function(x) all(x >= min(expected_ranges) & x <= max(expected_ranges))))
+print(paste("Range Check:", ifelse(range_check, "Passed", "Failed")))
+
+# 3. Numeric Check
+# Check if all values are numeric
+numeric_check <- all(sapply(data, is.numeric))
+print(paste("Numeric Check:", ifelse(numeric_check, "Passed", "Failed")))
+
+## Test simulated_vote_data
+# Load the dataset generated in R
+data2 <- read.csv("outputs/data/simulated_vote_data.csv", stringsAsFactors = FALSE)
+
+# Test 1: Check number of rows and columns
+num_rows <- nrow(data2)
+num_cols <- ncol(data2)
+
+expected_num_rows <- 50 # Expecting 50 states
+expected_num_cols <- 8  # Expecting 7 columns (plus 1 for state_name)
+
+if (num_rows == expected_num_rows && num_cols == expected_num_cols) {
+  print("Test 1 passed: Correct number of rows and columns.")
+} else {
+  print("Test 1 failed: Incorrect number of rows or columns.")
+}
+
+# Test 2: Check for missing values
[…]

### 2024-02-12 scripts/Replication.R

@@ -1,0 +1,189 @@
+#### Preamble ####
+# Purpose: Paper Replication
+# Author: Adrian Ly, Hannah Yu
+# Date: 9 February 2024 
+# Contact: adrian.ly@mail.utoronto.ca
+# License: MIT
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(sf)
+library(readxl)
+library(knitr)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(data.table)
+library(RColorBrewer)
+library(ggpubr)
+library(scales)
+library(here)
+library(kableExtra)
+library(webshot)
+library(webshot2)
+
+#### Reading files ####
+ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+
+## CREATE FIGURE 1
+ghs_data <- ghs_data %>%
+  mutate(COUNTRY = case_when(
+    Country == "United States of America" ~ "United States",
+    Country == "Bolivia (Plurinational State of)" ~ "Bolivia",
+    # Add other necessary name changes here
+    TRUE ~ Country
+  ))
+
+world_ghs <- merge(world_map, ghs_data, by.x = "COUNTRY", by.y = "COUNTRY", all.x = TRUE)
+
+
+ggplot(data = world_ghs) +
+  geom_sf(aes(fill = `OVERALL SCORE`), color = "white") +
+  scale_fill_distiller(palette = "Spectral", na.value = "lightgrey", 
+                       name = "GHS Index Score", direction = 1,
+                       limits = c(min(ghs_data$`OVERALL SCORE`, na.rm = TRUE), max(ghs_data$`OVERALL SCORE`, na.rm = TRUE)),
+                       labels = comma) +  # Use comma to format numbers with commas
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5))
+
+# Save the plot to a file
+ggsave("outputs/graphs/GHS_Index_Map.png", width = 11, height = 8)
+
+## CREATE FIGURE 2
+
[…]

### 2024-02-12 scripts/Replication.R
updated replication

@@ -139,8 +139,16 @@ vote_data_by_state <- vote_share %>%
 vote_data_by_state <- vote_data_by_state |>
   mutate(gop_greater = if_else(total_gop_votes > total_dem_votes, 1, 0))
 
-# Create plot summarizing total number of votes by top 10 republican state
-sorted_states <- vote_data_by_state[order(-vote_data_by_state$total_diff), ]
+# Calculate proportion of Republican votes
+sorted_states <- vote_data_by_state
+sorted_states$proportion_republican <- sorted_states$total_gop_votes / sorted_states$total_votes
+
+# Sort by proportion of Republican votes
+sorted_states <- sorted_states[order(-sorted_states$proportion_republican), ]
+
+# Select top 10 states
+top_10_states <- head(sorted_states, 10)
+
 
 # Extract the top 10 states with the biggest difference
 top_10_states <- head(sorted_states, 10)
@@ -161,7 +169,7 @@ ggplot(plot_data_long, aes(x = state_name, y = votes, fill = party)) +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   scale_fill_manual(values = c("republican_votes" = "red", "democrat_votes" = "blue"),
-                    labels = c("Republican Votes", "Democratic Votes"))
+                    labels = c("Democratic Votes", "Republican Votes"))
 
 ggsave("outputs/graphs/Voter_Data.png", width = 10, height = 8)
 
@@ -186,4 +194,36 @@ save_kable(kable_img, "outputs/graphs/table.html")
 
 # Capture the table as an image
 webshot::webshot("outputs/graphs/table.html", "outputs/graphs/table.png")
+
+## CREATE TABLE 2
+
+# Filter the data for the years 2019 to 2021
+life_exp_subset <- life_exp_melted[life_exp_melted$year_id >= 2019 & life_exp_melted$year_id <= 2021, ]
+
+# Pivot the data to create a table and rename columns
+life_exp_table <- reshape2::dcast(life_exp_subset, year_id ~ race, value.var = "life_exp")
+colnames(life_exp_table) <- c("Year", "All races and origins", "Hispanic", "AIAN", "Asian", "Black", "White")
+
+# Create a table using knitr::kable
+
+kable_img2 <- kable(life_exp_table, col.names = c("Year", "All Races and Origins", "Hispanic", "AIAN", "Asian", "Black", "White"))
+
[…]

### 2024-02-12 outputs/paper/paper.qmd
updated replication
finished discussion and conclusion

@@
 
-# Create plot summarizing total number of votes by top 10 republican state
-#sorted_states <- vote_data_by_state[order(-vote_data_by_state$total_diff), ]
-
 # Calculate proportion of Republican votes
@@
 
-\newpage
-
 # Discussion
@@
 
-Weaknesses and next steps should also be included.
+In our research, we identified several weaknesses that underscored the gap between the U.S.'s high pandemic preparedness ranking and its actual response to the COVID-19 crisis.
+Notably, there was a significant underutilization of preparedness capacities, as the resources and infrastructures in place were not fully mobilized or applied effectively during the pandemic.
+This disconnect highlights a critical area for improvement in aligning preparedness with real-time response capabilities.
+Furthermore, the response inadequately addressed intrinsic social vulnerabilities, exacerbating disparities in healthcare access and outcomes among different racial and socioeconomic groups, and underscoring the need for more inclusive and equitable public health strategies.
+The heavy politicization of the pandemic response, coupled with inconsistent messaging from health authorities, significantly eroded public trust and compliance, emphasizing the importance of depoliticizing public health measures and enhancing communication strategies.
+Additionally, shortcomings in data and surveillance systems, including delays in testing, inadequate genetic sequencing, and a lack of comprehensive surveillance, hindered informed decision-making and targeted interventions, pointing to a crucial need for investment in data infrastructure and capabilities.
+The analysis also suggests that the U.S.
+could benefit from more robust international benchmarking and learning from the diverse responses of other countries to health emergencies, which would involve understanding different strategies and their effectiveness beyond mere outcome comparisons.
+
+To address these challenges and strengthen the nation's resilience to future health emergencies, several next steps are recommended.
+Strengthening the linkages between preparedness and response is imperative, ensuring that capacities are not only available but also readily deployable and adaptable to the dynamics of a health crisis.
+This includes fostering agile and responsive systems capable of rapid mobilization.
+Addressing the social determinants of health is essential, with strategies aimed at mitigating the impact of social vulnerabilities through equitable healthcare access, support for marginalized communities, and targeted protective measures for vulnerable populations.
+Enhancing communication and public trust is vital, necessitating clear, consistent, and transparent communication from health authorities, alongside efforts to depoliticize health measures.
+Improving data and surveillance systems is critical to providing real-time insights, expanding genetic sequencing capabilities, and establishing standardized data protocols.
+Lastly, actively engaging in global health networks to exchange experiences, learn from global successes and failures, and collaborate on best practices is crucial for a more effective and equitable response that maximizes the full potential of preparedness capacities.
 
 # Conclusion
+
+Despite the United States' top ranking in the Global Health Security Index as the most prepared nation for pandemics, its actual response to the COVID-19 pandemic fell short of expectations, leading to disproportionately high death rates.
+This discrepancy is attributed to several critical factors highlighted in the paper.
+The U.S.
+did not fully capitalize on its pandemic preparedness resources, with early testing blunders and a lack of a unified national testing approach hindering its response.
+Additionally, inherent vulnerabilities such as the significant portion of the population in congregated settings like nursing homes and prisons increased susceptibility to virus spread and severe health outcomes.
+The politicization of the pandemic response further compounded these issues, resulting in varied adherence to public health guidelines across states and political lines, thereby undermining the response's effectiveness.
+Moreover, inconsistent public health communications and the challenge of accessing standardized, quality data impeded the implementation of localized, effective interventions.
+The pandemic also highlighted and intensified existing socio-economic disparities, disproportionately impacting marginalized communities who faced higher risks and adverse outcomes, a situation that the response efforts failed to sufficiently mitigate.
+
 \newpage

### 2024-02-13 outputs/paper/paper.qmd
wrote methdology and data cleaning, fixed code

@@
 library(kableExtra)
+library(gt)
 
@@
 
-ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
-deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+# ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+# world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+# life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+# vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+# deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+
+ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
+world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
+life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
+vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
+deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
 
@@
 The datasets utilized in this paper were mainly obtained from the `original paper` [@citearticle].
-Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and @citedata2.
+The data used by the paper was compiled by various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11.
+Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and @citedata2.The political party support data we utilized from @citedata1 was compiled from @data12.
 
@@
 
-`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], and `scales` [@citeScales].
+This paper will to replicate the data that was originally presented by @citearticle, as previously mentioned.
+We then transform the replication to our own data analysis incorporating the COVID statistics adn political party support data.
+We seek to use our initial replication and following own data anlysis to provide a fresh part of the original paper, enhancing the with more data.
+Different from the original paper that spend a lot of time talking about how badly prepared US was against other country, We also specifically diverge our focus of the political and socioeconomic factors that impacted US's COVID outcome.
+
+`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], 'gt' [@citegt] and `scales` [@citeScales].
+
+## Data cleaning
+From all the data provided from @citearticle, @citedata1, and @citedata2, we mainly utilize the @data3, @data8, @data12, and @citedata2.
+The data from @citearticle are quite organized and do not require excessive cleaning for us to proceed.
+
+Since we aim to analyze people's political preference on a state level and the data we collected from @data12 is on a county level, we create a new dataframe "vote_data_by_state" to summarize all the votes by state.
+We utilize the "group_by" function to form new variables on a state level.
+We calculate each of the sum of the total votes, total Republican party votes, total Democratic party votes, total vote difference between the two parties by county and created new variables that calculate the sum of each of these.
+We also create variables on the average percentage of vote each party receives and the difference by each state.
+
+To find the total number of votes in Top 10 states with highest proportion of Republican votes, we create another new dataframe "sorted_states" from the "vote_data_by_state" dataframe.
+We include a new variable "proportion_republican" that calculates the proportion of votes the Republican party receives from each state and ordered the states by the highest proportion of Republican votes.
+
+@datadata2 does not have column names for each column.To retrieve data by column for we assigned column names "State" and "Death_Rate" to the variables of interests.
+And then, to find the top 10 states with highest death rates, we ranked the states based on death rate.
+
+Lastly, to create the final figure that incorporates the COVID death rate and people's political preference by state, we merged the two data from @data12 and @citedata2.
+And subsequently did our work from there.
 
 ## Data Measurement
-
-## Data cleaning
 
@@
 
-# Print the table using kableExtra
-kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c") %>%
-  kable_styling(full_width = FALSE)
+# Create a gt table
+table_gt <- gt(table) %>%
+  tab_header(
+    title = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)"
+  )
+
+# Print the gt table
+table_gt
 
@@
 #| warning: false
+
 # Filter the data for the years 2019 to 2021
 life_exp_subset <- life_exp_melted[life_exp_melted$year_id >= 2019 & life_exp_melted$year_id <= 2021, ]
-
 # Pivot the data to create a table and rename columns
@@
 colnames(life_exp_table) <- c("Year", "All races and origins", "Hispanic", "AIAN", "Asian", "Black", "White")
-
-# Create a table using knitr::kable
-
-kable(life_exp_table, col.names = c("Year", "All Races and Origins", "Hispanic", "AIAN", "Asian", "Black", "White"))
+# Create a gt table
+life_exp_gt <- gt(life_exp_table, rowname_col = "Race") %>%
+  tab_header(
+    title = "Life Expectancy by Race (2019-2021)"
+  )
+# Print the gt table
+life_exp_gt
 

### 2024-02-13 outputs/paper/paper.qmd
updated paper, llm usage and readme

--- 2024-02-12 outputs/paper/paper.qmd
@@
 The datasets utilized in this paper were mainly obtained from the `original paper` [@citearticle].
-Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and @citedata2.
+The data used by the paper was compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11.
+Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and [@citedata2].
+The political party support data we utilized from @citedata1 was compiled from @data12.
 
@@
 
-`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], and `scales` [@citeScales].
+This paper will replicate the data that was originally presented by @citearticle, as previously mentioned.
+We then transform the replication to our data analysis incorporating the COVID statistics and political party support data.
+We seek to use our initial replication and following own data analysis to provide a fresh part of the original paper, enhancing the with more data.
+Different from the original paper which spent a lot of time talking about how badly prepared the US was against other countries, We also specifically diverge our focus to the political and socioeconomic factors that impacted the US's COVID outcome.
+
+`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], `gt` [@citegt] and `scales` [@citeScales].
+
+## Data cleaning
+
+From all the data provided from @citearticle, @citedata1, and @citedata2, we mainly utilize the @data3, @data8, @data12, and @citedata2.
+The data from @citearticle are quite organized and do not require excessive cleaning for us to proceed.
+
+Since we aim to analyze people's political preferences on a state level and the data we collected from @data12 is on a county level, we create a new data frame "vote_data_by_state" to summarize all the votes by state.
+We utilize the "group_by" function to form new variables on a state level.
+We calculated each the sum of the total votes, total Republican party votes, total Democratic party votes, and total vote difference between the two parties by county and created new variables that calculate the sum of each of these.
+We also create variables on the average percentage of votes each party receives and the difference by each state.
+
+To find the total number of votes in the Top 10 states with the highest proportion of Republican votes, we create another new data frame "sorted_states" from the "vote_data_by_state" data frame.
+We include a new variable "proportion_republican" that calculates the proportion of votes the Republican party receives from each state and orders the states by the highest proportion of Republican votes.
+The data from @data2 does not have column names for each column.
+To retrieve data by column we assigned column names "State" and "Death_Rate" to the variables of interest.
+And then, to find the top 10 states with the highest death rates, we ranked the states based on death rate.
+Lastly, to create the final figure that incorporates the COVID death rate and people's political preference by state, we merged the two data from @data12 and @citedata2.
+And subsequently did our work from there.
 
@@
 
-## Data cleaning
+### Measurement Techniques
 
-## Data Visualization
+Manual Data Extraction involved careful review of reports and tables to ensure accurate data capture, particularly for datasets like life expectancy where this meticulous approach was paramount; this was coupled with Standardization and Adjustment processes where excess death data and age-standardized excess death rates underwent significant processing, including the comparison of observed deaths during the pandemic to expected deaths based on historical trends, with adjustments made for the age distribution of the population; concurrently, Geospatial Data Handling was employed for mapping the GHS Index scores, utilizing Geographic Information Systems (GIS) to visually represent pandemic preparedness across various countries, and this was complemented by Real-Time Data Aggregation, which entailed aggregating COVID-19 death data in real-time from authoritative sources such as the Johns Hopkins University Dashboard, necessitating continuous data monitoring and updates to maintain accuracy and relevance.
+
+### Adjustments and Considerations
+
+Accounting for underreporting through the measurement of excess deaths provides a more comprehensive view of the pandemic's impact by capturing deaths that may have been missed or misclassified as COVID-19, while the process of age standardization adjusts these excess death rates to a standard age distribution, enabling meaningful comparisons between countries with varying demographic profiles.
+However, the analysis had to navigate challenges associated with varying data quality and availability across different countries, especially given the diverse COVID-19 reporting standards and testing capacities.
+Moreover, the study also acknowledges the significant influence of political and social factors on the pandemic's outcomes, which, despite being more challenging to quantify, are essential for a holistic understanding of the pandemic's impact.
 
@@
 @fig-ELE illustrates the trend of life expectancy at birth across different racial groups over time.
-@tbl-life-exp provides a more detailed breakdown of the life expectancy before and during COVID for...
+@tbl-life-exp provides a more detailed breakdown of the life expectancy before and during COVID for Asian, American Indian, Alaska Native and other racial communities.
+
+@fig-GHS also allows us to visualize the Global Health Security (GHS) Index which is designed to assess a country's capabilities to prevent, detect, and respond to significant infectious disease outbreaks.
+A higher Global Health Security (GHS) Index score indicates a stronger health system capacity for pandemic readiness.
+The GHS Index assesses the readiness of countries for pandemics and other significant biological threat emergencies, using a comprehensive framework that includes measurements grouped into six broad categories: prevention, detection, response, health system capacity, compliance with international norms, and risk environment.
+Higher scores suggest that a country has a more robust health system capacity, with better capabilities to prevent, detect, respond to, and mitigate the spread of an epidemic.
+
+Conversely, a lower GHS Index score implies weaker health system capacities, suggesting that the country may be less prepared to handle pandemics and other significant biological threats.
+Lower scores indicate deficiencies in one or more of the critical areas assessed by the Index, which could potentially hinder the country's ability to effectively manage and control the spread of infectious diseases.
 
@@
   geom_bar(stat = "identity") +
-  labs(title = "Total Number of Votes by Top 10 Republican State", x = "State", y = "Total Votes", fill = "Party") +
+  labs(x = "State", y = "Total Votes", fill = "Party") +
   theme_minimal() +
@@
 
-# Print the table using kableExtra
-kable(table, caption = "Top 10 States with Highest Death Rates from COVID-19 (per 100,000 people)", align = "c") %>%
-  kable_styling(full_width = FALSE)
+# Create a gt table
+table_gt <- gt(table)
+
+# Print the gt table
+table_gt
 
 ```
+
+\newpage
 
@@
 colnames(life_exp_table) <- c("Year", "All races and origins", "Hispanic", "AIAN", "Asian", "Black", "White")
-
-# Create a table using knitr::kable
-
-kable(life_exp_table, col.names = c("Year", "All Races and Origins", "Hispanic", "AIAN", "Asian", "Black", "White"))
+# Create a gt table
+life_exp_gt <- gt(life_exp_table, rowname_col = "Race") 
+# Print the gt table
+life_exp_gt
 
@@
 
-Following from our previous analysis regarding individuals' political affiliations, we have developed Figure @fig-deaths-rates-votes, which encompasses all 50 states of the US along with their political leanings based on which party garnered the majority votes.
+Following our previous analysis regarding individuals' political affiliations, we have developed Figure @fig-deaths-rates-votes, which encompasses all 50 states of the US along with their political leanings based on which party garnered the majority votes.
 This information is juxtaposed against their respective COVID-19 death rates.
@@
   labs(x = "Republican Votes", y = "COVID-19 Death Rate") +
-  ggtitle("COVID-19 Death Rates vs. Republican Votes") +
   theme_minimal() + # Use minimal theme for cleaner appearance

### 2024-02-13 outputs/paper/paper.qmd
new table, new references, lots of other changes

@@
 
-ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
-deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
+world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
+life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
+vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
+deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
+
+# ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+# world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+# life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+# vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+# deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
 
@@
 
-# Data {#sec-data}
+@fig-GHS allows us to visualize the Global Health Security (GHS) Index which is designed to assess a country's capabilities to prevent, detect, and respond to significant infectious disease outbreaks.
+A higher Global Health Security (GHS) Index score indicates a stronger health system capacity for pandemic readiness.
+The GHS Index assesses the readiness of countries for pandemics and other significant biological threat emergencies, using a comprehensive framework that includes measurements grouped into six broad categories: prevention, detection, response, health system capacity, compliance with international norms, and risk environment.
+Higher scores suggest that a country has a more robust health system capacity, with better capabilities to prevent, detect, respond to, and mitigate the spread of an epidemic.
 
-## Source
-
-The datasets utilized in this paper were mainly obtained from the `original paper` [@citearticle].
-The data used by the paper was compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11.
-Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and [@citedata2].
-The political party support data we utilized from @citedata1 was compiled from @data12.
-
-@citedata1 discusses the long-term impacts of COVID-related school closures. From this source, we utilized the dataset on voting shares during the 2020 election by county.
-@citedata2 summarizes COVID-19 death rates in the United States as of March 2023, organized by state.
-Analyzing results from both datasets allows us to explore the relationship between political affiliation and COVID-19 outcomes.
-Our reproduction aims to fill these gaps and includes tables and graphs not presented in the original paper to support our findings.
-
-## Methodology
-
-This paper will replicate the data that was originally presented by @citearticle, as previously mentioned.
-We then transform the replication to our data analysis incorporating the COVID statistics and political party support data.
-We seek to use our initial replication and following own data analysis to provide a fresh part of the original paper, enhancing the with more data.
-Different from the original paper which spent a lot of time talking about how badly prepared the US was against other countries, We also specifically diverge our focus to the political and socioeconomic factors that impacted the US's COVID outcome.
-
-`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], `gt` [@citegt] and `scales` [@citeScales].
-
-## Data cleaning
-
-From all the data provided from @citearticle, @citedata1, and @citedata2, we mainly utilize the @data3, @data8, @data12, and @citedata2.
-The data from @citearticle are quite organized and do not require excessive cleaning for us to proceed.
-
-Since we aim to analyze people's political preferences on a state level and the data we collected from @data12 is on a county level, we create a new data frame "vote_data_by_state" to summarize all the votes by state.
-We utilize the "group_by" function to form new variables on a state level.
-We calculated each the sum of the total votes, total Republican party votes, total Democratic party votes, and total vote difference between the two parties by county and created new variables that calculate the sum of each of these.
-We also create variables on the average percentage of votes each party receives and the difference by each state.
-
-To find the total number of votes in the Top 10 states with the highest proportion of Republican votes, we create another new data frame "sorted_states" from the "vote_data_by_state" data frame.
-We include a new variable "proportion_republican" that calculates the proportion of votes the Republican party receives from each state and orders the states by the highest proportion of Republican votes.
-The data from @data2 does not have column names for each column.
-To retrieve data by column we assigned column names "State" and "Death_Rate" to the variables of interest.
-And then, to find the top 10 states with the highest death rates, we ranked the states based on death rate.
-Lastly, to create the final figure that incorporates the COVID death rate and people's political preference by state, we merged the two data from @data12 and @citedata2.
-And subsequently did our work from there.
-
-## Data Measurement
-
-### Measurement Techniques
-
-Manual Data Extraction involved careful review of reports and tables to ensure accurate data capture, particularly for datasets like life expectancy where this meticulous approach was paramount; this was coupled with Standardization and Adjustment processes where excess death data and age-standardized excess death rates underwent significant processing, including the comparison of observed deaths during the pandemic to expected deaths based on historical trends, with adjustments made for the age distribution of the population; concurrently, Geospatial Data Handling was employed for mapping the GHS Index scores, utilizing Geographic Information Systems (GIS) to visually represent pandemic preparedness across various countries, and this was complemented by Real-Time Data Aggregation, which entailed aggregating COVID-19 death data in real-time from authoritative sources such as the Johns Hopkins University Dashboard, necessitating continuous data monitoring and updates to maintain accuracy and relevance.
-
-### Adjustments and Considerations
-
-Accounting for underreporting through the measurement of excess deaths provides a more comprehensive view of the pandemic's impact by capturing deaths that may have been missed or misclassified as COVID-19, while the process of age standardization adjusts these excess death rates to a standard age distribution, enabling meaningful comparisons between countries with varying demographic profiles.
-However, the analysis had to navigate challenges associated with varying data quality and availability across different countries, especially given the diverse COVID-19 reporting standards and testing capacities.
-Moreover, the study also acknowledges the significant influence of political and social factors on the pandemic's outcomes, which, despite being more challenging to quantify, are essential for a holistic understanding of the pandemic's impact.
+Conversely, a lower GHS Index score implies weaker health system capacities, suggesting that the country may be less prepared to handle pandemics and other significant biological threats.
+Lower scores indicate deficiencies in one or more of the critical areas assessed by the Index, which could potentially hinder the country's ability to effectively manage and control the spread of infectious diseases.
 
@@
 
+# Data {#sec-data}
+
+## Source
+
+The datasets utilized in this paper were mainly obtained from the `original paper` [@citearticle].
+The data used by the paper was compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11.
+Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and [@citedata2].
+The political party support data we utilized from @citedata1 was compiled from @data12.
+
+@citedata1 discusses the long-term impacts of COVID-related school closures. From this source, we utilized the dataset on voting shares during the 2020 election by county.
+@citedata2 summarizes COVID-19 death rates in the United States as of March 2023, organized by state.
+Analyzing results from both datasets allows us to explore the relationship between political affiliation and COVID-19 outcomes.
+Our reproduction aims to fill these gaps and includes tables and graphs not presented in the original paper to support our findings.
+
+## Methodology
+
+This paper will replicate the data that was originally presented by @citearticle, as previously mentioned.
+We then transform the replication to our data analysis incorporating the COVID statistics and political party support data.
+We seek to use our initial replication and following own data analysis to provide a fresh part of the original paper, enhancing the with more data.
+Different from the original paper which spent a lot of time talking about how badly prepared the US was against other countries, We also specifically diverge our focus to the political and socioeconomic factors that impacted the US's COVID outcome.
+
+`R` [@citeR] was the language and environment used for the bulk of this analysis, alongside `tidyverse` [@citeTidyVerse], `sf` [@citeSF], `readxl` [@citeReadXL], `knitr` [@citeKnitR], `janitor` [@citeJanitor], `lubridate` [@citeLubridate], `dplyr` [@citeDplyr], `data.table` [@citeDataTable], `RColorBrewer` [@citeRColorBrewer], `ggpubr` [@citeggpubr], `ggplot2` [@citeggplot], `here` [@citeHere], `kableExtra` [@citeKableExtra], `webshot` [@citewebshot], `webshot2` [@citewebshot2], `gt` [@citegt] and `scales` [@citeScales].
+
+## Data cleaning
+
+From all the data provided from @citearticle, @citedata1, and @citedata2, we mainly utilize  @data3, @data8, @data12, and @citedata2.
+The data from @citearticle is well-organized and require minimal cleaning for analysis.
+
+The data collected from @data12 is at the county level as depicted on @tbl-sample.
+However, since our analysis focuses on people's political preferences at the state level, we create a new data frame called "vote_data_by_state" to aggregate voting data by state.
+We utilize the "group_by" function to generate new variables at the state level, calculating the total votes, total Republican party votes, total Democratic party votes, and the total vote difference between the two parties by county.
+Additionally, We also create variables for the average percentage of votes each party receives and the difference by each state.
+
+To determine the total number of votes in the Top 10 states with the highest proportion of Republican votes, we create another new data frame called "sorted_states" from the "vote_data_by_state" data frame.
+We include a new variable "proportion_republican" that calculates the proportion of votes the Republican party receives from each state and orders the states by the highest proportion of Republican votes.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-sample
+#| tbl-cap: Sample of Vote Share Data
+#| warning: false
+
+# Take a sample of the data (let's say the first 5 rows)
+vote_share_sample <- head(vote_share, 5)
+
+# Print the sample table
+kable(vote_share_sample, 
+      format = "markdown", 
+      col.names = c("State Name", "County FIPS", "County Name", "Votes GOP", "Votes DEM", "Total Votes", "Difference", "Percentage GOP", "Percentage DEM", "Percentage Point Difference")
+)
+
+```
+
+The data from @data2 lacks column names for each column.
+To retrieve data by column, we assign column names "State" and "Death_Rate" to the variables of interest.
+Subsequently, to find the top 10 states with the highest death rates, we rank the states based on death rate.
+Lastly, to create the final figure that incorporates the COVID death rate and people's political preferences by state, we merge the two datasets from @data12 and @citedata2 and proceed with our analysis.
+
+## Data Measurement
+
+### Measurement Techniques
+
+Manual Data Extraction involved careful review of reports and tables to ensure accurate data capture, particularly for datasets like life expectancy where this meticulous approach was paramount; this was coupled with Standardization and Adjustment processes where excess death data and age-standardized excess death rates underwent significant processing, including the comparison of observed deaths during the pandemic to expected deaths based on historical trends, with adjustments made for the age distribution of the population; concurrently, Geospatial Data Handling was employed for mapping the GHS Index scores, utilizing Geographic Information Systems (GIS) to visually represent pandemic preparedness across various countries, and this was complemented by Real-Time Data Aggregation, which entailed aggregating COVID-19 death data in real-time from authoritative sources such as the Johns Hopkins University Dashboard, necessitating continuous data monitoring and updates to maintain accuracy and relevance.
+
+### Adjustments and Considerations
+
+Accounting for underreporting through the measurement of excess deaths provides a more comprehensive view of the pandemic's impact by capturing deaths that may have been missed or misclassified as COVID-19, while the process of age standardization adjusts these excess death rates to a standard age distribution, enabling meaningful comparisons between countries with varying demographic profiles.
+However, the analysis had to navigate challenges associated with varying data quality and availability across different countries, especially given the diverse COVID-19 reporting standards and testing capacities.
+Moreover, the study also acknowledges the significant influence of political and social factors on the pandemic's outcomes, which, despite being more challenging to quantify, are essential for a holistic understanding of the pandemic's impact.
+
 # Results
@@
 @fig-ELE illustrates the trend of life expectancy at birth across different racial groups over time.
+Unfortunately, due to the unavailability of data over time for Asian, American Indian, and Alaska Native communities, they have been excluded from the time series graph.
 @tbl-life-exp provides a more detailed breakdown of the life expectancy before and during COVID for Asian, American Indian, Alaska Native and other racial communities.
 
-@fig-GHS also allows us to visualize the Global Health Security (GHS) Index which is designed to assess a country's capabilities to prevent, detect, and respond to significant infectious disease outbreaks.
-A higher Global Health Security (GHS) Index score indicates a stronger health system capacity for pandemic readiness.
-The GHS Index assesses the readiness of countries for pandemics and other significant biological threat emergencies, using a comprehensive framework that includes measurements grouped into six broad categories: prevention, detection, response, health system capacity, compliance with international norms, and risk environment.
-Higher scores suggest that a country has a more robust health system capacity, with better capabilities to prevent, detect, respond to, and mitigate the spread of an epidemic.
+An intriguing observation from @fig-ELE is the consistently higher life expectancy among Hispanic individuals compared to other groups, even amidst the challenges posed by COVID-19.
+On the other hand, Black individuals have consistently exhibited lower life expectancy, which further declined notably in 2021, reaching just over 70.8 years old.
+The life expectancy trends of white people and all other races and origins remain close together throughout the 14 years, with minimum variance.
 
-Conversely, a lower GHS Index score implies weaker health system capacities, suggesting that the country may be less prepared to handle pandemics and other significant biological threats.
-Lower scores indicate deficiencies in one or more of the critical areas assessed by the Index, which could potentially hinder the country's ability to effectively manage and control the spread of infectious diseases.
-
-Unfortunately, due to the unavailability of data over time for Asian, American Indian, and Alaska Native communities, they have been excluded from the time series graph.
-
-An intriguing observation is the consistently higher life expectancy among Hispanic individuals compared to other groups, even amidst the challenges posed by COVID-19.
-On the other hand, Black individuals l have consistently exhibited lower life expectancy, which further declined notably in 2021, reaching just over 70.8 years old.
-The life expectancy trends of white people and all other races and origins remain close together throughout the 14 years, with minimum variance.
-Specifically, the life expectancy for White individuals decreased from 78.8 to 76.4 years old from pre-pandemic levels in 2019 to 2021, while for Hispanic individuals, it dropped by 4.2 years, and for Black individuals, it declined by 4 years during the same period.
+While all racial groups experienced declines in their average life expectancy, the declines vary greatly.
+The life expectancy for White individuals decreased from 78.8 to 76.4 years old from pre-pandemic levels in 2019 to 2021, while for Hispanic individuals, it dropped by 4.2 years, and for Black individuals, it declined by 4 years during the same period.
+Our data aligns with the findings of @andrasfay_goldman_2022, which states that the Black and Hispanic populations experienced twice as much life expectancy decrease as the white population."
 
@@
 While the declines in life expectancy are smaller in magnitude from 2020 to 2021, they are notably minimal for Asians and Hispanics, with reductions of -0.1 and -0.2 years respectively.
-(Insert some sources tmrw)
 
-Numerous sources have indicated a correlation between a state's political affiliation and its handling of COVID issues (sources to be inserted tomorrow).
+Various studies have confirmed the existence of correlation between a state's political affiliation and its handling of COVID issues (@datz_2022).
 To testify to the claim, we collected the voting data for each of the 50 states for the 2020 presidential election.
@@
 Notably, We found that six of the ten top Republican states made a reappearance in the top death rates table; these states are Oklahoma, West Virginia, Arkansas, Alabama, Tennessee, and Kentucky.
-(Insert quotes about the republican party and COVID tmrw)
+The higher mortality rates observed in Republican-leaning states may be attributed to their preference for less stringent measures compared to Democratic-leaning states, which tend to favor stricter measures (@vandusky-allen_shvetsova_2021).
 

### 2024-02-13 outputs/paper/paper.tex
new table, new references, lots of other changes

@@
 \makeatother
-% definitions for citeproc citations
-\NewDocumentCommand\citeproctext{}{}
-\NewDocumentCommand\citeproc{mm}{%
-  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
-\makeatletter
- % allow citations to break across lines
- \let\@cite@ofmt\@firstofone
- % avoid brackets around text for \cite:
- \def\@biblabel#1{}
- \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
-\makeatother
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
- {\begin{list}{}{%
-  \setlength{\itemindent}{0pt}
-  \setlength{\leftmargin}{0pt}
-  \setlength{\parsep}{0pt}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
   % turn on hanging indent if param 1 is 1
   \ifodd #1
-   \setlength{\leftmargin}{\cslhangindent}
-   \setlength{\itemindent}{-1\cslhangindent}
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
   \fi
   % set entry spacing
-  \setlength{\itemsep}{#2\baselineskip}}}
- {\end{list}}
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
 \usepackage{calc}
-\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
 \newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
@@
 \usepackage{xcolor}
+\usepackage{caption}
 \KOMAoption{captions}{tableheading}
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
 \makeatletter
@@
 \makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
 \makeatother
 \makeatletter
-\@ifpackageloaded{caption}{}{\usepackage{caption}}
-\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
 \makeatother
@@
 \fi
-\usepackage{bookmark}
-
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
@@
 \author{Adrian Ly \and Sakhil Goel \and Hannah Yu}
-\date{February 12, 2024}
+\date{February 13, 2024}
 
@@
 \end{abstract}
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, frame hidden, enhanced, interior hidden, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 }
-\section{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 
-\emph{The introduction is self-contained and tells a reader everything
-they need to know including: 1) broader context to motivate; 2) some
-detail about what the paper is about; 3) a clear gap that needs to be
-filled; 4) what was done; 5) what was found; 6) why it is important; 7)
-the structure of the paper. A reader should be able to read only the
-introduction and know what was done, why, and what was found. Likely 3
-or 4 paragraphs, or 10 per cent of total. The estimand is clearly stated
-in the introduction.}
+Figure~\ref{fig-GHS} allows us to visualize the Global Health Security
+(GHS) Index which is designed to assess a country's capabilities to
+prevent, detect, and respond to significant infectious disease
+outbreaks. A higher Global Health Security (GHS) Index score indicates a
+stronger health system capacity for pandemic readiness. The GHS Index
+assesses the readiness of countries for pandemics and other significant
+biological threat emergencies, using a comprehensive framework that
+includes measurements grouped into six broad categories: prevention,
+detection, response, health system capacity, compliance with
+international norms, and risk environment. Higher scores suggest that a
+country has a more robust health system capacity, with better
+capabilities to prevent, detect, respond to, and mitigate the spread of
+an epidemic.
 
-\section{Data}\label{sec-data}
+Conversely, a lower GHS Index score implies weaker health system
+capacities, suggesting that the country may be less prepared to handle
+pandemics and other significant biological threats. Lower scores
+indicate deficiencies in one or more of the critical areas assessed by
+the Index, which could potentially hinder the country's ability to
+effectively manage and control the spread of infectious diseases.
 
-Our results are summarized in (\textbf{fig:fig-GHS?}).
+\begin{figure}
 
-\subsection{Source}\label{source}
+{\centering \includegraphics{paper_files/figure-pdf/fig-GHS-1.pdf}
+
+}
+
+\caption{\label{fig-GHS}Global Health Security Index Scores by Country}
+
+\end{figure}
+
+\hypertarget{sec-data}{%
+\section{Data}\label{sec-data}}
+
+\hypertarget{source}{%
+\subsection{Source}\label{source}}
 
 The datasets utilized in this paper were mainly obtained from the
-\texttt{original\ paper} (Nuzzo and Ledesma 2023). Additionally, to
-address the original paper's lack of US Covid statistics and political
-party support data, we incorporated information from Jack and Oster
-(2023) and Elflein (2023).
+\texttt{original\ paper} (Nuzzo and Ledesma 2023). The data used by the
+paper was compiled from various sources such as the Arias et al. (2022),
+Arias and Xu (2022), Bell and Nuzzo (2021), Systems Science and
+Engineering (2023), Dong, Du, and Gardner (2020), \emph{COVID-19 Excess
+Mortality Estimates 2020-2021} (2022), Ledesma et al. (2023), Health
+Statistics (2022), {``Death Rate, Crude (Per 1,000 People)
+(SP.DYN.CDRT.IN)''} (2022), Economic and Division (2022), and Wang et
+al. (2022). Additionally, to address the original paper's lack of US
+COVID statistics and political party support data, we incorporated
+information from Jack and Oster (2023) and (Elflein 2023). The political
+party support data we utilized from Jack and Oster (2023) was compiled
+from McGovern (2009-\/-2020).
 
@@
 school closures. From this source, we utilized the dataset on voting
-shares during the 2020 election by county. Elflein (2023) provides
-summaries of COVID-19 death rates in the United States as of March 2023,
-organized by state. Analyzing results from both datasets allows us to
-explore the relationship between political affiliation and COVID-19
-outcomes.Our reproduction aims to fill these gaps and also includes
-tables and graphs that were not presented in the original paper to
-support our findings.
+shares during the 2020 election by county. Elflein (2023) summarizes
+COVID-19 death rates in the United States as of March 2023, organized by
+state. Analyzing results from both datasets allows us to explore the
+relationship between political affiliation and COVID-19 outcomes. Our
+reproduction aims to fill these gaps and includes tables and graphs not
+presented in the original paper to support our findings.
 
-\subsection{Methodology}\label{methodology}
+\hypertarget{methodology}{%
+\subsection{Methodology}\label{methodology}}
+
+This paper will replicate the data that was originally presented by
+Nuzzo and Ledesma (2023), as previously mentioned. We then transform the
+replication to our data analysis incorporating the COVID statistics and
+political party support data. We seek to use our initial replication and
+following own data analysis to provide a fresh part of the original
+paper, enhancing the with more data. Different from the original paper
+which spent a lot of time talking about how badly prepared the US was
+against other countries, We also specifically diverge our focus to the
+political and socioeconomic factors that impacted the US's COVID
+outcome.
 
@@
 \texttt{kableExtra} (Zhu 2024), \texttt{webshot} (Chang 2023a),
-\texttt{webshot2} (Chang 2023b), \texttt{gt} (Iannone et al. 2024), and
+\texttt{webshot2} (Chang 2023b), \texttt{gt} (Iannone et al. 2024) and
 \texttt{scales} (Wickham, Pedersen, and Seidel 2023).
 
-\subsection{Data Measurement}\label{data-measurement}
+\hypertarget{data-cleaning}{%
+\subsection{Data cleaning}\label{data-cleaning}}
 
-\emph{A thorough discussion of measurement, relating to the dataset, is
-provided in the data section.}
+From all the data provided from Nuzzo and Ledesma (2023), Jack and Oster
+(2023), and Elflein (2023), we mainly utilize Bell and Nuzzo (2021),
+Health Statistics (2022), McGovern (2009-\/-2020), and Elflein (2023).
+The data from Nuzzo and Ledesma (2023) is well-organized and require
+minimal cleaning for analysis.
 
-\subsection{Data cleaning}\label{data-cleaning}
+The data collected from McGovern (2009-\/-2020) is at the county level
+as depicted on Table~\ref{tbl-sample}. However, since our analysis
+focuses on people's political preferences at the state level, we create
+a new data frame called ``vote\_data\_by\_state'' to aggregate voting
+data by state. We utilize the ``group\_by'' function to generate new
+variables at the state level, calculating the total votes, total
+Republican party votes, total Democratic party votes, and the total vote
+difference between the two parties by county. Additionally, We also
+create variables for the average percentage of votes each party receives
+and the difference by each state.
+
+To determine the total number of votes in the Top 10 states with the
+highest proportion of Republican votes, we create another new data frame
+called ``sorted\_states'' from the ``vote\_data\_by\_state'' data frame.
+We include a new variable ``proportion\_republican'' that calculates the
+proportion of votes the Republican party receives from each state and
+orders the states by the highest proportion of Republican votes.
+
+\hypertarget{tbl-sample}{}
+\begin{longtable}[]{@{}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0791}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0863}}
+  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1079}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0719}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0719}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0863}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0791}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1079}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1079}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.2014}}@{}}
+\caption{\label{tbl-sample}Sample of Vote Share Data}\tabularnewline
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+State Name
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+County FIPS
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+County Name
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Votes GOP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Votes DEM
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Total Votes
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Difference
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Percentage GOP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Percentage DEM
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Percentage Point Difference
+\end{minipage} \\
+\midrule\noalign{}
+\endfirsthead
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedright
+State Name
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+County FIPS
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+County Name
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Votes GOP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Votes DEM
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Total Votes
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Difference
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Percentage GOP
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Percentage DEM
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+Percentage Point Difference
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+Alabama & 1001 & Autauga County & 19838 & 7503 & 27770 & 12335 &
+0.7143680 & 0.2701837 & 0.4441844 \\
+Alabama & 1003 & Baldwin County & 83544 & 24578 & 109679 & 58966 &
+0.7617137 & 0.2240903 & 0.5376234 \\
+Alabama & 1005 & Barbour County & 5622 & 4816 & 10518 & 806 & 0.5345123
+& 0.4578817 & 0.0766305 \\
+Alabama & 1007 & Bibb County & 7525 & 1986 & 9595 & 5539 & 0.7842626 &
+0.2069828 & 0.5772798 \\
+Alabama & 1009 & Blount County & 24711 & 2640 & 27588 & 22071 &
+0.8957155 & 0.0956938 & 0.8000217 \\
+\end{longtable}
+
+The data from Arias and Xu (2022) lacks column names for each column. To
+retrieve data by column, we assign column names ``State'' and
+``Death\_Rate'' to the variables of interest. Subsequently, to find the
+top 10 states with the highest death rates, we rank the states based on
+death rate. Lastly, to create the final figure that incorporates the
+COVID death rate and people's political preferences by state, we merge
+the two datasets from McGovern (2009-\/-2020) and Elflein (2023) and
+proceed with our analysis.
+
+\hypertarget{data-measurement}{%
+\subsection{Data Measurement}\label{data-measurement}}
+
+\hypertarget{measurement-techniques}{%
+\subsubsection{Measurement Techniques}\label{measurement-techniques}}
+
+Manual Data Extraction involved careful review of reports and tables to
+ensure accurate data capture, particularly for datasets like life
+expectancy where this meticulous approach was paramount; this was
+coupled with Standardization and Adjustment processes where excess death
+data and age-standardized excess death rates underwent significant
+processing, including the comparison of observed deaths during the
+pandemic to expected deaths based on historical trends, with adjustments
+made for the age distribution of the population; concurrently,
+Geospatial Data Handling was employed for mapping the GHS Index scores,
+utilizing Geographic Information Systems (GIS) to visually represent
+pandemic preparedness across various countries, and this was
+complemented by Real-Time Data Aggregation, which entailed aggregating
+COVID-19 death data in real-time from authoritative sources such as the
+Johns Hopkins University Dashboard, necessitating continuous data
+monitoring and updates to maintain accuracy and relevance.
+
+\hypertarget{adjustments-and-considerations}{%
+\subsubsection{Adjustments and
+Considerations}\label{adjustments-and-considerations}}
+
+Accounting for underreporting through the measurement of excess deaths
+provides a more comprehensive view of the pandemic's impact by capturing
+deaths that may have been missed or misclassified as COVID-19, while the
+process of age standardization adjusts these excess death rates to a
+standard age distribution, enabling meaningful comparisons between
+countries with varying demographic profiles. However, the analysis had
+to navigate challenges associated with varying data quality and
+availability across different countries, especially given the diverse
+COVID-19 reporting standards and testing capacities. Moreover, the study
+also acknowledges the significant influence of political and social
+factors on the pandemic's outcomes, which, despite being more
+challenging to quantify, are essential for a holistic understanding of
+the pandemic's impact.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+Our results are summarized in the following figures.
+Figure~\ref{fig-ELE} illustrates the trend of life expectancy at birth
+across different racial groups over time. Unfortunately, due to the
+unavailability of data over time for Asian, American Indian, and Alaska
+Native communities, they have been excluded from the time series graph.
+Table~\ref{tbl-life-exp} provides a more detailed breakdown of the life
+expectancy before and during COVID for Asian, American Indian, Alaska
+Native and other racial communities.
+
+An intriguing observation from Figure~\ref{fig-ELE} is the consistently
+higher life expectancy among Hispanic individuals compared to other
+groups, even amidst the challenges posed by COVID-19. On the other hand,
+Black individuals have consistently exhibited lower life expectancy,
+which further declined notably in 2021, reaching just over 70.8 years
+old. The life expectancy trends of white people and all other races and
+origins remain close together throughout the 14 years, with minimum
+variance.
+
+While all racial groups experienced declines in their average life
+expectancy, the declines vary greatly. The life expectancy for White
+individuals decreased from 78.8 to 76.4 years old from pre-pandemic
+levels in 2019 to 2021, while for Hispanic individuals, it dropped by
+4.2 years, and for Black individuals, it declined by 4 years during the
+same period. Our data aligns with the findings of Andrasfay and Goldman
+(2022), which states that the Black and Hispanic populations experienced
+twice as much life expectancy decrease as the white population.''
 
@@
 
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-GHS-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-ELE-1.pdf}
 
@@
 
-\caption{\label{fig-GHS}Global Health Security Index Scores by Country}
+\caption{\label{fig-ELE}Estimates of Life Expectancy at Birth, by Race
+2006-2021}
 
-\end{figure}%
-
-\section{Results}\label{results}
-
-Our results are summarized in the following figures. \textbf{?@fig-ELE}
-illustrates the trend of life expectancy at birth across different
-racial groups over time. Table~\ref{tbl-life_exp} provides a more
-detailed breakdown of the life expectancy before and during COVID
-for\ldots{} Unfortunately, due to the unavailability of data over time
-for Asians and American Indian and Alaska Native communities, they have
-been excluded from the time series graph.
-
-An intriguing observation is the consistently higher life expectancy
-among Hispanic individuals compared to other groups, even amidst the
-challenges posed by COVID-19. On the other hand, Black individuals l
-have consistently exhibited lower life expectancy, which further
-declined notably in 2021, reaching just over 70.8 years old. The life
-expectancy trends of white people and all other races and origins remain
-close together throughout the 14 year time period, with minimum
-variance. Specifically, the life expectancy for White individuals
-decreased from 78.8 to 76.4 years old from pre-pandemic levels in 2019
-to 2021, while for Hispanic individuals, it dropped by 4.2 years, and
-for Black individuals, it declined by 4 years during the same period.
-
-\begin{table}
-
-\caption{\label{tbl-life\_exp}Life Expectancy by Race (2019-2021)}
-
-\centering{
-
-}
-
-\end{table}%
+\end{figure}
 
@@
 
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-CLE-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-CLE-1.pdf}
 
@@
 
-\end{figure}%
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-pdf/fig-Vote-1.pdf}
+
+}
+
+\caption{\label{fig-Vote}Total Number of Votes in Top 10 States with
+Highest Proportion of Republican Votes}
+
+\end{figure}
+
+\hypertarget{tbl-deaths}{}
+\begin{longtable}{rlr}
+\caption{\label{tbl-deaths}Top 10 States with Highest Death Rates from COVID-19 (per 100,000
+people) }\tabularnewline
+
+\toprule
+Rank & State & Deaths \\ 
+\midrule\addlinespace[2.5pt]
+1 & Arizona & 455 \\ 
+2 & Oklahoma & 454 \\ 
+3 & Mississippi & 449 \\ 
+4 & West Virginia & 444 \\ 
+5 & New Mexico & 432 \\ 
+6 & Arkansas & 431 \\ 
+7 & Alabama & 429 \\ 
+8 & Tennessee & 428 \\ 
+9 & Michigan & 423 \\ 
+10 & Kentucky & 406 \\ 
+\bottomrule
+\end{longtable}
+
+\newpage
+
+\hypertarget{tbl-life-exp}{}
+\begin{longtable}{rrrrrrr}
+\caption{\label{tbl-life-exp}Life Expectancy by Race (2019-2021) }\tabularnewline
+
+\toprule
+Year & All races and origins & Hispanic & AIAN & Asian & Black & White \\ 
+\midrule\addlinespace[2.5pt]
+2019 & 78.8 & 81.9 & 71.8 & 85.6 & 74.8 & 78.8 \\ 
+2020 & 77.0 & 77.9 & 67.1 & 83.6 & 71.5 & 77.4 \\ 
+2021 & 76.1 & 77.7 & 65.2 & 83.5 & 70.8 & 76.4 \\ 
+\bottomrule
+\end{longtable}
 
 Figure~\ref{fig-CLE} serves as a valuable addition to our previous
-analysis with the inclusion of Asians and American Indians and Alaska
+analysis with the inclusion of Asians, American Indians, and Alaska
 Natives during the critical period from 2019 to 2021. With these
@@
 
-With a chart that depicts the change of life expectancy each year by
-ethnicity, we are able to better gain a more comprehensive understanding
-of the impact of COVID by ethnicity. While the declines in life
-expectancy are smaller in magnitude from 2020 to 2021, they are notably
-minimal for Asians and Hispanics, with reductions of -0.1 and -0.2 years
-respectively. (Insert some sources tmrw)
+With a chart that depicts the change in life expectancy each year by
+ethnicity, we can better gain a more comprehensive understanding of the
+impact of COVID by ethnicity. While the declines in life expectancy are
+smaller in magnitude from 2020 to 2021, they are notably minimal for
+Asians and Hispanics, with reductions of -0.1 and -0.2 years
+respectively.
 
-Numerous sources have indicated a correlation between a state's
-political affiliation and its handling of COVID issues (sources to be
-inserted tomorrow). To testify the claim, we collected the voting data
-for each of the 50 states for the 2020 presidential election. We then
-selected the ten states with the highest proportions of Republican
-votes.These voting patterns are visualized in Figure
-Figure~\ref{fig-Vote}, where red symbols represent Republican votes and
-blue symbols represent Democratic votes. Among the top ten states,
-Tennessee recorded the highest total number of votes, while Wyoming
-boasted the highest proportion of Republican votes.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-Vote-1.pdf}
-
-}
-
-\caption{\label{fig-Vote}Total Number of Votes in Top 10 States with
-Highest Proportion of Republican Votes}
-
-\end{figure}%
+Various studies have confirmed the existence of correlation between a
+state's political affiliation and its handling of COVID issues (Datz
+(2022)). To testify to the claim, we collected the voting data for each
+of the 50 states for the 2020 presidential election. We then selected
+the ten states with the highest proportions of Republican votes. These
+voting patterns are visualized in Figure Figure~\ref{fig-Vote}, where
+red symbols represent Republican votes and blue symbols represent
+Democratic votes. Among the top ten states, Tennessee recorded the
+highest total number of votes, while Wyoming boasted the highest
+proportion of Republican votes.
 
 Given the variation in population sizes among states, direct comparisons
-of COVID case numbers are inherently flawed. As an alternative, we
+of COVID-19 case numbers are inherently flawed. As an alternative, we
 employed death rates per 100,000 people as a metric for evaluating each
 state's COVID preparedness and situation. We then produced
-Table~\ref{tbl-deaths} that ranks the top ten states with the highest
-death rates from COVID per 100,000 people to examine the potential
-correlation between party preferences and COVID related deaths. Notably,
+Table~\ref{tbl-deaths} that rank the top ten states with the highest
+death rates from COVID-19 per 100,000 people to examine the potential
+correlation between party preferences and COVID-related deaths. Notably,
 We found that six of the ten top Republican states made a reappearance
 in the top death rates table; these states are Oklahoma, West Virginia,
-Arkansas, Alabama, Tennessee, and Kentucky. (Insert quotes about
-republican party and covid tmrw)
+Arkansas, Alabama, Tennessee, and Kentucky. The higher mortality rates
+observed in Republican-leaning states may be attributed to their
+preference for less stringent measures compared to Democratic-leaning
+states, which tend to favor stricter measures (VanDusky-Allen and
+Shvetsova (2021)).
 
-\begin{table}
-
-\caption{\label{tbl-deaths}Top 10 States with Highest Death Rates from
-COVID-19 (per 100,000 people)}
-
-\centering{
-
-}
-
-\end{table}%
-
-Following from our previous analysis regarding individuals' political
+Following our previous analysis regarding individuals' political
 affiliations, we have developed Figure
-Figure~\ref{fig-deaths_rates_votes}, which encompasses all 50 states of
+Figure~\ref{fig-deaths-rates-votes}, which encompasses all 50 states of
 the US along with their political leanings based on which party garnered
 the majority votes. This information is juxtaposed against their
-respective COVID death rates.
+respective COVID-19 death rates.
 
@@
 distributed, and notably one Democratic-leaning state has the lowest
-death rate.An intriguing observation is that although many
+death rate. An intriguing observation is that although many
 Republican-leaning states demonstrate higher COVID death rates, it is
@@
 
-(may add more)
-
 \begin{figure}
 
-\centering{
-
-\includegraphics{paper_files/figure-pdf/fig-deaths_rates_votes-1.pdf}
+{\centering \includegraphics{paper_files/figure-pdf/fig-deaths-rates-votes-1.pdf}
 
@@
 
-\caption{\label{fig-deaths\_rates\_votes}COVID-19 Death Rates
+\caption{\label{fig-deaths-rates-votes}COVID-19 Death Rates
 vs.~Republican Votes}
 
-\end{figure}%
+\end{figure}
 
-\newpage
-
-\section{Discussion}\label{discussion}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
 This begs the question as to why we are seeing these results. There
-isn't exactly a single answer to this question, however we can certainly
-point out some considerable factors to this result.
+isn't exactly a single answer to this question, however, we can
+certainly point out some considerable factors to this result.
 
+\hypertarget{influence-of-political-polarization-on-adherence-to-health-guidelines.}{%
 \subsection{Influence of political polarization on adherence to health
-guidelines.}\label{influence-of-political-polarization-on-adherence-to-health-guidelines.}
+guidelines.}\label{influence-of-political-polarization-on-adherence-to-health-guidelines.}}
 
@@
 
-Various studies and our own results have shown that areas with higher
+Various studies and our results have shown that areas with higher
 support for one political party exhibited distinct behaviors and
 compliance levels with health recommendations, which directly correlated
-with COVID-19 case rates and mortality. An news article from
+with COVID-19 case rates and mortality. A news article from
 \texttt{ABC\ News} (Diab and Kumar 2023) shows that the top states with
@@
 presidential voting data published by \texttt{CNN}, we have both states
-having the electoral vote of democrat with Washington wining by 58\%
+having the electoral vote of Democrat with Washington winning by 58\%
 (\emph{2020 Election Results by State, Washington} 2020) and Arizona
@@
 Another news article by \texttt{ContagionLive} (Parkinson 2023) also
-makes the claim of both Arizona and Washington having the highest
-COVID-19 mortality. This polarization has not only influenced individual
-behavior but also shaped state and local health policies, further
-entrenching the disparities in health outcomes.
+claims both Arizona and Washington have the highest COVID-19 mortality.
+This polarization has not only influenced individual behavior but also
+shaped state and local health policies, further entrenching the
+disparities in health outcomes.
 
-The adherence to health guidelines are evident in the varied health
+The adherence to health guidelines is evident in the varied health
 outcomes observed across the United States. Regions with lower
@@
 leaving certain communities more vulnerable to the virus and its
-variants. In order to mitigate the influence of political polarization
-on public health, it is imperative to depoliticize health guidelines and
-focus on evidence-based approaches to disease prevention and control.
-Building trust in health institutions and promoting bipartisan support
-for public health measures are essential steps towards achieving higher
-compliance and better health outcomes. Engaging trusted community
-leaders and utilizing targeted communication strategies can also help
-bridge the divide and encourage adherence to health guidelines.
+variants. To mitigate the influence of political polarization on public
+health, it is imperative to depoliticize health guidelines and focus on
+evidence-based approaches to disease prevention and control. Building
+trust in health institutions and promoting bipartisan support for public
+health measures are essential steps toward achieving higher compliance
+and better health outcomes. Engaging trusted community leaders and
+utilizing targeted communication strategies can also help bridge the
+divide and encourage adherence to health guidelines.
 
+\hypertarget{impact-of-government-transparency-and-consistent-communication-on-public-trust.}{%
 \subsection{Impact of government transparency and consistent
 communication on public
-trust.}\label{impact-of-government-transparency-and-consistent-communication-on-public-trust.}
+trust.}\label{impact-of-government-transparency-and-consistent-communication-on-public-trust.}}
 
@@
 
+\hypertarget{role-of-social-vulnerabilities-and-healthcare-access-disparities-in-pandemic-impact}{%
 \subsection{Role of social vulnerabilities and healthcare access
 disparities in pandemic
-impact.}\label{role-of-social-vulnerabilities-and-healthcare-access-disparities-in-pandemic-impact.}
+impact}\label{role-of-social-vulnerabilities-and-healthcare-access-disparities-in-pandemic-impact}}
 
+The COVID-19 pandemic starkly highlighted how social vulnerabilities and
+disparities in healthcare access exacerbated the impact of global health
+crises, contributing to significant variations in disease outcomes and
+underscoring the need for targeted public health strategies that address
+these disparities' root causes. Social vulnerabilities, such as
+socioeconomic status, race, ethnicity, and housing conditions,
+critically determined COVID-19 outcomes' severity, with populations in
+crowded housing, limited access to sanitation, and lower socioeconomic
+brackets experiencing higher transmission rates due to social distancing
+and hygiene maintenance challenges. An independent study done by the
+\texttt{Government\ of\ Canada} states that it ``identified that the
+risk of COVID-19 related deaths in Black, Asian and minority ethnic
+groups was nearly 1.5 times higher than White individuals'' (Emily
+Thompson 2021). Another news article done by the \texttt{MSN} (Dr.
+Sushama R. Chaphalkar 2024) states that `racial minority participants
+reported more negative impacts on health status, activity, and absence
+from work as compared to the White population.' The pandemic's economic
+toll further limited these groups' healthcare access, amplifying
+vulnerabilities. Disparities in healthcare access played a significant
+role in influencing COVID-19 morbidity and mortality, with communities
+facing healthcare facility shortages, provider scarcities, and barriers
+due to insurance or financial constraints at heightened risk. These
+disparities were evident in the uneven vaccine distribution and access,
+highlighting the advantages of regions with strong healthcare
+infrastructure. Marginalized populations, including racial and ethnic
+minorities, faced compounded risks from social vulnerabilities and
+healthcare disparities, evidenced by higher infection, hospitalization,
+and death rates due to factors like essential service employment and
+prevalent pre-existing conditions. A news article from \texttt{CNN}
+(Powell 2020) talks about how many essential workers, who cannot work
+from home, are from black and Latinx communities. These include
+healthcare professionals, grocery cashiers, delivery workers, and public
+transport employees. Despite their crucial roles, they often lack
+adequate pay, protection, and respect. Addressing these disparities in
+future pandemics requires public health strategies that prioritize
+equity and inclusivity, including community-based healthcare
+investments, enhanced vulnerable community outreach, and equitable
+healthcare resource access policies. By incorporating social
+determinants of health into public health preparedness plans, responses
+can effectively protect at-risk populations, making future public health
+responses more resilient, inclusive, and effective in safeguarding all
+population segments.
+
+\hypertarget{strategies-for-improving-real-time-data-collection-and-sharing-for-public-health-decisions.}{%
 \subsection{Strategies for improving real-time data collection and
 sharing for public health
-decisions.}\label{strategies-for-improving-real-time-data-collection-and-sharing-for-public-health-decisions.}
+decisions.}\label{strategies-for-improving-real-time-data-collection-and-sharing-for-public-health-decisions.}}
 
@@
 FHIR (Fast Healthcare Interoperability Resources) to facilitate
-efficient data sharing. Investing in digital surveillance systems, which
-employ AI and machine learning to sift through diverse data sources for
-early outbreak detection, is essential for rapid response to health
-threats. Furthermore, fostering public-private partnerships can harness
-the agility of the private sector and the public health expertise of
-governmental agencies to enhance data analytics capabilities. Ensuring
-the privacy and security of health data through robust governance
-frameworks and advanced encryption is paramount to maintaining public
-trust. Engaging communities in these initiatives ensures their relevance
-and fosters trust, while building global data sharing networks
-encourages international collaboration, crucial for a concerted response
-to pandemics. Collectively, these strategies are fundamental to
-bolstering public health decision-making and preparedness, making our
-health systems more resilient against the challenges posed by emerging
+efficient data sharing (\emph{What Is FHIR?} 2023). Investing in digital
+surveillance systems, which employ AI and machine learning to sift
+through diverse data sources for early outbreak detection, is essential
+for rapid response to health threats. Furthermore, fostering
+public-private partnerships can harness the agility of the private
+sector and the public health expertise of governmental agencies to
+enhance data analytics capabilities. Ensuring the privacy and security
+of health data through robust governance frameworks and advanced
+encryption is paramount to maintaining public trust. Engaging
+communities in these initiatives ensures their relevance and fosters
+trust while building global data-sharing networks encourages
+international collaboration, crucial for a concerted response to
+pandemics. Collectively, these strategies are fundamental to bolstering
+public health decision-making and preparedness, making our health
+systems more resilient against the challenges posed by emerging
 infectious diseases.
 
-\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}
+\hypertarget{weaknesses-and-next-steps}{%
+\subsection{Weaknesses and next steps}\label{weaknesses-and-next-steps}}
 
-Weaknesses and next steps should also be included.
+In our research, we identified several weaknesses that underscored the
+gap between the U.S.'s high pandemic preparedness ranking and its actual
+response to the COVID-19 crisis. Notably, there was a significant
+underutilization of preparedness capacities, as the resources and
+infrastructures in place were not fully mobilized or applied effectively
+during the pandemic. This disconnect highlights a critical area for
+improvement in aligning preparedness with real-time response
+capabilities. Furthermore, the response inadequately addressed intrinsic
+social vulnerabilities, exacerbating disparities in healthcare access
+and outcomes among different racial and socioeconomic groups, and
+underscoring the need for more inclusive and equitable public health
+strategies. The heavy politicization of the pandemic response, coupled
+with inconsistent messaging from health authorities, significantly
+eroded public trust and compliance, emphasizing the importance of
+depoliticizing public health measures and enhancing communication
+strategies. Additionally, shortcomings in data and surveillance systems,
+including delays in testing, inadequate genetic sequencing, and a lack
+of comprehensive surveillance, hindered informed decision-making and
+targeted interventions, pointing to a crucial need for investment in
+data infrastructure and capabilities. The analysis also suggests that
+the U.S. could benefit from more robust international benchmarking and
+learning from the diverse responses of other countries to health
+emergencies, which would involve understanding different strategies and
+their effectiveness beyond mere outcome comparisons.
+
+To address these challenges and strengthen the nation's resilience to
+future health emergencies, several next steps are recommended.
+Strengthening the linkages between preparedness and response is
+imperative, ensuring that capacities are not only available but also
+readily deployable and adaptable to the dynamics of a health crisis.
+This includes fostering agile and responsive systems capable of rapid
+mobilization. Addressing the social determinants of health is essential,
+with strategies aimed at mitigating the impact of social vulnerabilities
+through equitable healthcare access, support for marginalized
+communities, and targeted protective measures for vulnerable
+populations. Enhancing communication and public trust is vital,
+necessitating clear, consistent, and transparent communication from
+health authorities, alongside efforts to depoliticize health measures.
+Improving data and surveillance systems is critical to providing
+real-time insights, expanding genetic sequencing capabilities, and
+establishing standardized data protocols. Lastly, actively engaging in
+global health networks to exchange experiences, learn from global
+successes and failures, and collaborate on best practices is crucial for
+a more effective and equitable response that maximizes the full
+potential of preparedness capacities.
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+Despite the United States' top ranking in the Global Health Security
+Index as the most prepared nation for pandemics, its actual response to
+the COVID-19 pandemic fell short of expectations, leading to
+disproportionately high death rates. This discrepancy is attributed to
+several critical factors highlighted in the paper. The U.S. did not
+fully capitalize on its pandemic preparedness resources, with early
+testing blunders and a lack of a unified national testing approach
+hindering its response. Additionally, inherent vulnerabilities such as
+the significant portion of the population in congregated settings like
+nursing homes and prisons increased susceptibility to virus spread and
+severe health outcomes. The politicization of the pandemic response
+further compounded these issues, resulting in varied adherence to public
+health guidelines across states and political lines, thereby undermining
+the response's effectiveness. Moreover, inconsistent public health
+communications and the challenge of accessing standardized, quality data
+impeded the implementation of localized, effective interventions. The
+pandemic also highlighted and intensified existing socio-economic
+disparities, disproportionately impacting marginalized communities who
+faced higher risks and adverse outcomes, a situation that the response
+efforts failed to sufficiently mitigate.
 
@@
 
-\section*{References}\label{references}
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
 
-\phantomsection\label{refs}
+\hypertarget{refs}{}
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-citesource3}
+\leavevmode\vadjust pre{\hypertarget{ref-citesource3}{}}%
 \emph{2020 Election Results by State, Arizona}. 2020. CNN.
@@
 
-\bibitem[\citeproctext]{ref-citesource2}
+\leavevmode\vadjust pre{\hypertarget{ref-citesource2}{}}%
 \emph{2020 Election Results by State, Washington}. 2020. CNN.
@@
 
-\bibitem[\citeproctext]{ref-citeDataTable}
+\leavevmode\vadjust pre{\hypertarget{ref-andrasfay_goldman_2022}{}}%
+Andrasfay, Theresa, and Noreen Goldman. 2022. {``Reductions in US Life
+Expectancy from COVID-19 by Race and Ethnicity: Is 2021 a Repetition of
+2020?''} \emph{medRxiv: The Preprint Server for Health Sciences},
+January, 2021.10.17.21265117.
+https://doi.org/\url{https://doi.org/10.1101/2021.10.17.21265117}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data1}{}}%
+Arias, Elizabeth, Betzaida Tejada-Vera, Kenneth Kochanek, and Farida
+Ahmad. 2022. {``Provisional Life Expectancy Estimates for 2021.''}
+\emph{Atlanta, Georgia}, August.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data2}{}}%
+Arias, Elizabeth, and Jiaquan Xu. 2022. {``United States Life Tables,
+2020.''} \emph{Atlanta, Georgia}, August.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citeDataTable}{}}%
 Barrett, Tyson, Matt Dowle, Arun Srinivasan, Jan Gorecki, Michael
@@
 
-\bibitem[\citeproctext]{ref-citewebshot}
+\leavevmode\vadjust pre{\hypertarget{ref-data3}{}}%
+Bell, Julie A, and Jennifer B Nuzzo. 2021. {``Global Health Security
+Index: Advancing Collective Action and Accountability Amid Global
+Crisis.''} Washington D.C.:
+\url{https://www.ghsindex.org/report-model/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citewebshot}{}}%
 Chang, Winston. 2023a. \emph{Webshot: Take Screenshots of Web Pages}.
@@
 
-\bibitem[\citeproctext]{ref-citewebshot2}
+\leavevmode\vadjust pre{\hypertarget{ref-citewebshot2}{}}%
 ---------. 2023b. \emph{Webshot2: Take Screenshots of Web Pages}.
@@
 
-\bibitem[\citeproctext]{ref-citesource1}
+\leavevmode\vadjust pre{\hypertarget{ref-data6}{}}%
+\emph{COVID-19 Excess Mortality Estimates 2020-2021}. 2022. Seattle,
+United States of America: Institute for Health Metrics; Evaluation
+(IHME).
+
+\leavevmode\vadjust pre{\hypertarget{ref-datz_2022}{}}%
+Datz, Todd. 2022. {``Political Ideology of u.s. Elected Officials Linked
+with COVID-19 Health Outcomes.''} \emph{News}.
+\url{https://www.hsph.harvard.edu/news/press-releases/political-ideology-of-u-s-elected-officials-linked-with-covid-19-health-outcomes/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data9}{}}%
+{``Death Rate, Crude (Per 1,000 People) (SP.DYN.CDRT.IN).''} 2022. The
+World Bank;
+\url{https://databank.worldbank.org/reports.aspx?source=2\&series=SP.DYN.CDRT.IN\&country=\#}.
+2022.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citesource1}{}}%
 Diab, Dr. Alaa, and Dr. Keerthana Kumar. 2023. \emph{COVID-19 Death
@@
 
-\bibitem[\citeproctext]{ref-citedata2}
+\leavevmode\vadjust pre{\hypertarget{ref-data5}{}}%
+Dong, Ensheng, Hongru Du, and Lauren Gardner. 2020. {``An Interactive
+Web-Based Dashboard to Track COVID-19 in Real Time.''} \emph{The Lancet
+Infectious Diseases} 20 (5): 533--34.
+\url{https://doi.org/10.1016/S1473-3099(20)30120-1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citesource6}{}}%
+Dr. Sushama R. Chaphalkar, PhD. 2024. \emph{COVID-19 Recovery
+Disparities Uncovered Among Racial and Ethnic Groups}. MSN.
+\url{https://www.msn.com/en-gb/health/other/covid-19-recovery-disparities-uncovered-among-racial-and-ethnic-groups/ar-BB1hLTLw}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data10}{}}%
+Economic, United Nations Department of, and Social Affairs Population
+Division. 2022. {``World Population Prospects 2022 Demographic
+Indicators by Region, Subregion and Country, Annually for 1950-2100.''}
+\url{https://population.un.org/wpp/Download/Standard/MostUsed/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citedata2}{}}%
 Elflein, John. 2023. {``U.s. COVID-19 Death Rate by State.''}
@@
 
-\bibitem[\citeproctext]{ref-citeJanitor}
+\leavevmode\vadjust pre{\hypertarget{ref-citesource5}{}}%
+Emily Thompson, Nicole Atchessi, Rojiemiahd Edjoc. 2021. \emph{COVID-19
+Race Data Collection in Canada}. The Public Health Agency of Canada.
+\url{https://www.canada.ca/en/public-health/services/reports-publications/canada-communicable-disease-report-ccdr/monthly-issue/2021-47/issue-7-8-july-august-2021/covid-19-race-data-collection-canada.html}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citeJanitor}{}}%
 Firke, Sam. 2023. \emph{Janitor: Simple Tools for Examining and Cleaning
@@
 
-\bibitem[\citeproctext]{ref-citeLubridate}
+\leavevmode\vadjust pre{\hypertarget{ref-citeLubridate}{}}%
 Grolemund, Garrett, and Hadley Wickham. 2011. {``Dates and Times Made
@@
 
-\bibitem[\citeproctext]{ref-citeGT}
+\leavevmode\vadjust pre{\hypertarget{ref-data8}{}}%
+Health Statistics, National Center for. 2022. {``Life Expectancy.''} US
+Centers for Disease Control; Prevention;
+\url{https://www.cdc.gov/nchs/nvss/life-expectancy.htm}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citegt}{}}%
 Iannone, Richard, Joe Cheng, Barret Schloerke, Ellis Hughes, Alexandra
@@
 
-\bibitem[\citeproctext]{ref-citedata1}
+\leavevmode\vadjust pre{\hypertarget{ref-citedata1}{}}%
 Jack, Rebecca, and Emily Oster. 2023. {``COVID-19, School Closures, and
@@
 
-\bibitem[\citeproctext]{ref-citeggpubr}
+\leavevmode\vadjust pre{\hypertarget{ref-citeggpubr}{}}%
 Kassambara, Alboukadel. 2023. \emph{Ggpubr: 'Ggplot2' Based Publication
@@
 
-\bibitem[\citeproctext]{ref-citeHere}
+\leavevmode\vadjust pre{\hypertarget{ref-data7}{}}%
+Ledesma, J R, C R Isaac, S F Dowell, D L Blazes, G V Essix, K Budeski,
+Julie Bell, and Jennifer B Nuzzo. 2023. {``Evaluation of the Global
+Health Security Index as a Predictor of COVID-19 Excess Mortality
+Standardised for Under-Reporting and Age Structure.''} \emph{BMJ Global
+Health} 8 (7): e012203. \url{https://doi.org/10.1136/bmjgh-2023-012203}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data12}{}}%
+McGovern, Tony. 2009-\/-2020. {``United States General Election
+Presidential Results by County from 2009 to 2020.''} Github Repository.
+\url{https://github.com/tonmcg/US_County_Level_Election_Results_08-20}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citeHere}{}}%
 Müller, Kirill. 2020. \emph{Here: A Simpler Way to Find Your Files}.
@@
 
-\bibitem[\citeproctext]{ref-citeRColorBrewer}
+\leavevmode\vadjust pre{\hypertarget{ref-citeRColorBrewer}{}}%
 Neuwirth, Erich. 2022. \emph{RColorBrewer: ColorBrewer Palettes}.
@@
 
-\bibitem[\citeproctext]{ref-citearticle}
+\leavevmode\vadjust pre{\hypertarget{ref-citearticle}{}}%
 Nuzzo, Jennifer B., and Jorge R. Ledesma. 2023. \emph{Why Did the Best
@@
 
-\bibitem[\citeproctext]{ref-citesource4}
+\leavevmode\vadjust pre{\hypertarget{ref-citesource4}{}}%
 Parkinson, John. 2023. \emph{Which States Saw Greatest COVID-19
@@
 
-\bibitem[\citeproctext]{ref-citeSF}
+\leavevmode\vadjust pre{\hypertarget{ref-citeSF}{}}%
 Pebesma, Edzer. 2018. {``{Simple Features for R: Standardized Support
@@
 
-\bibitem[\citeproctext]{ref-citeR}
+\leavevmode\vadjust pre{\hypertarget{ref-citesource7}{}}%
+Powell, Catherine. 2020. \emph{Color of Covid: The Racial Justice
+Paradox of Our New Stay-at-Home Economy}. CNN.
+\url{https://www.cnn.com/2020/04/10/opinions/covid-19-people-of-color-labor-market-disparities-powell/index.html}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citeR}{}}%
 R Core Team. 2022. \emph{R: A Language and Environment for Statistical
@@
 
-\bibitem[\citeproctext]{ref-citeggplot}
+\leavevmode\vadjust pre{\hypertarget{ref-data4}{}}%
+Systems Science, Center for, and Engineering. 2023. {``COVID-19
+Dashboard.''} John Hopkins University.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vandusky-allen_shvetsova_2021}{}}%
+VanDusky-Allen, Julie, and Olga Shvetsova. 2021. {``How America's
+Partisan Divide over Pandemic Responses Played Out in the States.''}
+\emph{The Conversation}.
+\url{https://theconversation.com/how-americas-partisan-divide-over-pandemic-responses-played-out-in-the-states-157565}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-data11}{}}%
+Wang, Haidong, Kara R Paulson, Samuel A Pease, Samuel Watson, Haley
+Comfort, Pu Zheng, A Y Aravkin, et al. 2022. {``Estimating Excess
+Mortality Due to the COVID-19 Pandemic: A Systematic Analysis of
+COVID-19-Related Mortality, 2020--21.''} \emph{The Lancet} 399 (10334):
+1513--36. \url{https://doi.org/10.1016/S0140-6736(21)02796-3}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citesource8}{}}%
+\emph{What Is FHIR?} 2023. The Office of National Coordinator for Health
+Information Technology.
+\url{https://www.healthit.gov/sites/default/files/2019-08/ONCFHIRFSWhatIsFHIR.pdf}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-citeggplot}{}}%
 Wickham, Hadley. 2016. \emph{Ggplot2: Elegant Graphics for Data
@@
 
-\bibitem[\citeproctext]{ref-citeTidyVerse}
+\leavevmode\vadjust pre{\hypertarget{ref-citeTidyVerse}{}}%
 Wickham, Hadley, Mara Averick, Jennifer Bryan, Winston Chang, Lucy
@@
 
-\bibitem[\citeproctext]{ref-citeReadXL}
+\leavevmode\vadjust pre{\hypertarget{ref-citeReadXL}{}}%
 Wickham, Hadley, and Jennifer Bryan. 2023. \emph{Readxl: Read Excel
@@
 
-\bibitem[\citeproctext]{ref-citeDplyr}
+\leavevmode\vadjust pre{\hypertarget{ref-citeDplyr}{}}%
 Wickham, Hadley, Romain François, Lionel Henry, Kirill Müller, and Davis
@@
 
-\bibitem[\citeproctext]{ref-citeScales}
+\leavevmode\vadjust pre{\hypertarget{ref-citeScales}{}}%
 Wickham, Hadley, Thomas Lin Pedersen, and Dana Seidel. 2023.
@@
 
-\bibitem[\citeproctext]{ref-citeKnitR}
+\leavevmode\vadjust pre{\hypertarget{ref-citeKnitR}{}}%
 Xie, Yihui. 2014. {``Knitr: A Comprehensive Tool for Reproducible
@@
 
-\bibitem[\citeproctext]{ref-citeKableExtra}
+\leavevmode\vadjust pre{\hypertarget{ref-citeKableExtra}{}}%
 Zhu, Hao. 2024. \emph{kableExtra: Construct Complex Table with 'Kable'

### 2024-02-13 outputs/paper/paper.qmd
rerendered paper

@@
 
-ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
-world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
-life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
-vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
-deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
-
-# ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-# world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-# life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-# vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
-# deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
 
@@
 
-\newpage
-
 ```{r}

### 2024-02-13 outputs/paper/paper.qmd
data measurements and others

--- 2024-02-13 outputs/paper/paper.qmd
@@
 ## Data Measurement
+
+The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for life expectancy rates, and political affiliation.
+
+Our measurement for COVID death rates is drawn from a accumulated dataset last updated until March 2023.
+This temporal constraint means we would not be aware of the fluctuations of death rates over time, particularly during the peak of the pandemic in 2020.
+The data's primarily reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduces bias as excludes uncounted deaths from COVID.
+Especially during the critical period of health facilities capacity constraints,many individuals died outside of medical facilities without being documented.
+Therefore, our measurements for COVID death rates are only accounted for the documented death from COVID.
+
+In addition,it's important to note that the death rate presented here reflects the entire population of a state.
+However, COVID-19 disproportionately affects older individuals compared to younger adults in the prime of their lives.
+As a result, this data does not capture the effect of COVID across different age groups.
+Instead, it only offers a general overview the state's population.
+In summary, we are unable to provide detailed insights into age-specific or temporal variations based on this dataset.
+
+Similarly, our measurement of life expectancy suffer from the demographic limitations.
+Even though we categorize life expectancy by race, the absence of an age breakdown restricts us discern impacts across different age groups.
+Moreover, delays in reporting deaths undermine the timeliness and accuracy of our data, potentially skewing our understanding of population health trends.
+
+Our measurement for each state's political preference is calculated from the 2020 election votes.
+There may be many scenarios that prevented eligible voters from voting.
+However, we are only able to assess the the state's political preference based on the people who voted, thereby potentially introducing selection bias.
+Additionally, issues such as ballot loss or voter disenfranchisement further compound the reliability of our measurements.
 

### 2024-02-13 outputs/paper/paper.qmd
added intro and abstract and updated llm usage and readme

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In this research, we explored the complex interplay between local health infrastructure, socio-political factors, and COVID-19 outcomes across various U.S.
+states and counties to understand disparities in the pandemic's impact.
+Our findings reveal that political polarization and healthcare access disparities significantly influenced adherence to health guidelines and the effectiveness of the public health response.
+Additionally, social vulnerabilities and inconsistent public health messaging further exacerbated the pandemic's impact on marginalized communities.
+This study underscores the critical need for targeted public health strategies that address socio-political and infrastructural disparities to enhance pandemic preparedness and response, emphasizing the importance of equity and effective communication in safeguarding public health."
 format: pdf
@@
 
-@fig-GHS allows us to visualize the Global Health Security (GHS) Index which is designed to assess a country's capabilities to prevent, detect, and respond to significant infectious disease outbreaks.
+The COVID-19 pandemic has cast a stark light on the complexities and challenges of managing a global health crisis, particularly in the United States, which, despite being considered one of the best-prepared nations for a pandemic, experienced significant disparities in outcomes across its diverse states and counties.
+These disparities highlight the crucial interplay between local health infrastructure, socio-political factors, and the effectiveness of public health responses, underscoring the need for a nuanced understanding of how these elements influence pandemic outcomes.
+
+[@fig-GHS] gives us some context to motivate our research.
+This map allows us to visualize the Global Health Security (GHS) Index which is designed to assess a country's capabilities to prevent, detect, and respond to significant infectious disease outbreaks.
 A higher Global Health Security (GHS) Index score indicates a stronger health system capacity for pandemic readiness.
@@
 Lower scores indicate deficiencies in one or more of the critical areas assessed by the Index, which could potentially hinder the country's ability to effectively manage and control the spread of infectious diseases.
+
+Later on, this paper delves into the intricate dynamics between local health systems, socio-political influences, and their collective impact on the disparities observed in COVID-19 outcomes across the United States.
+By analyzing data from various states and counties, we aim to identify how differences in health infrastructure, political affiliations, and social vulnerabilities contributed to the varied effectiveness of the response to the pandemic.
+This investigation seeks to fill the gap in understanding the role of these factors in shaping public health outcomes during a pandemic, a critical area of inquiry given the unprecedented scale and impact of COVID-19.
+
+Our research uncovered that political polarization significantly affected adherence to health guidelines, with stark differences in compliance and outcomes between regions with varying political leanings.
+Furthermore, disparities in healthcare access and the prevalence of social vulnerabilities were closely correlated with the severity of the pandemic's impact, highlighting the exacerbating effect of pre-existing inequalities.
+These findings underscore the necessity of developing public health strategies that are sensitive to socio-political dynamics and infrastructural disparities, to enhance resilience against future health crises.
+
+This paper is structured to first provide a comprehensive overview of the broader context and motivations behind this study, followed by a detailed examination of the methodologies employed to investigate the research question.
+The results section presents the key findings, which are then discussed in the context of their implications for public health preparedness and response strategies.
+The conclusion summarizes the study's contributions to our understanding of the complex factors influencing pandemic outcomes and outlines recommendations for future research and policy development.
 

### 2024-02-13 outputs/paper/paper.qmd
changes

@@
   - Adrian Ly
-  - Sakhil Goel
   - Hannah Yu

### 2024-02-13 outputs/paper/paper.qmd
changes

@@
 
-ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
-world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
-life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
-vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
-deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+# ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
+# world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))
+# life_exp <- read.csv(here::here("inputs/data/life_table.csv"), stringsAsFactors = FALSE)
+# vote_share <- read.csv(here::here("inputs/data/Vote_share_2020_data.csv"), stringsAsFactors = FALSE)
+# deaths <- read.csv(here::here("inputs/data/covid-19-death-rates.csv"), stringsAsFactors = FALSE)
+ghs_data <- read_excel("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/GHS_index.xlsx")
+world_map <- st_read("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/world/World_Countries__Generalized_.shp")
+life_exp <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/life_table.csv", stringsAsFactors = FALSE)
+vote_share <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/Vote_share_2020_data.csv", stringsAsFactors = FALSE)
+deaths <- read.csv("C:/Users/35911/Documents/US-Covid-Analysis/inputs/data/covid-19-death-rates.csv", stringsAsFactors = FALSE)
 

### 2024-02-13 outputs/paper/paper.qmd
added some intro

--- 2024-02-13 outputs/paper/paper.qmd
@@
 
-The COVID-19 pandemic has cast a stark light on the complexities and challenges of managing a global health crisis, particularly in the United States, which, despite being considered one of the best-prepared nations for a pandemic, experienced significant disparities in outcomes across its diverse states and counties.
-These disparities highlight the crucial interplay between local health infrastructure, socio-political factors, and the effectiveness of public health responses, underscoring the need for a nuanced understanding of how these elements influence pandemic outcomes.
+The COVID-19 pandemic has exposed significant disparities in health outcomes across various regions of the United States, drawing attention to the intricate interplay of local factors in shaping disease trajectories.
+Understanding these dynamics is crucial for informing effective public health interventions and fostering resilience in the face of future pandemics.
+In this paper, building upon prior research from the American Economic Association, we delve into the multifaceted relationship between local health infrastructure, socio-political factors, and COVID-19 outcomes, aiming to provide insights into the mechanisms driving disparities and identify strategies for targeted intervention.
+
+Our analysis is grounded in comprehensive data sourced from the American Economic Association, enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
+While leveraging the initial findings from the AEA paper, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities.
+We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access, housing conditions, and ethnic demographics in shaping disparities in COVID-19 outcomes.
+One key finding that emerges from our study is the influence of political polarization, with Republican-leaning states exhibiting slightly higher clustering around elevated death rates, highlighting the importance of socio-political factors in shaping public health responses.
 
@@
 The conclusion summarizes the study's contributions to our understanding of the complex factors influencing pandemic outcomes and outlines recommendations for future research and policy development.
+
+As we navigate the complexities of the COVID-19 pandemic, it is imperative to not only understand the factors contributing to disparities but also to identify actionable strategies for mitigating these inequities and strengthening community resilience.
+In the subsequent sections of this paper, we first delve deeper into our methodology, data cleaning, and data measurement.
+Next, we present our findings in detail with graphs and tables.
+Finally, we discuss how these results came about and the implications of our research for public health policy and practice.
+By elucidating these critical relationships and expanding upon prior research, our study contributes to the broader discourse on pandemic preparedness and underscores the importance of tailored interventions in addressing underlying structural determinants of health disparities.
 

### 2024-02-13 outputs/paper/paper.qmd
Intro, abstract, and some edits

@@
 ---
-title: "How did local health infrastructure and socio-political factors within different states and counties in the United States affect the disparities in COVID-19 outcomes, and what lessons can be learned for more targeted public health preparedness and response strategies in future pandemics?"
+title: "Political Polarization, Public Trust, and Healthcare Disparities: Unraveling COVID-19's Impact and Future Preparedness"
 author: 
@@
   - Hannah Yu
+  - Sakhil Goel
 thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git"
@@
 date-format: long
-abstract: "In this research, we explored the complex interplay between local health infrastructure, socio-political factors, and COVID-19 outcomes across various U.S.
-states and counties to understand disparities in the pandemic's impact.
-Our findings reveal that political polarization and healthcare access disparities significantly influenced adherence to health guidelines and the effectiveness of the public health response.
-Additionally, social vulnerabilities and inconsistent public health messaging further exacerbated the pandemic's impact on marginalized communities.
-This study underscores the critical need for targeted public health strategies that address socio-political and infrastructural disparities to enhance pandemic preparedness and response, emphasizing the importance of equity and effective communication in safeguarding public health."
+abstract: "The COVID-19 pandemic has exposed deep disparities in health outcomes across different states and counties in the United States.
+We use data from the American Economic Association to conduct a comprehensive analysis of how variances in local health infrastructure and socio-political factors within the United States contributed to disparities in COVID-19 outcomes.
+Our analysis reveals that political polarization significantly influenced adherence to health guidelines, with Republican-leaning states demonstrating slightly higher clustering around elevated death rates.
+Moreover, social vulnerabilities, including disparities in healthcare access, housing conditions, and ethnic demographics, emerged as key contributors to substantial variations in disease outcomes.
+These results highlight the urgent requirement for customized public health readiness and response tactics, focusing on precise interventions to tackle deficiencies in community health infrastructure and socio-political factors."
 format: pdf
@@
 Lower scores indicate deficiencies in one or more of the critical areas assessed by the Index, which could potentially hinder the country's ability to effectively manage and control the spread of infectious diseases.
-
-Later on, this paper delves into the intricate dynamics between local health systems, socio-political influences, and their collective impact on the disparities observed in COVID-19 outcomes across the United States.
-By analyzing data from various states and counties, we aim to identify how differences in health infrastructure, political affiliations, and social vulnerabilities contributed to the varied effectiveness of the response to the pandemic.
-This investigation seeks to fill the gap in understanding the role of these factors in shaping public health outcomes during a pandemic, a critical area of inquiry given the unprecedented scale and impact of COVID-19.
-
-Our research uncovered that political polarization significantly affected adherence to health guidelines, with stark differences in compliance and outcomes between regions with varying political leanings.
-Furthermore, disparities in healthcare access and the prevalence of social vulnerabilities were closely correlated with the severity of the pandemic's impact, highlighting the exacerbating effect of pre-existing inequalities.
-These findings underscore the necessity of developing public health strategies that are sensitive to socio-political dynamics and infrastructural disparities, to enhance resilience against future health crises.
-
-This paper is structured to first provide a comprehensive overview of the broader context and motivations behind this study, followed by a detailed examination of the methodologies employed to investigate the research question.
-The results section presents the key findings, which are then discussed in the context of their implications for public health preparedness and response strategies.
-The conclusion summarizes the study's contributions to our understanding of the complex factors influencing pandemic outcomes and outlines recommendations for future research and policy development.
 
@@
 We then selected the ten states with the highest proportions of Republican votes.
-These voting patterns are visualized in Figure @fig-Vote, where red symbols represent Republican votes and blue symbols represent Democratic votes.
+These voting patterns are visualized in @fig-Vote, where red symbols represent Republican votes and blue symbols represent Democratic votes.
 Among the top ten states, Tennessee recorded the highest total number of votes, while Wyoming boasted the highest proportion of Republican votes.
@@
 
-Following our previous analysis regarding individuals' political affiliations, we have developed Figure @fig-deaths-rates-votes, which encompasses all 50 states of the US along with their political leanings based on which party garnered the majority votes.
+Following our previous analysis regarding individuals' political affiliations, we have developed @fig-deaths-rates-votes, which encompasses all 50 states of the US along with their political leanings based on which party garnered the majority votes.
 This information is juxtaposed against their respective COVID-19 death rates.

### 2024-02-13 scripts/01-simulate_data.R

@@ -2,7 +2,7 @@
 # Purpose: Simulate Data of Interest
 # Author: Adrian Ly, Hannah Yu
 # Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca
+# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca, s.goel@mail.utoronto.ca
 # License: MIT
 
 #### Workspace setup ####

### 2024-02-13 scripts/02-test_data.R

@@ -2,7 +2,7 @@
 # Purpose: Tests the Simulated Data
 # Author: Adrian Ly
 # Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca
+# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca, s.goel@mail.utoronto.ca
 # License: MIT
 
 #### Workspace setup ####

### 2024-02-13 scripts/Replication.R

@@ -2,7 +2,7 @@
 # Purpose: Paper Replication
 # Author: Adrian Ly, Hannah Yu
 # Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca
+# Contact: adrian.ly@mail.utoronto.ca, s.goel@mail.utoronto.ca
 # License: MIT
 
 

### 2024-02-13 scripts/01-simulate_data.R
fixed libraries

@@ -7,18 +7,9 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(sf)
-library(readxl)
 library(knitr)
-library(janitor)
-library(lubridate)
-library(dplyr)
-library(data.table)
-library(RColorBrewer)
-library(ggpubr)
-library(scales)
-library(here)
-library(kableExtra)
+
+
 
 #### Simulate data ####
 

### 2024-02-13 scripts/02-test_data.R

@@ -1,24 +1,14 @@
 #### Preamble ####
 # Purpose: Tests the Simulated Data
-# Author: Adrian Ly
+# Author: Adrian Ly, Hannah Yu
 # Date: 9 February 2024 
 # Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca, s.goel@mail.utoronto.ca
 # License: MIT
 
 #### Workspace setup ####
 library(tidyverse)
-library(sf)
-library(readxl)
 library(knitr)
-library(janitor)
-library(lubridate)
-library(dplyr)
-library(data.table)
-library(RColorBrewer)
-library(ggpubr)
-library(scales)
-library(here)
-library(kableExtra)
+
 
 #### Test data ####
 ## Test simulated_life_data

### 2024-02-13 scripts/03-replication.R

@@ -10,7 +10,6 @@
 library(tidyverse)
 library(sf)
 library(readxl)
-library(knitr)
 library(janitor)
 library(lubridate)
 library(dplyr)
@@ -21,8 +20,6 @@ library(scales)
 library(here)
 library(kableExtra)
 library(webshot)
-library(webshot2)
-
 #### Reading files ####
 ghs_data <- read_excel(here::here("inputs/data/GHS_index.xlsx"))
 world_map <- st_read(here::here("inputs/data/world/World_Countries__Generalized_.shp"))

### 2024-02-14 outputs/paper/paper.qmd
updated a few peer reviews

@@
   - Sakhil Goel
-thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git"
+thanks: "Code and data are available at: https://github.com/hannahyu07/US-Covid-Analysis.git 
+This reproduction was performed after a replication on the Social Science Reproduction platform: [link here](https://www.socialsciencereproduction.org/reproductions/c35e8e98-762d-4c00-a1a3-544ed0b2008d/index)"
 date: today
@@
 # Introduction
-
-*This reproduction was performed after a replication on the Social Science Reproduction platform: ****[link here](https://www.socialsciencereproduction.org/reproductions/c35e8e98-762d-4c00-a1a3-544ed0b2008d/index)***
 
@@
 
-The datasets utilized in this paper were mainly obtained from the `original paper` [@citearticle].
-The data used by the paper was compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11.
+The datasets utilized in the `original paper` [@citearticle] and this paper were mainly compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11.
 Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and [@citedata2].
@@
 
-The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for life expectancy rates, and political affiliation.
+The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for life expectancy rates [@citedefenition], and political affiliation.
 

### 2024-02-14 outputs/paper/paper.qmd
updated discussion

@@
 
-This begs the question as to why we are seeing these results. There isn't exactly a single answer to this question, however, we can certainly point out some considerable factors to this result.
+This begs the question as to why we are seeing these results.
+There isn't exactly a single answer to this question, however, we can certainly point out some considerable factors to this result.
+These factors can be broken down into the influence of political polarization, the impact of government transparency which ties into the effect it also had on public trust, and the role of social vulnerabilities and healthcare access disparities.
+We should also look at how to address these potential factors by implementing strategies in the future in regards to data collection and sharing.
+It is important to also note we were able to identify some weaknesses in our research that should be evaluated in the future.
 
-## Influence of political polarization on adherence to health guidelines.
+## Influence of political polarization on adherence to health guidelines
 

### 2024-02-14 outputs/paper/paper.qmd
Updated abstract

@@
 date-format: long
-abstract: "The COVID-19 pandemic has exposed deep disparities in health outcomes across different states and counties in the United States.
-We use data from the American Economic Association to conduct a comprehensive analysis of how variances in local health infrastructure and socio-political factors within the United States contributed to disparities in COVID-19 outcomes.
-Our analysis reveals that political polarization significantly influenced adherence to health guidelines, with Republican-leaning states demonstrating slightly higher clustering around elevated death rates.
-Moreover, social vulnerabilities, including disparities in healthcare access, housing conditions, and ethnic demographics, emerged as key contributors to substantial variations in disease outcomes.
-These results highlight the urgent requirement for customized public health readiness and response tactics, focusing on precise interventions to tackle deficiencies in community health infrastructure and socio-political factors."
+abstract: "The COVID-19 pandemic has exposed deep disparities in health outcomes across various regions in the United States.
+Using research conducted by @citearticle, we analyze how variances in local health infrastructure and socio-political factors influenced COVID-19 outcomes.
+Our analysis reveals that political polarization notably impacted adherence to health guidelines, particularly in Republican-leaning states, correlating with higher death rates.
+Furthermore, disparities in healthcare access and housing significantly affected the life expectancy of different ethnic groups.
+These findings emphasize the need for tailored public health strategies to address infrastructure and socio-political deficiencies, while future research should continue to explore these factors for better pandemic preparedness and response."
 format: pdf
@@
 
-Our analysis is grounded in comprehensive data sourced from the American Economic Association, enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
-While leveraging the initial findings from the AEA paper, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities.
+Our analysis is grounded in comprehensive data sourced from the paper, `Why Did the Best Prepared Country in
+the World Fare So Poorly during COVID?` [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
+While leveraging the initial findings from the @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities.
 We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access, housing conditions, and ethnic demographics in shaping disparities in COVID-19 outcomes.

### 2024-02-14 outputs/paper/paper.qmd
updated discussion and references

@@
 Equally important is enhancing data standardization and interoperability through universal standards like FHIR (Fast Healthcare Interoperability Resources) to facilitate efficient data sharing [@citesource8].
+Another notable example is the Digital Health Platform (DHP) [@citedhp] developed by the European Centre for Disease Prevention and Control (ECDC) [@citeecdc], which facilitates the real-time exchange and analysis of health data across EU member states.
+This platform has been instrumental in coordinating responses to health threats, including the COVID-19 pandemic, by providing timely and accurate data to inform public health decisions.
+Another digital platform worth mentioning is the HealthMap system [@citehealthmap], which utilizes online data sources and machine learning algorithms to detect and track global disease outbreaks.
+HealthMap's ability to aggregate and visualize data from various sources, including news reports and social media, has proven invaluable in early outbreak detection and response, showcasing the power of digital surveillance in public health.
 Investing in digital surveillance systems, which employ AI and machine learning to sift through diverse data sources for early outbreak detection, is essential for rapid response to health threats.
+An news article by `nature` [@citegoogle] talks about how Google has made an AI which claims to be "better bedside manner than human doctors — and makes better diagnoses".
 Furthermore, fostering public-private partnerships can harness the agility of the private sector and the public health expertise of governmental agencies to enhance data analytics capabilities.

### 2024-02-14 scripts/01-simulate_data.R
Updated simulation

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Simulate Data of Interest
-# Author: Adrian Ly, Hannah Yu
+# Author: Adrian Ly, Hannah Yu, Sakhil Goel
 # Date: 9 February 2024 
 # Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca, s.goel@mail.utoronto.ca
 # License: MIT
@@ -10,7 +10,6 @@ library(tidyverse)
 library(knitr)
 
 
-
 #### Simulate data ####
 
 # Simulate the average life expectancy of people by ethnicity
@@ -18,7 +17,7 @@ library(knitr)
 set.seed(123)
 
 # Define the years
-years <- 2006:2021
+years <- seq(2006, 2021)
 
 # Define life expectancy ranges for each group
 life_expectancy_ranges <- list(
@@ -30,27 +29,27 @@ life_expectancy_ranges <- list(
   "Non-Hispanic White" = c(77, 81)
 )
 
-# Generate the dataset
-data <- matrix(NA, nrow = length(years), ncol = length(life_expectancy_ranges) + 1)
-colnames(data) <- c("year_id", names(life_expectancy_ranges))
-data[, 1] <- years
-
-for (i in 1:length(years)) {
-  for (group in names(life_expectancy_ranges)) {
-    if (group == "All") {
-      life_expectancy <- runif(1, life_expectancy_ranges[[group]][1], life_expectancy_ranges[[group]][2])
-    } else {
-      life_expectancy <- runif(1, life_expectancy_ranges[[group]][1], life_expectancy_ranges[[group]][2])
-    }
-    data[i, which(colnames(data) == group)] <- round(life_expectancy, 1)
-  }
+# Define function to generate random numbers within a range
+generate_random_number <- function(range, n) {
+  return(round(runif(n, min = range[1], max = range[2]), 1))
 }
 
-# Save the dataset to a CSV file
-write.csv(data, "outputs/data/simulated_life_exp.csv", row.names = FALSE)
+# Generate life expectancy data
+simulated_life_exp <- tibble(
+  year = years,
+  All = generate_random_number(c(77, 80), n = length(years)),
+  Hispanic = generate_random_number(c(80, 83), n = length(years)),
+  Non_Hispanic_AIAN = generate_random_number(c(65, 72), n = length(years)),
[…]

### 2024-02-15 scripts/01-simulate_data.R
Updated simulation

@@ -34,7 +34,7 @@ generate_random_number <- function(range, n) {
   return(round(runif(n, min = range[1], max = range[2]), 1))
 }
 
-# Generate life expectancy data
+# Generate simulated life expectancy data
 simulated_life_exp <- tibble(
   year = years,
   All = generate_random_number(c(77, 80), n = length(years)),
@@ -50,10 +50,13 @@ print(head(simulated_life_exp))
 # Save the dataset to a CSV file
 write.csv(simulated_life_exp, "outputs/data/simulated_life_exp.csv", row.names = FALSE)
 
+# Confirm file creation
+file.exists("outputs/data/simulated_life_exp.csv")
 
 
 # Simulate the Vote at Each State
 set.seed(123)
+
 # Define the state names
 state_names <- c(
   "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", "Georgia",
@@ -64,35 +67,23 @@ state_names <- c(
   "Wisconsin", "Wyoming"
 )
 
-# Function to simulate vote data for a single state
-simulate_state_votes <- function() {
-  total_votes <- round(runif(1, 50000, 1000000))  # Simulate total votes (between 50,000 and 1,000,000)
-  votes_gop <- round(runif(1, 0.4 * total_votes, 0.6 * total_votes))  # Simulate GOP votes (between 40% and 60% of total votes)
-  votes_dem <- total_votes - votes_gop  # Calculate Democratic votes
-  diff <- votes_gop - votes_dem  # Calculate the difference between GOP and Democratic votes
-  per_gop <- runif(1, 0.4, 0.6)  # Simulate GOP vote share (between 40% and 60%)
-  per_dem <- 1 - per_gop  # Calculate Democratic vote share
-  per_point_diff <- abs(per_gop - per_dem)  # Calculate absolute difference in vote share
-  c(votes_gop, votes_dem, total_votes, diff, per_gop, per_dem, per_point_diff)
-}
+# Generate simulated vote data
+simulated_vote_data <- tibble(
+  state_name = rep(state_names, each = 1),
+  total_votes = round(runif(length(state_names), min = 50000, max = 1000000), 0),  # Simulate total votes (between 50,000 and 1,000,000)
+  votes_gop = round(runif(length(state_names), min = 0.4 * total_votes, max = 0.6 * total_votes), 0),  # Simulate GOP votes (between 40% and 60% of total votes)
[…]

### 2024-02-15 scripts/01-simulate_data.R
Added tests to simulation

@@ -19,16 +19,6 @@ set.seed(123)
 # Define the years
 years <- seq(2006, 2021)
 
-# Define life expectancy ranges for each group
-life_expectancy_ranges <- list(
-  "All" = c(77, 80),
-  "Hispanic" = c(80, 83),
-  "Non-Hispanic AIAN" = c(65, 72),
-  "Non-Hispanic Asian" = c(83, 87),
-  "Non-Hispanic Black" = c(70, 76),
-  "Non-Hispanic White" = c(77, 81)
-)
-
 # Define function to generate random numbers within a range
 generate_random_number <- function(range, n) {
   return(round(runif(n, min = range[1], max = range[2]), 1))
@@ -37,6 +27,7 @@ generate_random_number <- function(range, n) {
 # Generate simulated life expectancy data
 simulated_life_exp <- tibble(
   year = years,
+  # Every group has a different life expectancy range
   All = generate_random_number(c(77, 80), n = length(years)),
   Hispanic = generate_random_number(c(80, 83), n = length(years)),
   Non_Hispanic_AIAN = generate_random_number(c(65, 72), n = length(years)),
@@ -46,6 +37,40 @@ simulated_life_exp <- tibble(
 )
 
 print(head(simulated_life_exp))
+
+# Test that there are 7 columns
+num_cols <- ncol(simulated_life_exp)
+expected_num_cols <- 7
+test_num_cols <- num_cols == expected_num_cols
+test_num_cols
+
+# Test that there are 16 years
+simulated_life_exp$year |> length() == 16
+
+# Test that All is between 77 and 80
+simulated_life_exp$All |> min() >= 77
+simulated_life_exp$All |> max() <= 80
+
+# Test that Hispanic is between 80 and 83
+simulated_life_exp$Hispanic |> min() >= 80
+simulated_life_exp$Hispanic |> max() <= 83
+
+# Test that Non_Hispanic_AIAN is between 65 and 72
+simulated_life_exp$Non_Hispanic_AIAN |> min() >= 65
+simulated_life_exp$Non_Hispanic_AIAN |> max() <= 72
+
+# Test that Non_Hispanic_Asian is between 83 and 87
+simulated_life_exp$Non_Hispanic_Asian |> min() >= 83
+simulated_life_exp$Non_Hispanic_Asian |> max() <= 87
+
+# Test that Non_Hispanic_Black is between 70 and 76
[…]

### 2024-02-15 scripts/02-test_data.R

@@ -61,9 +61,9 @@ num_unique_states <- length(unique(data2$state_name))
 expected_num_states <- 50
 
 if (num_unique_states == expected_num_states) {
-  print("Test 4 passed: Data includes data for all 50 states.")
+  print("Test 3 passed: Data includes data for all 50 states.")
 } else {
-  print("Test 4 failed: Data does not include data for all 50 states.")
+  print("Test 3 failed: Data does not include data for all 50 states.")
 }
 
 

### 2024-02-15 outputs/paper/paper.qmd
modified data section and introduction

@@
 
-Our analysis is grounded in comprehensive data sourced from the paper, `Why Did the Best Prepared Country in
-the World Fare So Poorly during COVID?` [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
+Our analysis is grounded in comprehensive data sourced from the paper, "Why Did the Best Prepared Country in the World Fare So Poorly during COVID?" [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
 While leveraging the initial findings from the @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities.
@@
 As we navigate the complexities of the COVID-19 pandemic, it is imperative to not only understand the factors contributing to disparities but also to identify actionable strategies for mitigating these inequities and strengthening community resilience.
-In the subsequent sections of this paper, we first delve deeper into our methodology, data cleaning, and data measurement.
-Next, we present our findings in detail with graphs and tables.
-Finally, we discuss how these results came about and the implications of our research for public health policy and practice.
+In the subsequent sections of this paper, we first delve deeper into our methodology, data cleaning, and data measurement in @sec-data.
+Next, we present our findings in detail with graphs and tables in @sec-results.
+Finally, we discuss how these results came about and the implications of our research for public health policy and practice @sec-discussion and @sec-conclusion.
 By elucidating these critical relationships and expanding upon prior research, our study contributes to the broader discourse on pandemic preparedness and underscores the importance of tailored interventions in addressing underlying structural determinants of health disparities.
@@
 The datasets utilized in the `original paper` [@citearticle] and this paper were mainly compiled from various sources such as the @data1, @data2, @data3, @data4, @data5, @data6, @data7, @data8, @data9, @data10, and @data11.
-Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and [@citedata2].
+Additionally, to address the original paper's lack of US COVID statistics and political party support data, we incorporated information from @citedata1 and @citedata2.
 The political party support data we utilized from @citedata1 was compiled from @data12.
@@
 
-This paper will replicate the data that was originally presented by @citearticle, as previously mentioned.
-We then transform the replication to our data analysis incorporating the COVID statistics and political party support data.
-We seek to use our initial replication and following own data analysis to provide a fresh part of the original paper, enhancing the with more data.
-Different from the original paper which spent a lot of time talking about how badly prepared the US was against other countries, We also specifically diverge our focus to the political and socioeconomic factors that impacted the US's COVID outcome.
+This paper aims to replicate the data originally presented by @citearticle, as mentioned previously.
+Subsequently, we will integrate this replication into our own data analysis, incorporating COVID statistics and political party support data.
+Our objective is to leverage both the initial replication and our subsequent analysis to contribute a fresh perspective to the original paper, enriching it with additional data.
+In contrast to the original paper, which extensively discussed the inadequacies of US preparedness compared to other countries, we specifically diverge our focus to examine the political and socioeconomic factors that influenced the US's COVID outcomes.
 
@@
 
-The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for life expectancy rates [@citedefenition], and political affiliation.
+The transformation of real-world phenomena into quantifiable data points in our dataset is influenced by various factors, leading to inherent limitations in our  measurements for COVID death rates, life expectancy rates [@citedefenition], and political affiliation.
 
-Our measurement for COVID death rates is drawn from a accumulated dataset last updated until March 2023.
+The determination of whether someone is deceased and whether they had COVID involves several steps and considerations.
+Medical professionals typically confirm a person's death through various means, including clinical examination, vital signs assessment, and sometimes laboratory tests.
+For COVID-related deaths, individuals are often tested for the presence of the virus through diagnostic tests such as PCR or antigen tests.
+Additionally, medical history and symptoms consistent with COVID-19 may also contribute to the determination of COVID-related fatalities.
+However, it's important to note that not all COVID-related deaths may be accurately identified or reported, leading to potential underestimation or misclassification in official statistics.
+
+Incorporating this understanding into our analysis, our measurement for COVID death rates is drawn from a accumulated dataset last updated until March 2023.
 This temporal constraint means we would not be aware of the fluctuations of death rates over time, particularly during the peak of the pandemic in 2020.
@@
 
-Similarly, our measurement of life expectancy suffer from the demographic limitations.
-Even though we categorize life expectancy by race, the absence of an age breakdown restricts us discern impacts across different age groups.
+Similarly, our measurement of life expectancy suffers from demographic limitations.
+Life expectancy, defined as the estimate of the average age that members of a particular population group will be when they die, is a crucial metric in understanding population health dynamics [@citedefenition].
+Even though we categorize life expectancy by race, the absence of an age breakdown restricts our ability to discern impacts across different age groups.
 Moreover, delays in reporting deaths undermine the timeliness and accuracy of our data, potentially skewing our understanding of population health trends.
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-# Conclusion
+# Conclusion {#sec-conclusion}
 

### 2024-02-15 outputs/paper/paper.qmd
fixed weakness

@@
 
-In our research, we identified several weaknesses that underscored the gap between the U.S.'s high pandemic preparedness ranking and its actual response to the COVID-19 crisis.
-Notably, there was a significant underutilization of preparedness capacities, as the resources and infrastructures in place were not fully mobilized or applied effectively during the pandemic.
-This disconnect highlights a critical area for improvement in aligning preparedness with real-time response capabilities.
-Furthermore, the response inadequately addressed intrinsic social vulnerabilities, exacerbating disparities in healthcare access and outcomes among different racial and socioeconomic groups, and underscoring the need for more inclusive and equitable public health strategies.
-The heavy politicization of the pandemic response, coupled with inconsistent messaging from health authorities, significantly eroded public trust and compliance, emphasizing the importance of depoliticizing public health measures and enhancing communication strategies.
-Additionally, shortcomings in data and surveillance systems, including delays in testing, inadequate genetic sequencing, and a lack of comprehensive surveillance, hindered informed decision-making and targeted interventions, pointing to a crucial need for investment in data infrastructure and capabilities.
-The analysis also suggests that the U.S.
-could benefit from more robust international benchmarking and learning from the diverse responses of other countries to health emergencies, which would involve understanding different strategies and their effectiveness beyond mere outcome comparisons.
-
-To address these challenges and strengthen the nation's resilience to future health emergencies, several next steps are recommended.
-Strengthening the linkages between preparedness and response is imperative, ensuring that capacities are not only available but also readily deployable and adaptable to the dynamics of a health crisis.
-This includes fostering agile and responsive systems capable of rapid mobilization.
-Addressing the social determinants of health is essential, with strategies aimed at mitigating the impact of social vulnerabilities through equitable healthcare access, support for marginalized communities, and targeted protective measures for vulnerable populations.
-Enhancing communication and public trust is vital, necessitating clear, consistent, and transparent communication from health authorities, alongside efforts to depoliticize health measures.
-Improving data and surveillance systems is critical to providing real-time insights, expanding genetic sequencing capabilities, and establishing standardized data protocols.
-Lastly, actively engaging in global health networks to exchange experiences, learn from global successes and failures, and collaborate on best practices is crucial for a more effective and equitable response that maximizes the full potential of preparedness capacities.
+In our research, we identified several weaknesses in our data that is worth mentioning.
+The pandemic highlighted significant disparities in global surveillance, with countries possessing advanced healthcare systems paradoxically reporting higher COVID-19 mortality rates due to more accurate detection capabilities, underscoring the need for universal surveillance standards to accurately monitor diseases worldwide.
+The United States, in particular, faced challenges with delayed data reporting, impeding swift public health responses and decision-making.
+The effects of this is self-evident as it makes us question if the data we are given is even correct.
+Furthermore, the often incomplete demographic data obscured the unequal effects of the pandemic on diverse communities, preventing targeted interventions for those most at risk, such as racial and ethnic minorities, thereby deepening existing health inequities.
+Additionally, the pandemic revealed a critical lack of infrastructure capable of integrating various data types, from clinical to public health surveillance, limiting a comprehensive understanding of the pandemic's scope and the impact of measures taken.
+Addressing these data weaknesses necessitates establishing global data collection, reporting, and sharing standards, including case definitions, demographic information, and data formats to ensure consistency across borders.
+Investments in digital health technologies are essential for enabling real-time surveillance while maintaining privacy and security.
+Policies should be enacted to ensure the collection of detailed demographic data to identify and address health disparities, ensuring equitable intervention distribution.
+Developing frameworks that allow for the seamless integration of diverse data sources will offer a more detailed pandemic perspective, necessitating cross-sector collaboration.
+Enhancing data literacy among health professionals, policymakers, and the public will foster informed decision-making and trust in health interventions.
+Lastly, international collaboration in data sharing and analysis is crucial for the rapid dissemination of insights and best practices, bolstering global preparedness and response for future health crises.
 

### 2024-02-15 outputs/paper/paper.qmd
fixed format

@@
       format = "markdown", 
-      col.names = c("State Name", "County FIPS", "County Name", "Votes GOP", "Votes DEM", "Total Votes", "Difference", "Percentage GOP", "Percentage DEM", "Percentage Point Difference")
+      col.names = c("State Name", "County  FIPS", "County Name", "Votes GOP", "Votes DEM", "Total Votes", "Differ- ence", "Percent- age GOP", "Percent- age DEM", "Percentage Point Difference")
 )

### 2024-02-15 outputs/paper/paper.qmd
fixed discussions

@@
 
-Various studies and our results have shown that areas with higher support for one political party exhibited distinct behaviors and compliance levels with health recommendations, which directly correlated with COVID-19 case rates and mortality.
-A news article from `ABC News` [@citesource1] shows that the top states with the highest COVID-19 deaths are Arizona, and Washington with 581 deaths and 526 deaths respectively per 100,000 people.
-According to 2020 presidential voting data published by `CNN`, we have both states having the electoral vote of Democrat with Washington winning by 58% [@citesource2] and Arizona winning by 49.4% [@citesource3].
-Another news article by `ContagionLive` [@citesource4] also claims both Arizona and Washington have the highest COVID-19 mortality.
-This polarization has not only influenced individual behavior but also shaped state and local health policies, further entrenching the disparities in health outcomes.
+Various studies and our results have consistently demonstrated a correlation between political affiliation and adherence to health guidelines during the COVID-19 pandemic.
+This correlation extends to distinct behaviors and compliance levels with health recommendations, directly impacting COVID-19 case rates and mortality.
+One study, in particular, revealed a noteworthy trend: while Democratic-leaning states initially faced higher cases and death rates at the onset of the pandemic, Republican-leaning states ultimately suffered more severely throughout the latter half of 2020 [@march 11_2021_2021].
+This shift in severity can be attributed to several factors.
+Initially, many Democratic states were the epicenter of the virus, experiencing high case and death rates due to the initial entry of the virus into these regions.
+However, as the pandemic progressed, Republican-leaning states saw a surge in cases, partially due to their relatively loose COVID-19 regulations, such as the lack of strict mask requirements [@march 11_2021_2021].
 
@@
 Initially, inconsistencies in recommendations, such as on mask usage, challenged the principle of clear, consistent, and science-based communication essential for an effective public health response.
+
 Moreover, the transparency of government actions and decision-making processes is crucial in building and maintaining public trust, especially during health crises.
 The level of public trust was greatly affected by the openness and accuracy with which governments, at all levels, communicated about the evolving situation, the reasoning behind guidelines, and the measures taken to combat the virus, emphasizing the importance of transparent reporting of data related to case counts, hospitalizations, vaccine distribution, and side effects.
+
 Furthermore, consistent communication from public health officials and government leaders is key to ensuring adherence to health guidelines, where inconsistencies, such as changes in mask-wearing guidelines without clear explanations, have led to public confusion.
 The direct correlation between government transparency, consistent communication, and public behavior is self-evident, with populations receiving clear and transparent information being more likely to adhere to guidelines, participate in testing and tracing efforts, and accept vaccination.
+
 Drawing lessons from the pandemic, strategies for improving government transparency and communication in future health emergencies should include establishing centralized information hubs, ensuring regular and predictable communication from health authorities, engaging community leaders in information dissemination, and harnessing digital platforms and social media to amplify public health messages, thus reinforcing public trust and compliance.
@@
 The COVID-19 pandemic starkly highlighted how social vulnerabilities and disparities in healthcare access exacerbated the impact of global health crises, contributing to significant variations in disease outcomes and underscoring the need for targeted public health strategies that address these disparities' root causes.
+
 Social vulnerabilities, such as socioeconomic status, race, ethnicity, and housing conditions, critically determined COVID-19 outcomes' severity, with populations in crowded housing, limited access to sanitation, and lower socioeconomic brackets experiencing higher transmission rates due to social distancing and hygiene maintenance challenges.
 An independent study done by the `Government of Canada` states that it "identified that the risk of COVID-19 related deaths in Black, Asian and minority ethnic groups was nearly 1.5 times higher than White individuals" [@citesource5].
-Another news article done by the `MSN` [@citesource6] states that 'racial minority participants reported more negative impacts on health status, activity, and absence from work as compared to the White population.' The pandemic's economic toll further limited these groups' healthcare access, amplifying vulnerabilities.
+Another news article done by the `MSN` [@citesource6] states that 'racial minority participants reported more negative impacts on health status, activity, and absence from work as compared to the White population.'
+
+The pandemic's economic toll further limited these groups' healthcare access, amplifying vulnerabilities.
 Disparities in healthcare access played a significant role in influencing COVID-19 morbidity and mortality, with communities facing healthcare facility shortages, provider scarcities, and barriers due to insurance or financial constraints at heightened risk.
 These disparities were evident in the uneven vaccine distribution and access, highlighting the advantages of regions with strong healthcare infrastructure.
+
 Marginalized populations, including racial and ethnic minorities, faced compounded risks from social vulnerabilities and healthcare disparities, evidenced by higher infection, hospitalization, and death rates due to factors like essential service employment and prevalent pre-existing conditions.
@@
 Despite their crucial roles, they often lack adequate pay, protection, and respect.
+
 Addressing these disparities in future pandemics requires public health strategies that prioritize equity and inclusivity, including community-based healthcare investments, enhanced vulnerable community outreach, and equitable healthcare resource access policies.
@@
 Equally important is enhancing data standardization and interoperability through universal standards like FHIR (Fast Healthcare Interoperability Resources) to facilitate efficient data sharing [@citesource8].
+
 Another notable example is the Digital Health Platform (DHP) [@citedhp] developed by the European Centre for Disease Prevention and Control (ECDC) [@citeecdc], which facilitates the real-time exchange and analysis of health data across EU member states.
 This platform has been instrumental in coordinating responses to health threats, including the COVID-19 pandemic, by providing timely and accurate data to inform public health decisions.
+
 Another digital platform worth mentioning is the HealthMap system [@citehealthmap], which utilizes online data sources and machine learning algorithms to detect and track global disease outbreaks.
 HealthMap's ability to aggregate and visualize data from various sources, including news reports and social media, has proven invaluable in early outbreak detection and response, showcasing the power of digital surveillance in public health.
+
 Investing in digital surveillance systems, which employ AI and machine learning to sift through diverse data sources for early outbreak detection, is essential for rapid response to health threats.
@@
 Furthermore, fostering public-private partnerships can harness the agility of the private sector and the public health expertise of governmental agencies to enhance data analytics capabilities.
+
 Ensuring the privacy and security of health data through robust governance frameworks and advanced encryption is paramount to maintaining public trust.
 Engaging communities in these initiatives ensures their relevance and fosters trust while building global data-sharing networks encourages international collaboration, crucial for a concerted response to pandemics.
+
 Collectively, these strategies are fundamental to bolstering public health decision-making and preparedness, making our health systems more resilient against the challenges posed by emerging infectious diseases.
@@
 The pandemic highlighted significant disparities in global surveillance, with countries possessing advanced healthcare systems paradoxically reporting higher COVID-19 mortality rates due to more accurate detection capabilities, underscoring the need for universal surveillance standards to accurately monitor diseases worldwide.
+
 The United States, in particular, faced challenges with delayed data reporting, impeding swift public health responses and decision-making.
@@
 Furthermore, the often incomplete demographic data obscured the unequal effects of the pandemic on diverse communities, preventing targeted interventions for those most at risk, such as racial and ethnic minorities, thereby deepening existing health inequities.
+
 Additionally, the pandemic revealed a critical lack of infrastructure capable of integrating various data types, from clinical to public health surveillance, limiting a comprehensive understanding of the pandemic's scope and the impact of measures taken.
 Addressing these data weaknesses necessitates establishing global data collection, reporting, and sharing standards, including case definitions, demographic information, and data formats to ensure consistency across borders.
+
 Investments in digital health technologies are essential for enabling real-time surveillance while maintaining privacy and security.
@@
 Developing frameworks that allow for the seamless integration of diverse data sources will offer a more detailed pandemic perspective, necessitating cross-sector collaboration.
+
 Enhancing data literacy among health professionals, policymakers, and the public will foster informed decision-making and trust in health interventions.
@@
 This discrepancy is attributed to several critical factors highlighted in the paper.
+
 The U.S.
@@
 Additionally, inherent vulnerabilities such as the significant portion of the population in congregated settings like nursing homes and prisons increased susceptibility to virus spread and severe health outcomes.
+
 The politicization of the pandemic response further compounded these issues, resulting in varied adherence to public health guidelines across states and political lines, thereby undermining the response's effectiveness.

### 2024-02-15 outputs/paper/paper.qmd
fixed ref

@@
 This correlation extends to distinct behaviors and compliance levels with health recommendations, directly impacting COVID-19 case rates and mortality.
-One study, in particular, revealed a noteworthy trend: while Democratic-leaning states initially faced higher cases and death rates at the onset of the pandemic, Republican-leaning states ultimately suffered more severely throughout the latter half of 2020 [@march 11_2021_2021].
+One study, in particular, revealed a noteworthy trend: while Democratic-leaning states initially faced higher cases and death rates at the onset of the pandemic, Republican-leaning states ultimately suffered more severely throughout the latter half of 2020 [@march_11_2021_2021].
 This shift in severity can be attributed to several factors.
 Initially, many Democratic states were the epicenter of the virus, experiencing high case and death rates due to the initial entry of the virus into these regions.
-However, as the pandemic progressed, Republican-leaning states saw a surge in cases, partially due to their relatively loose COVID-19 regulations, such as the lack of strict mask requirements [@march 11_2021_2021].
+However, as the pandemic progressed, Republican-leaning states saw a surge in cases, partially due to their relatively loose COVID-19 regulations, such as the lack of strict mask requirements [@march_11_2021_2021].
 

### 2024-02-15 scripts/01-simulate_data.R
Updated Simulation

@@ -71,6 +71,9 @@ simulated_life_exp$Non_Hispanic_Black |> max() <= 76
 simulated_life_exp$Non_Hispanic_White |> min() >= 77
 simulated_life_exp$Non_Hispanic_White |> max() <= 81
 
+# Check if all values are numeric
+numeric_check <- all(sapply(simulated_life_exp, is.numeric))
+print(paste("Numeric Check:", ifelse(numeric_check, "Passed", "Failed")))
 
 # Save the dataset to a CSV file
 write.csv(simulated_life_exp, "outputs/data/simulated_life_exp.csv", row.names = FALSE)
@@ -119,6 +122,22 @@ expected_num_cols <- 8
 test_num_cols <- num_cols == expected_num_cols
 test_num_cols
 
+# Check for missing values
+if (!anyNA(data2)) {
+  print("Test passed: No missing values found.")
+} else {
+  print("Test failed: Missing values detected.")
+}
+
+# Check for all 50 states
+num_unique_states <- length(unique(data2$state_name))
+expected_num_states <- 50
+
+if (num_unique_states == expected_num_states) {
+  print("Test passed: Data includes data for all 50 states.")
+} else {
+  print("Test failed: Data does not include data for all 50 states.")
+}
 
 # Write simulated vote data to CSV
 write.csv(simulated_vote_data, "outputs/data/simulated_vote_data.csv", row.names = FALSE)

### 2024-02-15 outputs/paper/paper.qmd
Updated intro

@@
 Understanding these dynamics is crucial for informing effective public health interventions and fostering resilience in the face of future pandemics.
-In this paper, building upon prior research from the American Economic Association, we delve into the multifaceted relationship between local health infrastructure, socio-political factors, and COVID-19 outcomes, aiming to provide insights into the mechanisms driving disparities and identify strategies for targeted intervention.
+In this paper, building upon prior research, we delve into the multifaceted relationship between local health infrastructure, socio-political factors, and COVID-19 outcomes, aiming to provide insights into the mechanisms driving disparities and identify strategies for targeted intervention.
 
-Our analysis is grounded in comprehensive data sourced from the paper, "Why Did the Best Prepared Country in the World Fare So Poorly during COVID?" [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
-While leveraging the initial findings from the @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities.
-We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access, housing conditions, and ethnic demographics in shaping disparities in COVID-19 outcomes.
+Our analysis is grounded in comprehensive data sourced from the paper, **"Why Did the Best Prepared Country in the World Fare So Poorly during COVID?"** [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
+While leveraging the initial findings from @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities.
+We particularly focus on the impact of political polarization on adherence to health guidelines and the role of social vulnerabilities such as healthcare access and housing conditions on the life expectancy of various racial and ethnic groups.
 One key finding that emerges from our study is the influence of political polarization, with Republican-leaning states exhibiting slightly higher clustering around elevated death rates, highlighting the importance of socio-political factors in shaping public health responses.
@@
 [@fig-GHS] gives us some context to motivate our research.
-This map allows us to visualize the Global Health Security (GHS) Index which is designed to assess a country's capabilities to prevent, detect, and respond to significant infectious disease outbreaks.
-A higher Global Health Security (GHS) Index score indicates a stronger health system capacity for pandemic readiness.
-The GHS Index assesses the readiness of countries for pandemics and other significant biological threat emergencies, using a comprehensive framework that includes measurements grouped into six broad categories: prevention, detection, response, health system capacity, compliance with international norms, and risk environment.
+This map allows us to visualize the Global Health Security (GHS) Index which is designed to assess a country's capabilities to handle significant infectious disease outbreaks.
+A higher on the GHS Index indicates a stronger health system capacity for pandemic readiness.
+The GHS Index evaluates countries' readiness for pandemics and other significant biological threat emergencies, using a comprehensive framework.
+This framework includes measurements grouped into six broad categories: prevention, detection, response, health system capacity, compliance with international norms, and risk environment.
 Higher scores suggest that a country has a more robust health system capacity, with better capabilities to prevent, detect, respond to, and mitigate the spread of an epidemic.
@@
 As we navigate the complexities of the COVID-19 pandemic, it is imperative to not only understand the factors contributing to disparities but also to identify actionable strategies for mitigating these inequities and strengthening community resilience.
-In the subsequent sections of this paper, we first delve deeper into our methodology, data cleaning, and data measurement in @sec-data.
-Next, we present our findings in detail with graphs and tables in @sec-results.
-Finally, we discuss how these results came about and the implications of our research for public health policy and practice @sec-discussion and @sec-conclusion.
+In the subsequent sections of this paper, we outline our research process, findings, and implications as follows: @sec-data delves into our methodology, encompassing data collection, cleaning, and measurement techniques.
+In @sec-results, we present detailed findings accompanied by graphs, tables, and other visual aids for clarity.
+Following this, @sec-discussion engages in a discussion on the significance and implications of our results for public health policy and practice, addressing unexpected findings, limitations, and avenues for further research.
+Finally, @sec-conclusion offers a conclusion and recommendations, summarizing key insights and providing actionable recommendations for policymakers, practitioners, and future research directions.
 By elucidating these critical relationships and expanding upon prior research, our study contributes to the broader discourse on pandemic preparedness and underscores the importance of tailored interventions in addressing underlying structural determinants of health disparities.

### 2024-02-15 outputs/paper/paper.qmd
added estimand paragraph in intro , deleted files

@@
 In this paper, building upon prior research, we delve into the multifaceted relationship between local health infrastructure, socio-political factors, and COVID-19 outcomes, aiming to provide insights into the mechanisms driving disparities and identify strategies for targeted intervention.
+
+Our analysis seeks to estimate several key parameters, including the impact of political polarization on adherence to health guidelines, the association between social vulnerabilities (such as healthcare access, housing conditions, and ethnic demographics) and disparities in COVID-19 outcomes, and the influence of community health systems on disease trajectories.
+These estimands guide our methodology, data analysis, and interpretation of results, allowing us to comprehensively address the research questions.
 

### 2024-02-15 outputs/paper/paper.qmd
updated paper and llm

@@
 From all the data provided from @citearticle, @citedata1, and @citedata2, we mainly utilize  @data3, @data8, @data12, and @citedata2.
-The data from @citearticle is well-organized and require minimal cleaning for analysis.
+The data from @citearticle is well-organized and requires minimal cleaning for analysis.
 
@@
 We utilize the "group_by" function to generate new variables at the state level, calculating the total votes, total Republican party votes, total Democratic party votes, and the total vote difference between the two parties by county.
-Additionally, We also create variables for the average percentage of votes each party receives and the difference by each state.
+We also create variables for the average percentage of votes each party receives and the difference by each state.
 
@@
 Subsequently, to find the top 10 states with the highest death rates, we rank the states based on death rate.
-Lastly, to create the final figure that incorporates the COVID death rate and people's political preferences by state, we merge the two datasets from @data12 and @citedata2 and proceed with our analysis.
+Lastly, to create the final figure incorporating the COVID death rate and people's political preferences by state, we merge the two datasets from @data12 and @citedata2 and proceed with our analysis.
 
@@
 
-Incorporating this understanding into our analysis, our measurement for COVID death rates is drawn from a accumulated dataset last updated until March 2023.
+Incorporating this understanding into our analysis, our measurement for COVID death rates is drawn from an accumulated dataset last updated until March 2023.
 This temporal constraint means we would not be aware of the fluctuations of death rates over time, particularly during the peak of the pandemic in 2020.
-The data's primarily reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduces bias as excludes uncounted deaths from COVID.
-Especially during the critical period of health facilities capacity constraints,many individuals died outside of medical facilities without being documented.
-Therefore, our measurements for COVID death rates are only accounted for the documented death from COVID.
+The data's primarily reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduce bias as excludes uncounted deaths from COVID.
+Especially during the critical period of health facility capacity constraints, many individuals died outside of medical facilities without being documented.
+Therefore, our measurements for COVID death rates only accounted for the documented deaths from COVID.
 
-In addition,it's important to note that the death rate presented here reflects the entire population of a state.
+In addition, it's important to note that the death rate presented here reflects the entire population of a state.
 However, COVID-19 disproportionately affects older individuals compared to younger adults in the prime of their lives.
 As a result, this data does not capture the effect of COVID across different age groups.
-Instead, it only offers a general overview the state's population.
+Instead, it only offers a general overview of the state's population.
 In summary, we are unable to provide detailed insights into age-specific or temporal variations based on this dataset.
@@
 Our measurement for each state's political preference is calculated from the 2020 election votes.
-There may be many scenarios that prevented eligible voters from voting.
-However, we are only able to assess the the state's political preference based on the people who voted, thereby potentially introducing selection bias.
+There may be many scenarios that prevent eligible voters from voting.
+However, we can only assess the state's political preference based on the people who voted, thereby potentially introducing selection bias.
 Additionally, issues such as ballot loss or voter disenfranchisement further compound the reliability of our measurements.
@@
 Unfortunately, due to the unavailability of data over time for Asian, American Indian, and Alaska Native communities, they have been excluded from the time series graph.
-@tbl-life-exp provides a more detailed breakdown of the life expectancy before and during COVID for Asian, American Indian, Alaska Native and other racial communities.
+@tbl-life-exp provides a more detailed breakdown of the life expectancy before and during COVID-19 for Asian, American Indian, Alaska Native, and other racial communities.
 
@@
 
-Various studies have confirmed the existence of correlation between a state's political affiliation and its handling of COVID issues (@datz_2022).
+Various studies have confirmed the existence of a correlation between a state's political affiliation and its handling of COVID issues [@datz_2022].
 To testify to the claim, we collected the voting data for each of the 50 states for the 2020 presidential election.
@@
 Notably, We found that six of the ten top Republican states made a reappearance in the top death rates table; these states are Oklahoma, West Virginia, Arkansas, Alabama, Tennessee, and Kentucky.
-The higher mortality rates observed in Republican-leaning states may be attributed to their preference for less stringent measures compared to Democratic-leaning states, which tend to favor stricter measures (@vandusky-allen_shvetsova_2021).
+The higher mortality rates observed in Republican-leaning states may be attributed to their preference for less stringent measures compared to Democratic-leaning states, which tend to favor stricter measures [@vandusky-allen_shvetsova_2021].
 
@@
 These factors can be broken down into the influence of political polarization, the impact of government transparency which ties into the effect it also had on public trust, and the role of social vulnerabilities and healthcare access disparities.
-We should also look at how to address these potential factors by implementing strategies in the future in regards to data collection and sharing.
+We should also look at how to address these potential factors by implementing strategies in the future regarding data collection and sharing.
 It is important to also note we were able to identify some weaknesses in our research that should be evaluated in the future.
@@
 Investing in digital surveillance systems, which employ AI and machine learning to sift through diverse data sources for early outbreak detection, is essential for rapid response to health threats.
-An news article by `nature` [@citegoogle] talks about how Google has made an AI which claims to be "better bedside manner than human doctors — and makes better diagnoses".
+A news article by `nature` [@citegoogle] talks about how Google has made an AI that claims to be "better bedside manner than human doctors — and makes better diagnoses".
 Furthermore, fostering public-private partnerships can harness the agility of the private sector and the public health expertise of governmental agencies to enhance data analytics capabilities.
@@
 
-In our research, we identified several weaknesses in our data that is worth mentioning.
+In our research, we identified several weaknesses in our data that are worth mentioning.
 The pandemic highlighted significant disparities in global surveillance, with countries possessing advanced healthcare systems paradoxically reporting higher COVID-19 mortality rates due to more accurate detection capabilities, underscoring the need for universal surveillance standards to accurately monitor diseases worldwide.
@@
 The United States, in particular, faced challenges with delayed data reporting, impeding swift public health responses and decision-making.
-The effects of this is self-evident as it makes us question if the data we are given is even correct.
+The effects of this are self-evident as it makes us question if the data we are given is even correct.
 Furthermore, the often incomplete demographic data obscured the unequal effects of the pandemic on diverse communities, preventing targeted interventions for those most at risk, such as racial and ethnic minorities, thereby deepening existing health inequities.
@@
 Enhancing data literacy among health professionals, policymakers, and the public will foster informed decision-making and trust in health interventions.
-Lastly, international collaboration in data sharing and analysis is crucial for the rapid dissemination of insights and best practices, bolstering global preparedness and response for future health crises.
+Lastly, international collaboration in data sharing and analysis is crucial for the rapid dissemination of insights and best practices, bolstering global preparedness and response to future health crises.
 

### 2024-02-15 outputs/paper/paper.qmd
Final Edits

@@
 
-Our analysis is grounded in comprehensive data sourced from the paper, **"Why Did the Best Prepared Country in the World Fare So Poorly during COVID?"** [@citearticle], enabling a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
+Grounded in comprehensive data sourced from the paper, **"Why Did the Best Prepared Country in the World Fare So Poorly during COVID?"** [@citearticle], our analysis offers a nuanced exploration of how community health systems and socio-political contexts intersect to influence disease outcomes.
 While leveraging the initial findings from @citearticle, we expand the scope of inquiry to investigate broader implications and unearth deeper insights into the complexities of COVID-19 disparities.
@@
 Following this, @sec-discussion engages in a discussion on the significance and implications of our results for public health policy and practice, addressing unexpected findings, limitations, and avenues for further research.
-Finally, @sec-conclusion offers a conclusion and recommendations, summarizing key insights and providing actionable recommendations for policymakers, practitioners, and future research directions.
+Finally, @sec-conclusion offers a conclusion that synthesizes key insights from our analysis.
 By elucidating these critical relationships and expanding upon prior research, our study contributes to the broader discourse on pandemic preparedness and underscores the importance of tailored interventions in addressing underlying structural determinants of health disparities.
@@
 This temporal constraint means we would not be aware of the fluctuations of death rates over time, particularly during the peak of the pandemic in 2020.
-The data's primarily reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduce bias as excludes uncounted deaths from COVID.
+The data's primary reliance on the reported statistics from hospitals, clinics, testing centres, and laboratories may introduce bias as it excludes uncounted deaths from COVID.
 Especially during the critical period of health facility capacity constraints, many individuals died outside of medical facilities without being documented.
@@
 
-Manual Data Extraction involved careful review of reports and tables to ensure accurate data capture, particularly for datasets like life expectancy where this meticulous approach was paramount; this was coupled with Standardization and Adjustment processes where excess death data and age-standardized excess death rates underwent significant processing, including the comparison of observed deaths during the pandemic to expected deaths based on historical trends, with adjustments made for the age distribution of the population; concurrently, Geospatial Data Handling was employed for mapping the GHS Index scores, utilizing Geographic Information Systems (GIS) to visually represent pandemic preparedness across various countries, and this was complemented by Real-Time Data Aggregation, which entailed aggregating COVID-19 death data in real-time from authoritative sources such as the Johns Hopkins University Dashboard, necessitating continuous data monitoring and updates to maintain accuracy and relevance.
+Manual Data Extraction involved careful review of reports and tables to ensure accurate data capture, particularly for datasets like life expectancy where this meticulous approach was paramount.
+This was coupled with Standardization and Adjustment processes where excess death data and age-standardized excess death rates underwent significant processing, including the comparison of observed deaths during the pandemic to expected deaths based on historical trends, with adjustments made for the age distribution of the population.
+Concurrently, Geospatial Data Handling was employed for mapping the GHS Index scores, utilizing Geographic Information Systems (GIS) to visually represent pandemic preparedness across various countries, and this was complemented by Real-Time Data Aggregation, which entailed aggregating COVID-19 death data in real-time from authoritative sources such as the Johns Hopkins University Dashboard, necessitating continuous data monitoring and updates to maintain accuracy and relevance.
 
@@
 As an alternative, we employed death rates per 100,000 people as a metric for evaluating each state's COVID preparedness and situation.
-We then produced @tbl-deaths that rank the top ten states with the highest death rates from COVID-19 per 100,000 people to examine the potential correlation between party preferences and COVID-related deaths.
-Notably, We found that six of the ten top Republican states made a reappearance in the top death rates table; these states are Oklahoma, West Virginia, Arkansas, Alabama, Tennessee, and Kentucky.
+We then produced @tbl-deaths, which ranks the top ten states with the highest death rates from COVID-19 per 100,000 people, to examine the potential correlation between party preferences and COVID-related deaths.
+Notably, we found that six of the ten top Republican states made a reappearance in the top death rates table; these states are Oklahoma, West Virginia, Arkansas, Alabama, Tennessee, and Kentucky.
 The higher mortality rates observed in Republican-leaning states may be attributed to their preference for less stringent measures compared to Democratic-leaning states, which tend to favor stricter measures [@vandusky-allen_shvetsova_2021].

### 2024-02-15 scripts/02-test_data.R

@@ -1,70 +1,0 @@
-#### Preamble ####
-# Purpose: Tests the Simulated Data
-# Author: Adrian Ly, Hannah Yu
-# Date: 9 February 2024 
-# Contact: adrian.ly@mail.utoronto.ca, realhannah.yu@mail.utoronto.ca, s.goel@mail.utoronto.ca
-# License: MIT
-
-#### Workspace setup ####
-library(tidyverse)
-library(knitr)
-
-
-#### Test data ####
-## Test simulated_life_data
-# Load the dataset generated in R
-data1 <- read.csv("outputs/data/simulated_life_exp.csv", stringsAsFactors = FALSE)
-
-# 1. Data Structure Check
-# Check the number of rows and columns
-structure_check <- dim(data)[1] == length(years) & dim(data)[2] == length(life_expectancy_ranges) + 1
-print(paste("Data Structure Check:", ifelse(structure_check, "Passed", "Failed")))
-
-# 2. Range Check
-# Define the expected ranges
-expected_ranges <- c(77:80, 80:83, 65:72, 83:87, 70:76, 77:81)
-# Check if all values fall within the expected ranges
-range_check <- all(sapply(data[, -1], function(x) all(x >= min(expected_ranges) & x <= max(expected_ranges))))
-print(paste("Range Check:", ifelse(range_check, "Passed", "Failed")))
-
-# 3. Numeric Check
-# Check if all values are numeric
-numeric_check <- all(sapply(data, is.numeric))
-print(paste("Numeric Check:", ifelse(numeric_check, "Passed", "Failed")))
-
-## Test simulated_vote_data
-# Load the dataset generated in R
-data2 <- read.csv("outputs/data/simulated_vote_data.csv", stringsAsFactors = FALSE)
-
-# Test 1: Check number of rows and columns
-num_rows <- nrow(data2)
-num_cols <- ncol(data2)
-
-expected_num_rows <- 50 # Expecting 50 states
-expected_num_cols <- 8  # Expecting 7 columns (plus 1 for state_name)
-
-if (num_rows == expected_num_rows && num_cols == expected_num_cols) {
-  print("Test 1 passed: Correct number of rows and columns.")
-} else {
-  print("Test 1 failed: Incorrect number of rows or columns.")
-}
-
-# Test 2: Check for missing values
-if (!anyNA(data2)) {
-  print("Test 2 passed: No missing values found.")
-} else {
-  print("Test 2 failed: Missing values detected.")
-}
-
[…]
